---
title: "Gelsemium root ATAC"
author: "Anne Pavia"
date: "2026-09-14"
output: html_document
---

>The purpose of this markdown is to provide information on how to integrate multiple 10x single cell multiome datasets across multiple replicates and tissues. In this example code, Gelsemium root multiome libraries were used to generate a single multiome object. 

```{r setup, include=FALSE}
library(tidyverse)
library(Seurat) #v5.2.1
library(SeuratObject) #v5.0.2
library(Signac) #v1.16.0
library(GenomicRanges)

set.seed(818)
```

#Step 1: Process GEX Data

##GEX Data Integration
```{r}
integrated_GEX <- readRDS("/scratch/ac05869/gelsemium_sc/seurat_data/GEL_MPI_root_cleaned_integrated.rds")
DefaultAssay(integrated_GEX)
unique(integrated_GEX@meta.data$Run)
```

#Step 2: Process ATAC Data

##Generate chromatin assays for each ATAC library and generate QC metrics.
```{r}
#Set up inputs
atac_samples <- c("GEL_BU_BQ", "GEL_BV_BR") #these are data directories for replicate samples within the data_dir path
data_dir = "/scratch/ac05869/gelsemium_sc/CR-ARC/"

#Load in features file (bed format). Features are the genomic coordinates of open chromatin regions. NOTE: In order to integrate ATAC libraries in the next step, you MUST use a common set of peaks/features for all ATAC libraries. Cat all bed files, sort, then use bedtools merge. 
peaks <- read.table(
  file = "/scratch/ac05869/gelsemium_sc/CR-ARC/merged_peaks/GEL_root.atac_peaks.sorted.merged.bed",
  col.names = c("chr","start","end")
)

#convert to genomic ranges. Necessary for the following step. 
gr_peaks <- makeGRangesFromDataFrame(peaks)

seurat_list <- list()

#path to fragment file. Fragment file contains all of the fragments and where they map to. Analogous to read alignment data for RNA-seq.
for (sample in atac_samples) {

  fragpath <- paste0(data_dir,sample,"/outs/atac_fragments.tsv.gz")

  #count fragments
  fragcounts <- CountFragments(fragments = fragpath)

  #filter fragments to only keep barcodes that more than 500 counts 
  atac.cells <- fragcounts[fragcounts$frequency_count > 500, "CB"]

  #create the fragment object
  atac.frags <- CreateFragmentObject(path = fragpath, cells = atac.cells)

  #create feature matrices. This step quantifies peaks in the library. 
  counts <- FeatureMatrix(
    fragments = atac.frags,
    features = gr_peaks,
    cells = rownames(atac.frags)
  )

  #create chrom assay and a seurat object
  chrom_assay <- CreateChromatinAssay(counts, 
                                      fragments = atac.frags)
  seurat_obj <- CreateSeuratObject(chrom_assay, 
                                   assay = "ATAC",
                                   project = sample)
  
  #include info about fragments to the assay
  rownames(fragcounts) <- fragcounts$CB
  
  
  seurat_obj$fragments <-fragcounts[colnames(seurat_obj), "frequency_count"]
  seurat_obj$mononucleosomal <- fragcounts[colnames(seurat_obj), "mononucleosomal"]
  seurat_obj$nucleosome_free <- fragcounts[colnames(seurat_obj), "nucleosome_free"]
  seurat_obj$reads_count <- fragcounts[colnames(seurat_obj), "reads_count"]

  #calculate QC metrics
  seurat_obj <- NucleosomeSignal(seurat_obj)
  seurat_obj <- FRiP(
  object = seurat_obj,
  assay = 'ATAC',
  total.fragments = "fragments"
)
  
  #set the orig.ident to be able to identify samples later on
  seurat_obj$orig.ident <- seurat_list[[sample]]
  
  #Store in list, using sample ID as name
  seurat_list[[sample]] <- seurat_obj
  
}
seurat_list
#Save all of the objects for later (if needed)
saveRDS(seurat_list[["GEL_BU_BQ"]], "/scratch/ac05869/gelsemium_sc/seurat_data/GEL_BU_BQ_ATAC.RDS")
saveRDS(seurat_list[["GEL_BV_BR"]], "/scratch/ac05869/gelsemium_sc/seurat_data/GEL_BV_BR_ATAC.RDS")

```

##Make graphs of the UMI (nCount) and features (peaks)
### Replicate 1
```{r}
#The goal is to determine what cut-offs of UMI and feature counts should be used to remove low-quality cells. 

##if you used the code above, you can access each seurat object and its metadata from the list
scatter <- seurat_list[["GEL_BU_BQ"]]@meta.data %>% 
  ggplot(aes(x = nCount_ATAC, y = nFeature_ATAC)) +
  geom_point() +
  geom_vline(xintercept = 100, color = "red") + #keep changing the lines until you feel it fits
  geom_vline(xintercept = 30000, color = "red") +
  geom_hline(yintercept = 175, color = "blue") +
  geom_hline(yintercept = 25000, color = "blue") +
  scale_x_log10() + #without the log scale, we would see much more cells towards the bottom of our graph
  scale_y_log10() +
  theme_classic()

#Check the output
print(scatter) #then go back up and adjust if needed

#Also check the FRiP (fragment of reads in peak) scores
summary(seurat_list[["GEL_BU_BQ"]]@meta.data$FRiP) #0.08024 0.17632 0.21563 0.29120 0.40561 0.69458 

#save scatter plot
print(scatter)
ggsave("/scratch/ac05869/gelsemium_sc/seurat_qc/GEL_BU_BQ_ATAC_scatterplot.png", height = 10, width = 14, bg = "white")
```
### Replicate 2

```{r}
scatter <- seurat_list[["GEL_BV_BR"]]@meta.data %>% 
  ggplot(aes(x = nCount_ATAC, y = nFeature_ATAC)) +
  geom_point() +
  geom_vline(xintercept = 100, color = "red") + #keep changing the lines until you feel it fits
  geom_vline(xintercept = 30000, color = "red") +
  geom_hline(yintercept = 175, color = "blue") +
  geom_hline(yintercept = 25000, color = "blue") +
  scale_x_log10() + #without the log scale, we would see much more cells towards the bottom of our graph
  scale_y_log10() +
  theme_classic()

#Check the output
print(scatter) # then go back up and adjust if needed

#Also check the FRiP (fragment of reads in peak) scores
summary(seurat_list[["GEL_BV_BR"]]@meta.data$FRiP) # 0.1166  0.1795  0.2064  0.3250  0.5132  0.7942 

#save scatter plot
print(scatter)
ggsave("/scratch/ac05869/gelsemium_sc/seurat_qc/GEL_BV_BR_ATAC_scatterplot.png", height = 10, width = 14, bg = "white")
```

##Filter low quality cells from each ATAC assay

Make sure the filtering parameters are those that you set in the scatter plots. In this case, both samples here used the same filtering criterion but this may need     to be individualized for each library.
```{r}
#call the samples again if needed
atac_samples <- c("GEL_BU_BQ", "GEL_BV_BR")

#create another empty list for our subsetted ATAC assays
seurat_list_sub <- list()

for (sample in atac_samples) {
  
  #create individual seurat objects
  obj <- seurat_list[[sample]]
  
  #set default assay
  DefaultAssay(obj) <- "ATAC"
  
  #subset cells based on criteria determined above. In this case, all four samples here used the same filtering criterion but this may need     to be individualized for each library.
  seurat_obj_sub <- subset(
  x = obj,
  subset = nCount_ATAC > 100 &
    nCount_ATAC < 30000 &
    nFeature_ATAC > 175 &
    nFeature_ATAC < 25000 &
    FRiP > 0.25)
  
  
  #Store in list, using sample ID as name
  seurat_list_sub[[sample]] <- seurat_obj_sub
}
seurat_list_sub
#save objects as needed.
saveRDS(seurat_list_sub[["GEL_BU_BQ"]], "/scratch/ac05869/gelsemium_sc/seurat_data/GEL_BU_BQ_sub_ATAC.RDS")
saveRDS(seurat_list_sub[["GEL_BV_BR"]], "/scratch/ac05869/gelsemium_sc/seurat_data/GEL_BV_BR_sub_ATAC.RDS")
```

## Load in data if picking up from here

If needed, load in saved subsetted objs
```{r}
#call the samples again if needed
atac_samples <- c("GEL_BU_BQ", "GEL_BV_BR")

seurat_list_sub <- list()

for (sample in atac_samples) {
  
  data_dir = "/scratch/ac05869/gelsemium_sc/seurat_data/"
  
  seurat_obj_path = paste0(data_dir,sample,"_sub_ATAC.RDS")
  
  seurat_obj <- readRDS(seurat_obj_path)
  
  #Store in list, using sample ID as name
  seurat_list_sub[[sample]] <- seurat_obj
}
```


##Normalize, find top features, perform dimensional reduction to prepare assays for integration.
```{r}
#make an empty list to save the normalized data
seurat_norm <- list()

for (sample in atac_samples) {
  
  obj <- seurat_list_sub[[sample]]
  
  #set default assay to ATAC
  DefaultAssay(obj) <- "ATAC"
  
  #Normalize the data with TFIDF
  ##normalizes across cells and across peaks to account for sequence depth diff.s and to give higher values to rarer peaks. 
  obj <- RunTFIDF(obj)
  
  #Find top features
  ##Not like GEX where we can find highly variable features, because ATAC data is not that dynamic. We want to filter out features present in   less than n cells.
  ##min.cutoff specifies the minimum percentile, selecting the 95% most common features. This can be changed to whatever you like.
  obj <- FindTopFeatures(obj, min.cutoff = 'q5')
  
  #Perform linear dimensional reduction
  ##returns a reduced dimensional representation of the data, analogous to PCA
  obj <- RunSVD(obj)
  
    #Store in list, using sample ID as name
  seurat_norm[[sample]] <- obj
}

#Generate individual objects to be able to find integration anchors
root_R1_ATAC <- seurat_norm[["GEL_BU_BQ"]]
root_R2_ATAC <- seurat_norm[["GEL_BV_BR"]]

#clean up env
rm(seurat_norm, seurat_list_sub, obj, seurat_obj)
```

##Integrate ATAC Assays
```{r}
#generate an object list
obj_list <- list(root_R1_ATAC, root_R2_ATAC)

#Find integration anchors
anchors <- FindIntegrationAnchors(
  object.list = obj_list,
  reduction = "rlsi",
  dims = 2:30
)

#Integrate (this step takes forever!)
integrated_ATAC <- IntegrateData(anchorset = anchors, dims = 2:30)

#check and make sure the integrated object looks correct
integrated_ATAC

#save this integrated object for later on
saveRDS(integrated_ATAC, "/scratch/ac05869/gelsemium_sc/seurat_data/GEL_root_ATAC_integrated_14Sep26.RDS")
```

#Step 3: Find cells with high quality GEX and ATAC data

##Get an intersection of cell barcodes from GEX and ATAC
```{r}
#Add columns to the ATAC integrated obj's metadata layer. Drop the suffixes for the atac cb to be able to match with the GEX cb.
integrated_ATAC@meta.data <- integrated_ATAC@meta.data %>% 
  mutate(cell.bc = row.names(.)) %>% 
  mutate(cell.bc.2 = str_remove(cell.bc, "-1"))

#Generate a list of shared barcodes between the ATAC and GEX data
shared.cells.meta <- integrated_ATAC@meta.data %>% 
  filter(cell.bc.2 %in% Cells(integrated_GEX))

#Check how many cells with high quality GEX and ATAC data there are. 
dim(shared.cells.meta)
#[1] 2895   11
```

##Filter the ATAC and GEX data to get the same number of nuclei for each assay to be able to combine 
```{r}
#subset cells in the ATAC integrated obj by cell barcodes (note that after this the cell names will look like CB-1_1)
combined.atac_sub <- subset(
  x = integrated_ATAC,
  cells = row.names(shared.cells.meta))

#check number of cells 
combined.atac_sub

# An object of class Seurat 
# 63036 features across 2895 samples within 2 assays 
# Active assay: integrated (31518 features, 31518 variable features)
#  1 layer present: data
#  1 other assay present: ATAC

#subset cells in the GEX integrated obj by cell barcodes (note that after this the cb will look like CB_1)
combined.gex_sub <- subset(
  x = integrated_GEX,
  cells = row.names(shared.cells.meta) %>%
    str_remove("-1")
)

#check number of cells 
combined.gex_sub

# An object of class Seurat 
# 24415 features across 2895 samples within 2 assays 
# Active assay: RNA (21415 features, 3000 variable features)
#  3 layers present: scale.data, data, counts
#  1 other assay present: integrated
#  2 dimensional reductions calculated: pca, umap

#rename cells in the ATAC assay to be able to merge! 
combined.atac_sub_renamed <- RenameCells(combined.atac_sub, new.names = str_remove(string = Cells(combined.atac_sub), pattern="-1"))

#check and make sure the string has been properly removed from atac bc
head(Cells(combined.atac_sub_renamed))
tail(Cells(combined.atac_sub_renamed))
```

##Add GEX and ATAC Objects into one Seurat object
```{r}
#Set the filtered GEX data to a new name that will be the multiome obj
root_multiome <- combined.gex_sub

#set default assay of this multiome obj to GEX
DefaultAssay(root_multiome) <- "RNA"

#set the default assay of the ATAC data to ATAC
DefaultAssay(combined.atac_sub_renamed) <- "ATAC"

#Set the filtered ATAC data as a new assay within the multiome object under the layer "ATAC"
root_multiome[["ATAC"]] <- combined.atac_sub_renamed[["ATAC"]]  

#confirm obj has both modalities present and the expected number of cells
root_multiome

# An object of class Seurat 
# 55933 features across 2895 samples within 3 assays 
# Active assay: RNA (21415 features, 3000 variable features)
#  3 layers present: scale.data, data, counts
#  2 other assays present: integrated, ATAC
#  2 dimensional reductions calculated: pca, umap
 
#save data as needed
saveRDS(root_multiome, "/scratch/ac05869/gelsemium_sc/seurat_data/GEL_root_multiome_GEX_ATAC_14Sep26.RDS")
```

#Step 4: First pass data visualization and cell type annotation
##load in multiome obj if needed
```{r}
#load in multiome obj
root_multiome <- readRDS("/scratch/ac05869/gelsemium_sc/seurat_data/GEL_root_multiome_GEX_ATAC_14Sep26.RDS")
```

##Run UMAP on GEX
```{r}
#run SCTranform which normalizes the data, finds highly variable features, then scales the data
DefaultAssay(root_multiome) <- "RNA"

root_multiome <- SCTransform(root_multiome, verbose = F)

root_multiome <- RunPCA(root_multiome)

root_multiome <- RunUMAP(root_multiome, dims = 1:50, reduction.name = 'umap.rna', reduction.key = 'rnaUMAP_', min.dist = 0.01, repulsion.strength = 0.5, n.neighbors = 50, spread = 0.5)
```

##Run UMAP on ATAC
```{r}
#Perform similar normalization, feature identification, and dimensionality reduction as before. 
DefaultAssay(root_multiome) <- "ATAC"

root_multiome <- RunTFIDF(root_multiome)

root_multiome <- FindTopFeatures(root_multiome, min.cutoff = 'q10')

root_multiome <- RunSVD(root_multiome)

root_multiome <- RunUMAP(root_multiome, reduction = 'lsi', dims = 2:30, reduction.name = "umap.atac", reduction.key = "atacUMAP_", min.dist = 0.001, n.neighbors = 25, spread = 0.5, repulsion.strength = 3)
```

##Run multimodal WNN analysis
```{r}
#This analysis integrates both modalities and learns "weights" for each cell to determine how important each modality is to each cell. It can then find neighbors of cells with similar weights across modalities and can help improve the resolution of cell states.  

root_multiome <- FindMultiModalNeighbors(root_multiome, reduction.list = list("pca", "lsi"), dims.list = list(1:50, 2:50))

root_multiome <- RunUMAP(root_multiome, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnnUMAP_", min.dist = 0.01, repulsion.strength = 0.5, n.neighbors = 50, spread = 0.5)

root_multiome <- FindClusters(root_multiome, graph.name = "wsnn", algorithm = 3, verbose = F)
```

##Generate UMAPs by cluster
```{r}
p1 <- DimPlot(root_multiome, reduction = "umap.rna", label = T, repel = T) + ggtitle("RNA")

p2 <- DimPlot(root_multiome, reduction = "umap.atac", label = T, repel = T) + ggtitle("ATAC")

p3 <- DimPlot(root_multiome, reduction = "wnn.umap", label = T, repel = T) + ggtitle("WNN")

p1 + p2 + p3 & NoLegend() & theme(plot.title = element_text(hjust= 0.5))

ggsave("/scratch/ac05869/gelsemium_sc/seurat_qc/GEL_root_multiome_cluster_UMAPs_14Sep26.svg", width = 18, height = 6, bg = "white")
```

##Generate UMAPs by library
```{r}
#Note that the GEX library info is the default orig identification for the object
Idents(root_multiome) <- "orig.ident"

p1 <- DimPlot(root_multiome, reduction = "umap.rna", label = F, repel = T) + ggtitle("RNA")

p2 <- DimPlot(root_multiome, reduction = "umap.atac", label = F, repel = T) + ggtitle("ATAC")

p3 <- DimPlot(root_multiome, reduction = "wnn.umap", label = F, repel = T) + ggtitle("WNN")

p1 + p2 + p3 & theme(plot.title = element_text(hjust= 0.5))

ggsave("/scratch/ac05869/gelsemium_sc/seurat_qc/GEL_root_multiome_library_UMAPs_14Sep26.svg", width = 18, height = 6, bg = "white")
```
## Save data
```{r}
DefaultAssay(root_multiome) <- "RNA"
Idents(root_multiome) <- "seurat_clusters"
#root_multiome@active.ident

#joinlayers
root_multiome <- JoinLayers(root_multiome)

saveRDS(root_multiome, "/scratch/ac05869/gelsemium_sc/seurat_data/GEL_root_multiome_GEX_ATAC_14Sep26.RDS")
saveRDS(root_multiome, "/work/crblab/ac05869/sc_analyses/seurat_object/GEL_root_multiome.rds")
root_multiome <- readRDS(file = "/scratch/ac05869/gelsemium_sc/seurat_data/GEL_root_multiome_GEX_ATAC_14Sep26.RDS")
```
##Annotation data - both nuclear mitochondrial and chloroplast
```{r}
repr_gene_models <- read_delim("/work/crblab/ac05869/genomes/gese_mpi_v2/gese_mpi_v2.working_models.repr.gff3",
                               delim = "\t", col_names = F)
anno <- read_delim("/work/crblab/ac05869/genomes/gese_mpi_v2/gese_mpi_v2.functional_annotation.txt", col_names = c("gene_ID", "gene_anno"))

repr_models <- repr_gene_models %>% 
  dplyr::filter(X3 == "mRNA") %>% 
  dplyr::select(X1, X4, X5, X7, X9) %>% 
  dplyr::rename(
    Chr = X1,
    start = X4,
    end = X5,
    strand = X7
  ) %>%
  separate(X9, c("ID", "gene_ID", "Parent"), sep = ";") %>%
  mutate(across(everything(), ~ str_trim(str_replace_all(.x, '"', ''))))%>%
  mutate(gene_ID = gsub("^Name=", "", gene_ID)) %>%
  mutate(Parent = gsub("^Parent=", "", Parent))%>%
  dplyr::select(c(Chr, start, end, strand, gene_ID, Parent)) %>%
  left_join(anno, by = "gene_ID")

head(repr_models)
dim(repr_models) #28509

repr_models$Parent <- gsub("_", "-", repr_models$Parent)
head(repr_models)
```

## Find Markers
```{r}
mpi.root.markers <- FindAllMarkers(root_multiome, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25) #min.pct is the minimum fraction that cells cluster needs to be expressing that gene for it to count. logfc.threshhold is the log-fold change in expression between clusters expressing that gene
head(mpi.root.markers)
#mpi.root.markers$gene <- gsub(pattern = "-", replacement = "_", x = mpi.root.markers$gene)
head(mpi.root.markers)

mpi.root.markers_annotated <- mpi.root.markers %>%
  left_join(repr_models, by = c("gene" = "Parent")) %>% 
  group_by(cluster)

head(mpi.root.markers_annotated)
write_delim(mpi.root.markers_annotated, "/scratch/ac05869/gelsemium_sc/seurat_data/gese_mpi_v2.root_markers_annotated_minpct0.25_minlfc0.25.tsv", delim = "\t")
#mpi.root.markers_annotated <- read_delim("/scratch/ac05869/gelsemium_sc/seurat_data/gese_mpi_v2.root_markers_annotated_minpct0.25_minlfc0.25.tsv", delim = "\t")
```

## Find Average Expression
```{r}
# replace spaces with underscores '_' so ggplot2 doesn't fail
orig.levels <- levels(root_multiome)
Idents(root_multiome) <- gsub(pattern = " ", replacement = "_", x = Idents(root_multiome))
orig.levels <- gsub(pattern = " ", replacement = "_", x = orig.levels)
levels(root_multiome) <- orig.levels
cluster.averages <- AverageExpression(root_multiome, return.seurat = TRUE)
head(cluster.averages)

# How can I extract expression matrix
cluster_avg.data <- as.data.frame(GetAssayData(cluster.averages, layer = "data")[, ])
cluster_avg.data$gene <- rownames(cluster_avg.data)
cluster_avg.data$gene <- gsub(pattern = "-", replacement = "_", x = cluster_avg.data$gene)
head(cluster_avg.data)
write.table(cluster_avg.data, file = "/scratch/ac05869/gelsemium_sc/seurat_data/gese_mpi_v2.root_multiome_cluster_av.norm.counts.tsv", sep = "\t", quote = FALSE)

```


# Plot Pathway genes
```{r}
path <- read_delim("/scratch/ac05869/gelsemium_sc/gelsemium_sempervirens_discovered_genes.tsv", delim = "\t") %>%
  mutate(order = row_number())

head(repr_models)
path$gene_ID <- gsub(pattern = "_", replacement = "-", x = path$gene_ID)
path$gene_ID <- gsub("\\.\\d+$", "", path$gene_ID)

path <- path %>% 
  left_join(repr_models %>% select(-gene_ID), by = c("gene_ID" = "Parent"))
head(path)
rev_path <- path[order(nrow(path):1),]
head(rev_path)
```

```{r}
dotplot_data <- DotPlot(root_multiome, features = unique(rev_path$gene_ID))$data
head(dotplot_data)
p1_nice_df<- dotplot_data %>% 
  inner_join(rev_path %>% 
               distinct(gene_ID, .keep_all = T), by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(Chr, end = -5),
    LocusID = reorder(LocusID, -order),
    ylab = paste0(features.plot),  #for plot y axis
    ylab2 = paste0(gene),  # make this new variable
    ylab2 = reorder(ylab2, -order),
    ylab2 = as.factor(ylab2))
```


```{r}
p1 <- p1_nice_df %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "") +
  theme_classic() +
  theme(text = element_text(size = 18, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 1, face = "bold"),
        legend.key.height = unit(0.7, "lines"),
        legend.position = "right",
        legend.box = "vertical")
p1

ggsave(file = "/scratch/ac05869/gelsemium_sc/R_output/GEL_root_pathway_dotplot.png", height = 8, width = 7)
ggsave(file = "/scratch/ac05869/gelsemium_sc/R_output/GEL_root_pathway_dotplot.svg", height = 8, width = 7)
```
```{r}
DimPlot(root_multiome, reduction = "umap.rna", label = T, label.size = 5, repel = T, group.by = "seurat_clusters") +
  labs(title = "MPI Gelsemium root")+
  theme_classic() +
  theme(legend.position = "right")
ggsave(file = "/scratch/ac05869/gelsemium_sc/R_output/GEL_root_rna_umap.png", height = 6, width = 6)
ggsave(file = "/scratch/ac05869/gelsemium_sc/R_output/GEL_root_rna_umap.svg", height = 6, width = 6)
```

