---
title: "KRT leaf ATAC preprocessing haplotype 1"
author: "Anne Pavia"
date: "2025-June-3"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('/scratch/ac05869/10X_Multiome/KRT_leaf/ATAC_analysis/wd')
knitr::opts_knit$set(root.dir = '/scratch/ac05869/10X_Multiome/KRT_leaf/ATAC_analysis/wd')
```

# Packages 
```{r}
#install.packages("hdf5r")
library(hdf5r)
library(tidyverse)
library(Seurat) 
library(Signac)
library(GenomicRanges)
library(readxl)
library(RColorBrewer)
library(viridis)
library(rcartocolor)
library(patchwork)
library(showtext)
showtext_auto(enable = F)
```
#WD
paste into terminal:
setwd('/scratch/ac05869/10X_Multiome/KRT_leaf/ATAC_analysis/wd')
cd /scratch/ac05869/10X_Multiome/KRT_leaf/ATAC_analysis/wd

# KRT_AC_AE: Rep 1
## Data 
```{r}
KRT1_counts <- Read10X_h5(filename = "../AC_cellranger_out/filtered_feature_bc_matrix.h5")

KRT1 <- CreateSeuratObject(
  counts = KRT1_counts$`Gene Expression`,
  assay = "RNA",
  project = "Rep 1"
)

KRT1[["ATAC"]] <- CreateChromatinAssay(
  counts = KRT1_counts$Peaks,
  sep = c(":", "-"),
  fragments = "../AC_cellranger_out/atac_fragments.tsv.gz"
)

KRT1

```
138322 features
5271 samples
2 assays: counts('Gene Expression) and ATAC(chromatin assay)

## Filter nuclei
```{r}
DefaultAssay(KRT1) <- "ATAC"
KRT1 <- NucleosomeSignal(object = KRT1)
```

```{r}
VlnPlot(
  object = KRT1,
  features = c('nCount_RNA', 'nCount_ATAC',
               'nucleosome_signal'),
  pt.size = 0,
  ncol = 3
) 


```
```{r}
A<- KRT1$nCount_RNA %>% 
  as.data.frame() %>% 
  ggplot(aes(x = `.`)) + 
  geom_histogram(bins = 100, color = "white") +
  geom_vline(xintercept = 5000, color = "tomato1") +
  geom_vline(xintercept = 300, color = "tomato1") +
  ggtitle("Filtering nCount_RNA") +
  theme_classic()

B<- KRT1$nCount_ATAC %>% 
  as.data.frame() %>% 
  ggplot(aes(x = `.`)) + 
  geom_histogram(bins = 100, color = "white") +
  geom_vline(xintercept = 40000, color = "tomato1") +
  geom_vline(xintercept = 3000, color = "tomato") +
  ggtitle("Filtering nCount_ATAC") +
  theme_classic()

filtering <-
  wrap_plots(A + labs(tag = "(a)"),
           B + labs(tag = "(b)"),
           ncol = 1, nrow = 2,
           widths = c(1), 
           heights = c(0.5, 0.5))
filtering
ggsave(file = "../R_output/KRT_AC_AE_filter.svg", height = 4, width = 5, bg = "white")
ggsave(file = "../R_output/KRT_AC_AE_filter.png", height = 4, width = 5, bg = "white")
```


```{r}
KRT1 <- subset(
  x = KRT1,
  subset =  nCount_ATAC < 40000 &
    nCount_RNA < 5000 &
    nCount_ATAC > 3000 & 
    nCount_RNA > 300 &
    nucleosome_signal < 2
)

KRT1

```
138322 features
4077 samples

## GEX processing
```{r}
DefaultAssay(KRT1) <- "RNA"
KRT1 <- SCTransform(KRT1) 
KRT1 <- RunPCA(KRT1) 
KRT1 <- RunUMAP(KRT1, dims = 1:50, reduction.name = "umap.rna",
                reduction.kep = "rnaUMAP_")
```

## ATAC processing
```{r}
DefaultAssay(KRT1) <- "ATAC"
KRT1 <- FindTopFeatures(KRT1, min.cutoff = "q0")
KRT1 <- RunTFIDF(KRT1)
KRT1 <- RunSVD(KRT1)
KRT1 <- RunUMAP(KRT1, reduction = 'lsi', dims = 2:50, reduction.name = "umap.atac", reduction.key = "atacUMAP_")
```

## Joint UMAP
```{r}
KRT1 <- FindMultiModalNeighbors(
  object = KRT1,
  reduction.list = list("pca", "lsi"), 
  dims.list = list(1:50, 2:50),
  modality.weight.name = "RNA.weight",
  verbose = TRUE
)

# build a joint UMAP visualization
KRT1 <- RunUMAP(
  object = KRT1,
  nn.name = "weighted.nn",
  reduction.name = "wnn.umap",
  reduction.key = "wnnUMAP_"
)

```
```{r}
KRT1 <- FindClusters(KRT1, graph.name = "wsnn", algorithm = 3)
```


```{r}
KRT1_rna <- DimPlot(KRT1, label = TRUE, repel = TRUE, reduction = "umap.rna") +
  NoLegend() +
  theme_void() +
  ggtitle("RNA")

KRT1_atac <- DimPlot(KRT1, label = TRUE, repel = TRUE, reduction = "umap.atac") +
  NoLegend() +
  theme_void()+
  ggtitle("ATAC")

KRT1_joint <- DimPlot(KRT1, label = TRUE, repel = TRUE, reduction = "wnn.umap") +
  NoLegend() +
  theme_void() +
  ggtitle("Joint")

wrap_plots(
  KRT1_rna, KRT1_atac, KRT1_joint
) &
  NoLegend() &
  theme(plot.title = element_text(hjust = 0.5)) &
  theme(plot.margin = margin(t = 0, r = 3, b = 0, l = 0.5, unit = "lines")) 

ggsave("../R_output/KRT1_UMAPs.svg", height = 3, width = 8, bg = "white")
ggsave("../R_output/KRT1_UMAPs.png", height = 3, width = 8, bg = "white")
```
# Write out filtered object
```{r}
KRT1$run <- "KRT1"
DefaultAssay(KRT1) <- "RNA"

KRT1

saveRDS(KRT1, "../R_output/KRT1_filtered.Rds")
```

KRT1: 5271 nuclei

```{r}
KRT1$nCount_RNA %>% summary()
KRT1$nFeature_RNA %>% summary()

KRT1$nCount_ATAC %>% summary()
KRT1$nFeature_ATAC %>% summary()
```
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    103    1031    1451    1580    1940   13141 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
     99     856    1162    1212    1488    7062 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    119    3876    8839   11227   15082  129690 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
     61    1871    4010    4850    6598   35421 

# KRT_AD_AF: Rep 2 (KRT2)
## Data 
```{r}
KRT2_counts <- Read10X_h5(filename = "../AD_cellranger_out/filtered_feature_bc_matrix.h5")

KRT2 <- CreateSeuratObject(
  counts = KRT2_counts$`Gene Expression`,
  assay = "RNA",
  project = "Rep 2"
)

KRT2[["ATAC"]] <- CreateChromatinAssay(
  counts = KRT2_counts$Peaks,
  sep = c(":", "-"),
  fragments = "../AD_cellranger_out/atac_fragments.tsv.gz"
)

KRT2
```
 101664 features across 5322 samples within 2 assays 

## Filter nuclei
```{r}
DefaultAssay(KRT2) <- "ATAC"
KRT2 <- NucleosomeSignal(object = KRT2)
```

```{r}
VlnPlot(
  object = KRT2,
  features = c('nCount_RNA', 'nCount_ATAC',
               'nucleosome_signal'),
  pt.size = 0,
  ncol = 3
) 


```
```{r}
C<- KRT2$nCount_RNA %>% 
  as.data.frame() %>% 
  ggplot(aes(x = `.`)) + 
  geom_histogram(bins = 100, color = "white") +
  geom_vline(xintercept = 5000, color = "tomato1") +
  geom_vline(xintercept = 500, color = "tomato1") +
  ggtitle("Filtering nCount_RNA") +
  theme_classic()

D<- KRT2$nCount_ATAC %>% 
  as.data.frame() %>% 
  ggplot(aes(x = `.`)) + 
  geom_histogram(bins = 100, color = "white") +
  geom_vline(xintercept = 40000, color = "tomato1") +
  geom_vline(xintercept = 2000, color = "tomato") +
  ggtitle("Filtering nCount_ATAC") +
  theme_classic()

filtering <-
  wrap_plots(C + labs(tag = "(a)"),
           D + labs(tag = "(b)"),
           ncol = 1, nrow = 2,
           widths = c(1), 
           heights = c(0.5, 0.5))
filtering
ggsave(file = "../R_output/KRT_AD_AF_filter.svg", height = 4, width = 5, bg = "white")
ggsave(file = "../R_output/KRT_AD_AF_filter.png", height = 4, width = 5, bg = "white")
```


```{r}
KRT2 <- subset(
  x = KRT2,
  subset =  nCount_ATAC < 40000 &
    nCount_RNA < 5000 &
    nCount_ATAC > 2000 & 
    nCount_RNA > 500 &
    nucleosome_signal < 2
)

KRT2
```
101664 features across 4262 samples

## GEX processing
```{r}
DefaultAssay(KRT2) <- "RNA"
KRT2 <- SCTransform(KRT2) 
KRT2 <- RunPCA(KRT2) 
KRT2 <- RunUMAP(KRT2, dims = 1:50, reduction.name = "umap.rna",
                reduction.kep = "rnaUMAP_")
```

## ATAC processing
```{r}
DefaultAssay(KRT2) <- "ATAC"
KRT2 <- FindTopFeatures(KRT2, min.cutoff = "q0")
KRT2 <- RunTFIDF(KRT2)
KRT2 <- RunSVD(KRT2)
KRT2 <- RunUMAP(KRT2, reduction = 'lsi', dims = 2:50, reduction.name = "umap.atac", reduction.key = "atacUMAP_")
```

## Joint UMAP
```{r}
KRT2 <- FindMultiModalNeighbors(
  object = KRT2,
  reduction.list = list("pca", "lsi"), 
  dims.list = list(1:50, 2:50),
  modality.weight.name = "RNA.weight",
  verbose = TRUE
)

# build a joint UMAP visualization
KRT2 <- RunUMAP(
  object = KRT2,
  nn.name = "weighted.nn",
  reduction.name = "wnn.umap",
  reduction.key = "wnnUMAP_"
)

```

```{r}
KRT2 <- FindClusters(KRT2, graph.name = "wsnn", algorithm = 3)
```


```{r}
KRT2_rna <- DimPlot(KRT2, label = TRUE, repel = TRUE, reduction = "umap.rna") +
  NoLegend() +
  theme_void() +
  ggtitle("RNA")

KRT2_atac <- DimPlot(KRT2, label = TRUE, repel = TRUE, reduction = "umap.atac") +
  NoLegend() +
  theme_void()+
  ggtitle("ATAC")

KRT2_joint <- DimPlot(KRT2, label = TRUE, repel = TRUE, reduction = "wnn.umap") +
  NoLegend() +
  theme_void() +
  ggtitle("Joint")

wrap_plots(
  KRT2_rna, KRT2_atac, KRT2_joint
) &
  NoLegend() &
  theme(plot.title = element_text(hjust = 0.5)) &
  theme(plot.margin = margin(t = 0, r = 3, b = 0, l = 0.5, unit = "lines")) 

ggsave("../R_output/KRT2_UMAPs.svg", height = 3, width = 8, bg = "white")
ggsave("../R_output/KRT2_UMAPs.png", height = 3, width = 8, bg = "white")
```
# Write out filtered object
```{r}
KRT2$run <- "KRT2"
DefaultAssay(KRT2) <- "RNA"

KRT2

saveRDS(KRT2, "../R_output/KRT2_filtered.Rds")
```

KRT2: 4262 nuclei

```{r}
KRT2$nCount_RNA %>% summary()
KRT2$nFeature_RNA %>% summary()

KRT2$nCount_ATAC %>% summary()
KRT2$nFeature_ATAC %>% summary()
```
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    506    1269    1719    1818    2212    4979 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    420    1008    1318    1361    1638    3239 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   2017    5339    9258   11329   15200   39878 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    833    2490    4169    4854    6508   15036 
    
