---
title: "ATAC_integration"
author: "Anne Pavia"
date: "4-June-2025"
output:  
  html_notebook:
    number_sections: yes
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/scratch/ac05869/10X_Multiome/KRT_leaf/ATAC_analysis/wd')
setwd('/scratch/ac05869/10X_Multiome/KRT_leaf/ATAC_analysis/wd')
```
# Packages 
```{r}
library(tidyverse)
library(Seurat) 
library(Signac)
library(GenomicRanges)
library(glmGamPoi)
library(readxl)

library(RColorBrewer)
library(viridis)
library(rcartocolor)

library(patchwork)
library(showtext)
```

# Data 
## Load a common set of peaks
Use merged peaks: 

On GARCR, do:
Combine the atac_peaks.bed file from replicates. 
Sort them. 
Remove duplicates (if any).
Then use bedtools merge to make a single bed file. 

```{r}
arc_peaks <- read_delim("../bedtools/atac_peaks.sorted.merged.bed",
                          delim = "\t", col_names = F, col_types = cols())

arc_peaks$X4 <- "name"
head(arc_peaks)
```
### Produce GRange from bed
```{r}
arc_peaks_granges <- lapply(split(arc_peaks, arc_peaks$X4), function(i){
  GRanges(seqnames = i$X1,
          ranges = IRanges(
            start = i$X2,
            end = i$X3
          ))
})

arc_peaks_granges
```
```{r}
gr <- arc_peaks_granges$name
#arc_peaks_nice <- keepStandardChromosomes(arc_peaks_granges$name, pruning.mode = "coarse")
#arc_peaks_nice
```

## Rep 1: KRT_AC_AE (KRT1)
```{r}
KRT1_metadata <- read.csv(
  "../AC_cellranger_out/per_barcode_metrics.csv",
  header = T,
  row.names = 1
)

KRT1_metadata
```

```{r}
KRT1_frag_path <- "../AC_cellranger_out/atac_fragments.tsv.gz"
KRT1_frag_counts <- CountFragments(fragments = KRT1_frag_path)
KRT1_cells <- KRT1_frag_counts$CB
KRT1_fragment_object <- CreateFragmentObject(path = KRT1_frag_path, cells = KRT1_frag_counts$CB)
```


```{r}
KRT1_ATAC_matrix <- FeatureMatrix(
  fragments = KRT1_fragment_object,
  features = gr,
  cells = KRT1_frag_counts$CB
)

KRT1_ATAC_assay <- CreateChromatinAssay(
  counts = KRT1_ATAC_matrix,
  min.features = 1000,
  fragments = KRT1_fragment_object
) 

KRT1_ATAC_object <- CreateSeuratObject(
  counts = KRT1_ATAC_assay, 
  assay = "peaks",
  meta.data = KRT1_metadata, 
  min.cells = 0,
  min.features = 0)
```
```{r}
showtext_auto(enable = T)
KRT1_metadata %>% 
  mutate(FRiP = atac_peak_region_fragments/atac_fragments * 100) %>% 
  mutate(FRiT = atac_TSS_fragments/atac_fragments * 100) %>% 
  ggplot(aes(x = atac_fragments, y = FRiP)) +
  geom_point() +
  scale_x_log10() +
  #scale_y_log10() +
  theme_classic()

KRT1_metadata %>% 
  mutate(FRiP = atac_peak_region_fragments/atac_fragments * 100) %>% 
  mutate(FRiT = atac_TSS_fragments/atac_fragments * 100) %>% 
  ggplot(aes(x = atac_fragments, y = FRiT)) +
  geom_point() +
  scale_x_log10() +
  #scale_y_log10() +
  theme_classic()
showtext_auto(enable = F)
```

## Rep2: KRT_AD_AF (KRT2)
```{r}
KRT2_frag_path <- "../AD_cellranger_out/atac_fragments.tsv.gz"
KRT2_frag_counts <- CountFragments(fragments = KRT2_frag_path)

KRT2_metadata <- read.csv(
  "../AD_cellranger_out/per_barcode_metrics.csv",
  header = T,
  row.names = 1
)

KRT2_cells <- KRT2_frag_counts$CB
KRT2_fragment_object <- CreateFragmentObject(path = KRT2_frag_path, cells = KRT2_cells)
```

```{r}
KRT2_ATAC_matrix <- FeatureMatrix(
  fragments = KRT2_fragment_object,
  features = gr,   
  cells = KRT2_cells
)

KRT2_ATAC_assay <- CreateChromatinAssay(
  counts = KRT2_ATAC_matrix,
  min.features = 1000,
  fragments = KRT2_fragment_object
) 

KRT2_ATAC_object <- CreateSeuratObject(
  counts = KRT2_ATAC_assay, 
  assay = "peaks",
  meta.data = KRT2_metadata,
  min.cells = 0,
  min.features = 0)
```

```{r}
showtext_auto(enable = T)
KRT2_metadata %>% 
  mutate(FRiP = atac_peak_region_fragments/atac_fragments * 100) %>% 
  mutate(FRiT = atac_TSS_fragments/atac_fragments * 100) %>% 
  ggplot(aes(x = atac_fragments, y = FRiP)) +
  geom_point() +
  scale_x_log10() +
  #scale_y_log10() +
  theme_classic()

KRT2_metadata %>% 
  mutate(FRiP = atac_peak_region_fragments/atac_fragments * 100) %>% 
  mutate(FRiT = atac_TSS_fragments/atac_fragments * 100) %>% 
  ggplot(aes(x = atac_fragments, y = FRiT)) +
  geom_point() +
  scale_x_log10() +
  #scale_y_log10() +
  theme_classic()
showtext_auto(enable = F)

```


# Filter nuclei 
## Nucleosome signal 
```{r}
KRT1_ATAC_object <- NucleosomeSignal(KRT1_ATAC_object)
KRT2_ATAC_object <- NucleosomeSignal(KRT2_ATAC_object)
```

## Reads in peaks 
```{r}
KRT1_ATAC_object$read_in_peaks <- KRT1_ATAC_object$atac_peak_region_fragments / KRT1_ATAC_object$atac_fragments * 100

KRT2_ATAC_object$read_in_peaks <- KRT2_ATAC_object$atac_peak_region_fragments / KRT2_ATAC_object$atac_fragments * 100
```

## TSS fractions
```{r}
KRT1_ATAC_object$TSS <- KRT1_ATAC_object$atac_TSS_fragments / KRT1_ATAC_object$atac_fragments * 100

KRT2_ATAC_object$TSS <- KRT2_ATAC_object$atac_TSS_fragments / KRT2_ATAC_object$atac_fragments * 100
```

 
## Filter per sample 
### KRT1
```{r}
KRT1_ATAC_object@meta.data %>% head()
```

```{r}
KRT1_ATAC_object@meta.data %>% 
  ggplot(aes(y = nFeature_peaks, x = atac_fragments)) +
  geom_point() +
  geom_hline(yintercept = 1000, color = "red2") +
  geom_vline(xintercept = 2000, color = "red2") +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()
```

```{r}
KRT2_ATAC_object@meta.data %>% 
  ggplot(aes(y = nFeature_peaks, x = atac_fragments)) +
  geom_point() +
  geom_hline(yintercept = 1000, color = "red2") +
  geom_vline(xintercept = 2000, color = "red2") +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()
```



```{r}
KRT1_ATAC_object <- subset(
  x = KRT1_ATAC_object,
  subset = nCount_peaks > 1000 & 
    read_in_peaks > 25 &
    atac_fragments > 2000
    
)

KRT1_ATAC_object
```
90651 features across 5517
### KRT2
```{r}
KRT2_ATAC_object <- subset(
  x = KRT2_ATAC_object,
  subset = nCount_peaks > 1000 & 
    read_in_peaks > 25 &
    atac_fragments > 2000
    
)

KRT2_ATAC_object
```
90651 features across 4338 samples

# Pre-integration processing
```{r}
KRT1_ATAC_object <- FindTopFeatures(KRT1_ATAC_object, min.cutoff = 10)
KRT2_ATAC_object <- FindTopFeatures(KRT2_ATAC_object, min.cutoff = 10)
```

```{r}
KRT1_ATAC_object <- RunTFIDF(KRT1_ATAC_object, verbose = F)
KRT2_ATAC_object <- RunTFIDF(KRT2_ATAC_object, verbose = F)
```

```{r}
KRT1_ATAC_object <- RunSVD(KRT1_ATAC_object, verbose = F, approx = F)
KRT2_ATAC_object <- RunSVD(KRT2_ATAC_object, verbose = F, approx = F)
```

# Merge first 
```{r}
KRT1_ATAC_object$run <- "KRT1"
KRT2_ATAC_object$run <- "KRT2"

KRT_ATAC_combined <- merge(KRT1_ATAC_object,
                           list(KRT2_ATAC_object), 
                           #add.cell.ids = c("KRT1", "KRT2")
                           )

KRT_ATAC_combined <- FindTopFeatures(KRT_ATAC_combined, min.cutoff = 10)
KRT_ATAC_combined <- RunTFIDF(KRT_ATAC_combined, verbose = F)
KRT_ATAC_combined <- RunSVD(KRT_ATAC_combined, verbose = F)

KRT_ATAC_combined <- RunUMAP(KRT_ATAC_combined, reduction = "lsi", dims = 1:30)
```

 

# Integration 
```{r}
ATAC_integration_anchors <- FindIntegrationAnchors(
  object.list = list(KRT1_ATAC_object, KRT2_ATAC_object),
  anchor.features = rownames(KRT1_ATAC_object1),
  reduction = "rlsi",
  dims = 2:30
)

KRT_ATAC <- IntegrateEmbeddings(
  anchorset = ATAC_integration_anchors, 
  reductions = KRT_ATAC_combined[["lsi"]],
  new.reduction.name = "integrated_lsi",
  dims.to.integrate = 1:30
)

KRT_ATAC <- RunUMAP(KRT_ATAC, reduction = "integrated_lsi", dims = 2:30)

Integrated <- DimPlot(KRT_ATAC, group.by = "run") +
  ggtitle("Integrated")

Integrated
```

```{r}
DimPlot(KRT_ATAC, group.by = "run", split.by = "run") 

ggsave("../R_output/ATAC_UMAP_by_run.svg", height = 3, width = 6, bg = "white")
ggsave("../R_output/ATAC_UMAP_by_run.png", height = 3, width = 6, bg = "white")
```

# Clustering 
```{r}
KRT_ATAC <- RunTFIDF(KRT_ATAC, verbose = F)
KRT_ATAC <- FindTopFeatures(KRT_ATAC, min.cutoff = 'q5')
KRT_ATAC <- RunSVD(KRT_ATAC, verbose = F)
KRT_ATAC <- RunUMAP(KRT_ATAC, reduction = "integrated_lsi", dims = 2:30, 
                    min.dist = 0.001, repulsion.strength = 1, n.neighbors = 30, spread = 1,
                    reduction.name = "umap.atac", reduction.key = "atacUMAP_")

DimPlot(KRT_ATAC, group.by = "run", reduction = "umap.atac") 
```

```{r}
KRT_ATAC <- FindNeighbors(KRT_ATAC, reduction = "lsi", dims = 1:30)
KRT_ATAC <- FindClusters(KRT_ATAC, verbose = F, algorithm = 2)
DimPlot(KRT_ATAC, label = T, reduction = "umap.atac") + NoLegend()
```
```{r}
DefaultAssay(KRT_ATAC) <- "peaks"
saveRDS(KRT_ATAC, "../R_output/KRT_ATAC.Rds")
KRT_ATAC <- readRDS("../R_output/KRT_ATAC.Rds")
```

# Integrate with RNA 
## Pull RNA object 
```{r}
KRT_RNA <- readRDS("../../seurat_out/Integrated_RNA_Data/KRT_RNA_multiome_only_5june25.rds") 
DefaultAssay(KRT_RNA) <- "RNA"
```

```{r}
KRT_RNA
KRT_ATAC
```
RNA object: 72704 features across 12082 samples
ATAC object: 90651 features across 9855 samples
## Subset shared barcodes 
```{r}
shared_cells <- intersect(
  colnames(KRT_RNA),
  str_remove(colnames(KRT_ATAC), "-1")
)

KRT_RNA@meta.data$barcode <- row.names(KRT_RNA@meta.data)
KRT_ATAC@meta.data$barcode <- row.names(KRT_ATAC@meta.data)

KRT_RNA$cell <- colnames(KRT_RNA)
KRT_ATAC$cell<- str_remove(colnames(KRT_ATAC), "-1")


KRT_RNA_subset <- subset(
  x = KRT_RNA,
  subset = cell %in% shared_cells 
)

KRT_ATAC_subset <- subset(
  x = KRT_ATAC,
  subset = cell %in% shared_cells 
)


#colnames(KRT_RNA_subset)
dim(KRT_ATAC_subset)
dim(KRT_RNA_subset)

# row.names(KRT_ATAC@meta.data) <- str_remove(colnames(KRT_ATAC), "-1")
KRT_ATAC_subset_renamed <- RenameCells(KRT_ATAC_subset, 
            new.names = str_remove(colnames(KRT_ATAC_subset), "-1"))
```

## Append objects 
```{r}
KRT_multiome <- KRT_RNA_subset
DefaultAssay(KRT_multiome) <- "RNA"
KRT_multiome[["peaks"]] <- KRT_ATAC_subset_renamed[["peaks"]]

KRT_multiome
```

## Joint Visualization 
```{r}
DefaultAssay(KRT_multiome) <- "RNA"

KRT_multiome <- SCTransform(KRT_multiome, verbose = FALSE) %>% 
  RunPCA() %>% 
  RunUMAP(dims = 1:30, reduction.name = 'umap.rna', reduction.key = 'rnaUMAP_',
          min.dist = 0.001, repulsion.strength = 1, n.neighbors = 30, spread = 1)

DefaultAssay(KRT_multiome) <- "peaks" 
KRT_multiome <- RunTFIDF(KRT_multiome)

KRT_multiome <- FindTopFeatures(KRT_multiome, min.cutoff = "q5")
KRT_multiome <- RunSVD(KRT_multiome)
KRT_multiome <- RunUMAP(KRT_multiome, reduction = "lsi", 
                        dims = 2:30, reduction.name = "umap.atac", reduction.key = "atacUMAP_",
                        min.dist = 0.001, repulsion.strength = 1, n.neighbors = 30, spread = 1)

KRT_multiome <- FindMultiModalNeighbors(
  KRT_multiome, reduction.list = list("pca", "lsi"), 
  dims.list = list(1:30, 2:30), verbose = F
)

KRT_multiome <- RunUMAP(KRT_multiome, 
                        nn.name = "weighted.nn", reduction.name = "wnn.umap",
                        reduction.key = "wnnUMAP_",
                        min.dist = 0.001, repulsion.strength = 1, n.neighbors = 30, spread = 1,
                        verbose = F)

```

```{r}
saveRDS(KRT_multiome, "../R_output/KRT_multiome_pre-cluster_metadata.Rds")
```
##### Left off here until annotation of clusters is completed.
```{r}
KRT_multiome@meta.data <- KRT_multiome@meta.data %>% 
  mutate(cell_type = case_when(
    seurat_clusters == 0 | 
      seurat_clusters == 1 | 
      seurat_clusters == 3 | 
      seurat_clusters ==7  ~ "Mesophyll",
    seurat_clusters == 2 | 
      seurat_clusters == 5 ~ "Epidermis",
    seurat_clusters == 12 ~ "Guard cells",
    seurat_clusters == 11 |
      seurat_clusters == 6 |
      seurat_clusters == 10 |
      seurat_clusters == 14 |
      seurat_clusters == 15  ~ "Vasculature",
    seurat_clusters == 13 ~ "IPAP",
      seurat_clusters == 8 ~ "Idioblast",
    T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type,
                      levels = c("Mesophyll", 
                                            "Epidermis", "Guard cells",
                                             "Vasculature", 
                                             "IPAP", "Idioblast", "Unassigned"
                                            )))
KRT_multiome@meta.data %>% 
  group_by(seurat_clusters) %>% 
  count()

KRT_multiome@meta.data %>% 
  group_by(cell_type) %>% 
  count()
```

```{r}

Multi_RNA_umap <- DimPlot(KRT_multiome, reduction = "umap.rna", 
                          group.by = "cell_type") +
  ggtitle("RNA")

Multi_ATAC_umap <- DimPlot(KRT_multiome, reduction = "umap.atac", 
                          group.by = "cell_type") +
  ggtitle("ATAC")

Multi_joint_umap <- DimPlot(KRT_multiome, reduction = "wnn.umap", 
                          group.by = "cell_type") +
  ggtitle("Joint")

wrap_plots(
 Multi_RNA_umap,
 Multi_ATAC_umap,
 Multi_joint_umap,
 nrow = 1, guides = "collect"
) &
  scale_color_manual(values = c(brewer.pal(6, "Accent")[1:3],
                                brewer.pal(6, "Set2")[6],
                                brewer.pal(6, "Accent")[5:6],
                                "grey80")) &
  theme_void() &
  theme(
    legend.position = "bottom",
    plot.margin = margin(r = 1, l = 1, unit = "lines"),
    text = element_text(size = 14)
  ) 
  
ggsave("../R_output/KRT_multiome_UMAP_bulk_peaks.svg", height = 3, width = 7, bg = "white")
ggsave("../R_output/KRT_multiome_UMAP_bulk_peaks.png", height = 3, width = 7, bg = "white")
```
```{r}
saveRDS(KRT_multiome, "../R_output/KRT_multiome.Rds")
```