---
title: "CoExp of MPI and UGA kratom leaf libraries (MIT_AH_AI and KRT_AA_AB)"
author: "Anne Pavia"
date: "2025-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(igraph)
library(ggraph)
library(Seurat)
library(readxl)
library(patchwork)
library(RColorBrewer)
library(viridis)
library(showtext)
#showtext_auto()
set.seed(666)
options(bitmapType='cairo')
```


# Create Matrices
## Find Markers 
```{r}
leaf.combined <- readRDS("/scratch/ac05869/MIT_KRT_sc_analysis/seurat_data/KRT_MIT_cleaned_integrated_seurat_21feb25.rds")
DefaultAssay(leaf.combined) <- "RNA"
cluster.averages <- AverageExpression(leaf.combined, return.seurat = TRUE)
cluster.averages
cluster_avg.data <- as.matrix(GetAssayData(cluster.averages, layer = "data")[, ]) 
Exp_table <- cluster_avg.data
colnames(Exp_table) <- as.factor(c(0:22))

head(Exp_table)
dim(Exp_table)
```
#GOI
```{r}
Baits <- read_delim("/scratch/ac05869/MIT_KRT_sc_analysis/leaf_baits_orthologs_12mar25.txt", delim = "\t", col_names = F, col_types = cols())
Baits$X2 <- gsub(pattern = "_", replacement = "-", x = Baits$X2)
Baits$X2 <- gsub("\\.\\d+$", "", Baits$X2)
head(Baits)
```


```{r}
Exp_table_long <- Exp_table %>%
  as.data.frame() %>%
  rownames_to_column("gene_ID") %>% 
  pivot_longer(cols = !gene_ID, names_to = "cluster", values_to = "avg_exp") 
head(Exp_table_long)
```

# Gene Coexpression Analysis

```{r}
Exp_table_long_averaged <- Exp_table_long

head(Exp_table_long_averaged)
```
We start from the long (tidy) table we made earlier. I also pulled the metadata as well to guide the averaging process. by = c("library"="Run) inside full_join() deals with the fact that the library ID is called library in the long table, but Run in the metadata. Then we filter for dissection_method == "Hand. group_by() followed by summarise(mean = ...) takes each gene, tissue, and cluster, and computes the mean. The elegance of a tidyverse based workflow is that you do not have to do loops! You let group_by() do the heavy lifting. This could take a moment. This step is doing a lot of mean calculations.

## Z score
It standardize the expression pattern of each gene to mean = 0, sd = 1. It is not absolutely necessary, but I have found including this step to produce results that better capture the underlying biology.
```{r}
Exp_table_long_averaged_z <- Exp_table_long_averaged %>% 
  group_by(gene_ID) %>% 
  mutate(z.score = (avg_exp - mean(avg_exp))/sd(avg_exp)) %>% 
  ungroup()

head(Exp_table_long_averaged_z)
```


## Gene selection - Omitted



## Gene-wise correlation

Now we can correlate each gene to every other gene. The essence of this workflow is simple, so we will use a simple correlation. If you want, you can use fancier methods such as GENIE3

We will use the cor() function in R. But the cor() only take vector or matrix as input, so we need to go from long to wide again.

```{r}
#need wide z-score
z_score_wide <- Exp_table_long_averaged_z %>% 
  select(gene_ID, cluster, z.score) %>% 
  pivot_wider(names_from = cluster, values_from = z.score) %>% 
  as.data.frame()

row.names(z_score_wide) <- z_score_wide$gene_ID
head(z_score_wide)

number_of_clusters <- ncol(z_score_wide) - 1
number_of_clusters

```

## Alternate Method to Generate Edge Table for Co-expression networks with > 30,000 genes


```{r}
cor_matrix <- cor(t(z_score_wide[, -1]))
dim(cor_matrix)
```
Before we select edges (correlations), we need to deal with some redundant data. The correlation matrix is symmetrical along its diagonal. The diagonal will be 1, because it is correlating with itself. Everything else appears twice. We can take care of that by setting the upper (or lower) triangle of this matrix to NA. This step can take a while. The larger the matrix, the slower it is.

```{r}
cor_matrix_upper_tri <- cor_matrix
cor_matrix_upper_tri[lower.tri(cor_matrix_upper_tri)] <- NA
```


```{r}
# Create an empty dataframe to store the results
edge_table <- data.frame()

chunk_size <- 250  # Define the chunk size

num_rows <- nrow(cor_matrix_upper_tri) #number of rows in correlation matrix
num_chunks <- ceiling(num_rows / chunk_size) # number of chunks for data set

for (i in 1:num_chunks) {
  start <- (i - 1) * chunk_size + 1 
  end <- min(i * chunk_size, num_rows)
  
  # Extract a chunk of rows
  chunk <- cor_matrix_upper_tri[start:end, ]
  
  #Wide to long form
  chunk <- chunk %>%
    as.data.frame() %>%
    mutate(from = row.names(chunk)) %>%
    pivot_longer(cols = !from, names_to = "to", values_to = "r") %>%
    filter(is.na(r) == F) %>%
    filter(from != to)
  
  # Add the processed chunk to the result dataframe
  edge_table <- bind_rows(edge_table, chunk)
  
  # Print the current process
  cat("Processed chunk", i, "of", num_chunks, "\n")
}

edge_table <- edge_table %>%
  mutate(t = r*sqrt((number_of_clusters-2)/(1-r^2))) %>%
  mutate(p.value = case_when(
    t > 0 ~ pt(t, df = number_of_clusters-2, lower.tail = F),
    t <=0 ~ pt(t, df = number_of_clusters-2, lower.tail = T)
  )) %>%
  mutate(FDR = p.adjust(p.value, method = "fdr"))

saveRDS(edge_table, file = "/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/KRT_MIT_leaf_edge_table_all_genes.rds")
```

```{r}
edge_table <- readRDS("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/KRT_MIT_leaf_edge_table_all_genes.rds")
```

You can look at various adjusted p value cutoffs and the corresponding r value before proceeding. Let's say we just look at positively correlated genes.

```{r}
edge_table %>% 
  filter(r > 0) %>% 
  filter(FDR < 0.05) %>% 
  slice_min(order_by = abs(r), n = 10)

edge_table %>% 
  filter(r > 0) %>% 
  filter(FDR < 0.01) %>% 
  slice_min(order_by = abs(r), n = 10)
```


> FDR cutoff at 0.05: then your r values are 0.532 or larger. 
> FDR cutoff at 0.01, then your r values are 0.657  or larger. 

### Empirical determination using bait genes and rank distribution 
If I go into this analysis not knowing any biology, then I would proceed with a t approximation followed by some p value cutoff.
I think in real life, this is hardly the case. We usually know something a priori. 
This is where bait genes can be helpful. 
You can use the bait genes to determine the cutoff if you know two bait genes are involved in the same process. 
The underlying assumption is if two bait genes are involved in the same process, they might be co-expressed. 
Because this selection method is based on empirical observations, I argue this is better than using an arbitrary p value cutoff. 
> Skip this until after filtering the edge_table so it doesn't crash


You can also look at the distribution of r values. 
```{r}
edge_table %>% 
  slice_sample(n = 20000) %>% 
  ggplot(aes(x = r)) +
  geom_histogram(color = "white", bins = 100) +
  geom_vline(xintercept = 0.7, color = "tomato1", linewidth = 1.2) +
  theme_classic() +
  theme(
    text = element_text(size = 14),
    axis.text = element_text(color = "black")
  )

ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/r_histogram_r.7.svg", height = 3.5, width = 5, bg = "white")
ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/r_histogram_r.7.png", height = 3.5, width = 5, bg = "white")
```

Cutoff at r = 0.7

Note that there are many negatively correlated genes, we can look at those at well.
But for the sake of this example, let's just look at positively correlated genes. 

```{r}
edge_table_select <- edge_table %>% 
  filter(r >= 0.7)

dim(edge_table_select)
head(edge_table_select)

saveRDS(edge_table_select, file = "/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/KRT_MIT_leaf_edge_table_select_r.7.rds")

edge_table_select <- readRDS("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/KRT_MIT_leaf_edge_table_select_r.7.rds")
```

> We are now down to 36,467,739 edges.

### Check bait genes after filtering for high correlated genes
```{r}
head(Baits)
edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.16-1G012780") & #MsLAMT Ortholog
           str_detect(to,"Mitsp.v1.04-1G008440") | #MsSLS Ortholog
         str_detect(from, "Mitsp.v1.04-1G008440") &
           str_detect(to,"Mitsp.v1.16-1G012780")) 

edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.16-1G012780") & #MsLAMT orth
           str_detect(to,"Mitsp.v1.12-1G013480") | #MsSTRtr orth
         str_detect(from, "Mitsp.v1.12-1G013480") &
           str_detect(to,"Mitsp.v1.16-1G012780")  )

edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.10-1G015530") & #MsDCS1 orth (first alkaloid)
           str_detect(to,"Mitsp.v1.12-1G013480") | #MsSTRtr orth (last iridoid)
         str_detect(from, "Mitsp.v1.12-1G013480") &
           str_detect(to,"Mitsp.v1.10-1G015530"))

edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.10-1G015520") & #MsDCS2 orth (first alkaloid)
           str_detect(to,"Mitsp.v1.12-1G013480") | #MsSTRtr orth (last iridoid)
         str_detect(from, "Mitsp.v1.12-1G013480") &
           str_detect(to,"Mitsp.v1.10-1G015520"))

edge_table_select %>% 
  filter(str_detect(from, "Mitsp.v1.11-1G000250") & #MsEnolMT MIA pathway
           str_detect(to,"Mitsp.v1.12-1G013480") | #MsSTRtr orth (last iridoid)
         str_detect(from, "Mitsp.v1.12-1G013480") &
           str_detect(to,"Mitsp.v1.11-1G000250")  ) 

edge_table_select %>% 
  filter(str_detect(from, "Mitsp.v1.11-1G000250") & #MsEnolMT MIA pathway
           str_detect(to,"Mitsp.v1.16-1G012780") | #MsLAMT orth
         str_detect(from, "Mitsp.v1.16-1G012780") &
           str_detect(to,"Mitsp.v1.11-1G000250")  ) 

edge_table_select %>% 
  filter(str_detect(from, "Mitsp.v1.11-1G000250") & #MsEnolMT MIA pathway
           str_detect(to,"Mitsp.v1.10-1G021390") | #MsCO1
         str_detect(from, "Mitsp.v1.10-1G021390") &
           str_detect(to,"Mitsp.v1.11-1G000250")  ) 

edge_table_select %>% 
  filter(str_detect(from, "Mitsp.v1.11-1G000250") & #MsEnolMT MIA pathway
           str_detect(to,"Mitsp.v1.10-1G010920") | #MsDCR
         str_detect(from, "Mitsp.v1.10-1G010920") &
           str_detect(to,"Mitsp.v1.11-1G000250"))

edge_table_select %>% 
  filter(str_detect(from, "Mitsp.v1.11-1G000250") & #MsEnolMT MIA pathway
           str_detect(to,"Mitsp.v1.18-2G016780") | #MsSAS
         str_detect(from, "Mitsp.v1.18-2G016780") &
           str_detect(to,"Mitsp.v1.11-1G000250")  ) 

edge_table_select %>% 
  filter(str_detect(from, "Mitsp.v1.10-1G010920") & #MsDCR
           str_detect(to,"Mitsp.v1.18-2G016780") | #MsSAS
         str_detect(from, "Mitsp.v1.18-2G016780") &
           str_detect(to,"Mitsp.v1.10-1G010920")  ) 

edge_table_select %>% 
  filter(str_detect(from, "Mitsp.v1.10-1G010920") & #MsDCR
           str_detect(to,"Mitsp.v1.10-1G021390") | #MsCO1
         str_detect(from, "Mitsp.v1.10-1G021390") &
           str_detect(to,"Mitsp.v1.10-1G010920")  ) 
```


> Iridoid genes MsLAMT and MsSLS have correlation of r= 0.8130084
> Iridoid genes MsLAMT and MsSTRtr r=0.7900412
> STRtr and DCS1/STRtr and DCS2 = no correlation
> MsEnolMT and STRtr have no corr
> MsEnolMT and LAMT have corr r=0.8678462	
> MsEnolMT and MsCO1 have no correlation (for 2 dif orthologs of EnolMT: Mitsp.v1.11-1G000250 & Mitsp.v1.11-1G000230)
> MsEnolMT and MsDCR have no corr (for 2 dif orthologs of EnolMT: Mitsp.v1.11-1G000250 & Mitsp.v1.11-1G000230)
> MsEnolMT and MsSAS have no corr (for 2 dif orthologs of EnolMT: Mitsp.v1.11-1G000250 & Mitsp.v1.11-1G000230)
> MsDCR and MsSAS have corr r=0.8309151
> MsDCR and MsCO1 have corr r=0.8474654	

## Module detection
The main goal of a gene co-expression analysis to detect gene co-expression modules, groups of highly co-expressed genes. 
We will be the Leiden algorithm to detect module, which is a graph based clustering method. 
The Leiden method produces clusters in which members are highly interconnected. 
In gene co-expression terms, it looks for groups of genes that are highly correlated with each other. 
If you are interested, you can read more about it in this [review](https://www.nature.com/articles/s41598-019-41695-z ).

### Build graph object 
We will be using `igraph` to do some of the downstream analyses. It will do a lot of the heavy lifting for us. 
While you can get Leiden as a standalone package, Leiden is also part of the `igraph` package. 
The first thing to do is producing a graph object, also known as a network object. 

To make a graph object, you need a edge table. 
We already made that, which is `edge_table_select`, a edge table that we filtered based on some kind of r cutoff. 
Optionally, we can also provide a node table, which contains information about all the nodes present in this network. 
We can make that. 

We need to two things. 

1. Non-redundant gene IDs from the edge table
2. Functional annotation, which I [downloaded](http://spuddb.uga.edu/m82_uga_v1_download.shtml ).

```{r}
mitr_v1.funct_anno <- read_delim("/scratch/ac05869/nih/kratom/mitr_v1_anno/mitr_v1.functional_annotation.txt", delim = "\t", col_names = F, col_types = cols())
mitr_v1.funct_anno$X1 <- gsub(pattern = "_", replacement = "-", mitr_v1.funct_anno$X1)
mitr_v1.funct_anno$X1 <- gsub("\\.\\d+$", "", mitr_v1.funct_anno$X1)

head(mitr_v1.funct_anno)
```

```{r}
node_table <- data.frame(
  gene_ID = c(edge_table_select$from, edge_table_select$to) %>% 
    unique()
  ) %>% 
  left_join(mitr_v1.funct_anno, by = c("gene_ID"="X1"), multiple = "first") %>% 
  rename(functional_annotation = X2)

head(node_table)
dim(node_table)
```

> We have 60851 genes in this network, along with 36,467,739 edges.
Now let's make the network object. 

```{r}
my_network <- graph_from_data_frame(
  edge_table_select,
  vertices = node_table,
  directed = F
)

saveRDS(my_network, file = "/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/KRT_MIT_leaf_network_r.7.rds")
#my_network <- readRDS("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/KRT_MIT_leaf_network_r.7.rds")
```

`graph_from_data_frame()` is a function from the `igraph` package. 
It takes your edge table and node table and produce a graph (aka network) from it. 
Note that I selected the `directed = F` argument, because we made our network using correlation.
Correlation is non-directional, because cor(A,B) = cor(B,A). 

### What is the optimal resolution for module detection? 
The optimal resolution for module detection differs between networks. 
A key factor that contributes to the difference in optimal resolution is to what extent are nodes inter-connected. 

Since this is a simple workflow, we can determine the optimal resolution using heuristics. 
We can test a range of resolutions and monitor two key performance indexes:

1. Optimize number of modules that have >= 5 genes.
2. Optimize number of genes that are contained in modules that have >= 5 genes. 

Because: 

* Too low resolution leads to forcing genes with different expression patterns into the same module.
* Too high resolution leads to many genes not contained in any one module. 

```{r}
optimize_resolution <- function(network, resolution){
  modules = network %>% 
    cluster_leiden(resolution = resolution,
                   objective_function = "modularity")
  
  parsed_modules = data.frame(
    gene_ID = names(membership(modules)),
    module = as.vector(membership(modules)) 
    )
  
  num_module_5 = parsed_modules %>% 
    group_by(module) %>% 
    count() %>% 
    arrange(-n) %>% 
    filter(n >= 5) %>% 
    nrow() %>% 
    as.numeric()
  
  num_genes_contained = parsed_modules %>% 
    group_by(module) %>% 
    count() %>% 
    arrange(-n) %>% 
    filter(n >= 5) %>% 
    ungroup() %>% 
    summarise(sum = sum(n)) %>% 
    as.numeric()
  
  cbind(num_module_5, num_genes_contained) %>% 
    as.data.frame()

}
```

Here I wrote a function to detect module, pull out number of modules that have >= 5 genes, and count number of genes contained in modules that have >= 5 genes. All in one function. 

Then I can test a list of resolutions in this function. 
Let's test a range of resolution from 0.25 to 5, in steps of 0.25.  
```{r}
 optimization_results <- purrr::map_dfr(
  .x = seq(from = 0.25, to = 5, by = 0.25),
  .f = optimize_resolution, 
  network = my_network
) %>% 
  cbind(
   resolution = seq(from = 0.25, to = 5, by = 0.25)
  ) %>% 
  as.data.frame() %>% 
  rename(num_module = num_module_5,
         num_contained_gene = num_genes_contained)

head(optimization_results)
```

This could take a while. 
We have the results organized into one tidy data table. We can graph it. 
```{r}
Optimize_num_module <- optimization_results %>% 
  ggplot(aes(x = resolution, y = num_module)) +
  geom_line(size = 1.1, color = "dodgerblue2") +
  geom_point(size = 3) +
  geom_vline(xintercept = 2, size = 1, linetype = 4) +
  labs(x = "resolution parameter",
       y = "num. modules\nw/ >=5 genes") +
  theme_classic() +
  theme(
    text = element_text(size = 14),
    axis.text = element_text(color = "black")
  )

Optimize_num_gene <- optimization_results %>% 
  ggplot(aes(x = resolution, y = num_contained_gene)) +
  geom_line(size = 1.1, color = "violetred2") +
  geom_point(size = 3) +
  geom_vline(xintercept = 2, size = 1, linetype = 4) +
  labs(x = "resolution parameter",
       y = "num. genes in\nmodules w/ >=5 genes") +
  theme_classic() +
  theme(
    text = element_text(size = 14),
    axis.text = element_text(color = "black")
  )

wrap_plots(Optimize_num_module, Optimize_num_gene, nrow = 2)

ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/Optimize_resolution.svg", height = 5, width = 3.2, bg ="white")
ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/Optimize_resolution.png", height = 5, width = 3.2, bg ="white")
```

You can see that there is a big jump for num. modules w/ >= 5 genes going from # to # resolution.
The number of modules stabilizes at resolution >=#
However, if you look at number of contained genes, the story is a little different. 
The number of contained genes is very stable until resolution > 2.5, after which the number of genes continues to diminish. 

> In this case going with res. = 3

### Graph based clustering
The next step is detect modules from the graph object. 
```{r}
modules <- cluster_leiden(my_network, resolution = 3, 
                          objective_function = "modularity")

```

`cluster_leiden()` runs the Leiden algorithm for you. 
`resolution_parameter` controls how many clusters you will get. The larger it is, the more clusters. 
You can play around with the resolution and see what you get. 
The underlying math of `objective_function` is beyond me, but it specifies how the modules are computed. 

Let's say we move on with module detection using a resolution of # 
Next, we need to link the module membership to the gene IDs.
```{r}
my_network_modules <- data.frame(
  gene_ID = names(membership(modules)),
  module = as.vector(membership(modules)) 
) %>% 
  inner_join(node_table, by = "gene_ID")

my_network_modules %>% 
  group_by(module) %>% 
  count() %>% 
  arrange(-n) %>% 
  filter(n >= 5)

my_network_modules %>% 
  group_by(module) %>% 
  count() %>% 
  arrange(-n) %>% 
  filter(n >= 5) %>% 
  ungroup() %>% 
  summarise(sum = sum(n))

```

Looks like there are ~24 modules that have 5 or more genes, comprising ~60629 genes.

Not all genes are contained in modules. They are just lowly connected genes. 

Note that Leiden clustering has a stochastic aspect. The membership maybe slightly different every time you run it. 
Moving forward we will only use modules that have 5 or more genes. 

```{r}
module_5 <- my_network_modules %>% 
  group_by(module) %>% 
  count() %>% 
  arrange(-n) %>% 
  filter(n >= 5)

my_network_modules <- my_network_modules %>% 
  filter(module %in% module_5$module)

head(my_network_modules)
#View(my_network_modules)

saveRDS(my_network_modules, file = "/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/my_network_modules_5_genes.rds")
```

### Module quality control
We have a bunch of different modules now, how do we know if they make any sense? 
One way to QC these modules is looking at our bait genes. 

Orthologs of CO1: 
Mitsp.v1.01-1G019680, Mitsp.v1.01-1G019690, Mitsp.v1.01-1G019700, Mitsp.v1.01-2G001330, Mitsp.v1.01-2G001340, Mitsp.v1.01-2G001350, Mitsp.v1.01-2G019310, Mitsp.v1.01-2G019320, Mitsp.v1.01-2G019330, Mitsp.v1.03-1G014910, Mitsp.v1.03-2G015030, Mitsp.v1.10-1G021390, Mitsp.v1.10-1G021550, Mitsp.v1.10-2G022130, Mitsp.v1.10-2G022370, Mitsp.v1.20-1G017760, Mitsp.v1.20-1G017770, Mitsp.v1.20-2G016840, Mitsp.v1.20-2G016850
```{r}
head(Baits)
(Iridoids <- my_network_modules %>% 
  filter(gene_ID == "Mitsp.v1.16-1G012780" | #Iridoid MsLAMT, first switch of clusters from MEP
           gene_ID == "Mitsp.v1.04-1G008440" | # MsSLS1
           gene_ID == "Mitsp.v1.07-1G009190" | #SLTr
           gene_ID == "Mitsp.v1.16-2G020400" | #TDC
           gene_ID == "Mitsp.v1.21-1G001170" | #STR
           gene_ID == "Mitsp.v1.12-1G013480")) #STRtr

(CO1_ortho <- my_network_modules %>% 
  filter(gene_ID == "Mitsp.v1.10-2G022130" | #MIA CO1 top hit
           gene_ID =="Mitsp.v1.01-1G019680" |
           gene_ID =="Mitsp.v1.01-1G019690" |
           gene_ID =="Mitsp.v1.01-1G019700" |
           gene_ID =="Mitsp.v1.01-2G001330" |
           gene_ID =="Mitsp.v1.01-2G001340" |
           gene_ID =="Mitsp.v1.01-2G001350" |
           gene_ID =="Mitsp.v1.01-2G019310" |
           gene_ID =="Mitsp.v1.01-2G019320" |
           gene_ID =="Mitsp.v1.01-2G019330" |
           gene_ID =="Mitsp.v1.03-1G014910" |
           gene_ID =="Mitsp.v1.03-2G015030" |
           gene_ID =="Mitsp.v1.10-1G021390" | #ortholog used in baits
           gene_ID =="Mitsp.v1.10-1G021550" |
           gene_ID =="Mitsp.v1.10-2G022370" |
           gene_ID =="Mitsp.v1.20-1G017760" |
           gene_ID =="Mitsp.v1.20-1G017770" |
           gene_ID =="Mitsp.v1.20-2G016840" |
           gene_ID =="Mitsp.v1.20-2G016850" 
         ) )

my_network_modules %>% 
  filter(gene_ID == "Mitsp.v1.16-1G012780" | ##MsDCR (MIA)
           gene_ID == "Mitsp.v1.04-1G008440" | ##MsSAS (MIA)
           gene_ID == "Mitsp.v1.10-1G021390") ##MsCO1 (MIA)

```

> (Iridoid) STR in mod 9
> (Iridoid) SLS, SLTr, and TDC in module 11
> (Iridoid) LAMT and STRtr in mod 7
> (MIA) MsSAS in module 11 
> (MIA) DCR in mod 7
> (MIA) CO1 orthologs are in modules: 4, 10, 20, 11 (top ortholog Mitsp.v1.10-1G021390 in 10)

edge_table_select %>% 
  filter(str_detect(from, "Mitsp.v1.10-1G010920") & #MsDCR
           str_detect(to,"Mitsp.v1.18-2G016780") | #MsSAS
         str_detect(from, "Mitsp.v1.18-2G016780") &
           str_detect(to,"Mitsp.v1.10-1G010920")  ) 

edge_table_select %>% 
  filter(str_detect(from, "Mitsp.v1.10-1G010920") & #MsDCR
           str_detect(to,"Mitsp.v1.10-1G021390") | #MsCO1
         str_detect(from, "Mitsp.v1.10-1G021390") &
           str_detect(to,"Mitsp.v1.10-1G010920")  ) 

If they are highly correlated, then they are in the same module

## Module-treatment correspondance
The next key task is understanding the expression pattern of the clusters. 
Again, the essence of this workflow is simple, so we will use a simple method: peak expression.
To do that, we append the module membership data back to the long table containing z scores. 

```{r}
Exp_table_long_averaged_z_modules <- Exp_table_long_averaged_z %>% 
  inner_join(my_network_modules, by = "gene_ID")

head(Exp_table_long_averaged_z_modules)
#View(Exp_table_long_averaged_z_modules)
```

```{r}
#Annotate clusters by cell type
Exp_table_long_averaged_z_modules <- Exp_table_long_averaged_z_modules %>%
 mutate(cell_type = case_when(
    cluster == 0 | 
      cluster == 3 | 
      cluster == 6 | 
      cluster == 7 | 
      cluster == 13 |
      cluster == 14 ~ "Mesophyll",
    cluster == 1 | 
      cluster == 9 |
      cluster == 11 | 
      cluster == 15 | 
      cluster == 18 ~ "Epidermis",
    cluster == 4 |
      cluster == 8 |
      cluster == 16 ~ "Xylem",
    cluster == 2 |
      cluster == 12 |
      cluster == 20 ~ "Phloem",
    cluster == 5 |
      cluster == 10 ~ "Proliferating",
    cluster == 19 ~ "MEP+",
      cluster == 21 ~ "IPAP",
    T ~ "Unassigned"))
    
Exp_table_long_averaged_z_modules

write_excel_csv(Exp_table_long_averaged_z_modules, "/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/Exp_table_long_averaged_z_modules_r0.7_3res.csv")
```

Now we can produce summary statistics for each cluster and look at their expression pattern using mean. 
```{r}
modules_mean_z <- Exp_table_long_averaged_z_modules %>% 
  group_by(module, cluster, cell_type) %>% 
  summarise(mean.z = mean(z.score)) %>% 
  ungroup()

head(modules_mean_z)
#View(modules_mean_z)

modules_mean_z2 <- Exp_table_long_averaged_z_modules %>% 
  group_by(cluster, module, cell_type) %>% 
  summarise(mean.z = mean(z.score)) %>% 
  ungroup()

head(modules_mean_z2)
#View(modules_mean_z2)
```

Then we look at at which cluster is each module most highly expressed. 
```{r}
module_peak_exp <- modules_mean_z %>% 
  group_by(cluster) %>% 
  group_by(module) %>% #added this because some modules did not have peak expression when only grouped by cluster
  slice_max(order_by = mean.z, n = 1)

module_peak_exp
#View(module_peak_exp)
```
Again, `group_by()` is doing a lot of heavy lifting here. 

### More module QC
You can also QC the clusters via a line graph
> Notes
* module 7: where iridoid bait genes are. 

```{r}
ggplot(Exp_table_long_averaged_z_modules, aes(x=cluster, y=z.score)) +
  geom_line(aes(group = gene_ID), color = "grey70") +
  geom_line(
    data = modules_mean_z,
    aes(y = mean.z, group = module), 
   linewidth = 1.1
  ) +
  labs(x = NULL,
       y = "z score") +
  theme_classic() +
  facet_wrap( ~ module, ncol = 5) +
  theme(
    text = element_text(size = 16),
    axis.text = element_text(color = "black"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
    panel.spacing = unit(1, "line")
  )

#ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/module_line_plots_3res.svg", height = 4, width = 8.2, bg = "white")
#ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/module_line_plots_3res.png", height = 4, width = 8.2, bg = "white")
```

### Heat map representation of clusters 
A good way to present these modules is to make a heat map. 
To make an effective heatmap though, we need to take care of a few things.

* reorder x and y axis
* take care of outliers 

#### Check outliers 
Let's take care of outliers first 
```{r}
modules_mean_z$mean.z %>% summary()
```
You can see that the distribution of averaged z scores are more or less symmetrical from the 1st to 3rd quartiles. 
```{r}
quantile(modules_mean_z$mean.z, 0.95)
```
The 95th percentile of averaged z score is 1.31 . We can probably roughly clipped the z-scores at 1.5 or -1.5

> 1.31, clipped at 1.5

```{r}
modules_mean_z <- modules_mean_z %>% 
  mutate(mean.z.clipped = case_when(
    mean.z > 1.5 ~ 1.5,
    mean.z < -1.5 ~ -1.5,
    T ~ mean.z
  ))
```

> This sets z scores > 1.5 or < -1.5 to 1.5 or -1.5, respectively. The rest remain unchanged.  

#### Reorder rows and columns 
Let's say we graph modules on y axis, and stage/tissue (cluster) on x-axis.
Reordering columns are easy, we just do it by hand. 
We already did it before. We can copy and paste that down here. 
```{r}
modules_mean_z2 <- modules_mean_z %>% 
#  mutate(cluster = factor(cluster, levels = c(0:22))) %>%
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll", "Epidermis", "Guard cells",
                              "Xylem", "Phloem",
                              "Proliferating", "MEP+", "IPAP", "Unassigned"
                            ))) %>% 
  mutate(cluster = factor(cluster, levels = c(
    "0", "3", "6", "7", "13", "14", #M
    "1","9", "11", "15", "18", #Epi
    "4", "8", "16", #X
    "2", "12", "20", #Ph
    "5", "10", #prolif
    "19", ##Unassigned/MEP+
    "21", #IPAP
   "17", "22" #Unknown
  )))
head(modules_mean_z2)
```

```{r}
my_data_peak_values <- modules_mean_z2 %>% 
  group_by(cluster) %>% #this used to be module
  slice_max(order_by = mean.z, n = 1, with_ties = F) %>% 
  rename(peaked_at = cluster) %>% 
  select(-mean.z, cell_type)

head(my_data_peak_values)

number_of_peaks <- my_data_peak_values %>% 
  group_by(peaked_at) %>% 
  count() %>% 
  arrange(-n)
head(number_of_peaks)

my_data_peak_values_reordered <- my_data_peak_values %>% 
  inner_join(number_of_peaks, by = "peaked_at") %>% 
  arrange(-n)  
my_data_peak_values_reordered

my_data_reordered <- modules_mean_z2 %>% 
  inner_join(number_of_peaks, by = c("cluster" = "peaked_at")) %>% 
  mutate(cluster = reorder(cluster, -n)) %>%  # this reorders the columns
  select(-n) %>% 
  inner_join(my_data_peak_values_reordered, by = "module", multiple = "first") %>% 
  mutate(peaked_at = reorder(peaked_at, -n)) %>% 
  mutate(order_rows = as.numeric(peaked_at)) %>% 
  mutate(module = reorder(module, -order_rows)) # this reorders the rows by the "peaked_at" column. 
head(my_data_reordered)

module_heatmap <- my_data_reordered %>% 
  ggplot(aes(x = cluster, y = module)) +
  geom_tile(aes(fill = mean.z.clipped.x), color = "grey80") +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu")), limits = c(-1.5, 1.5),
                       breaks = c(-1.5, 0, 1.5), labels = c("< -1.5", "0", "> 1.5")) +
  labs(fill = "mean.z") +
  theme_classic() +
  theme(
    text = element_text(size = 16),
    axis.text = element_text(color = "black"),
    axis.text.x = element_text(color = "black"),
    strip.text = element_blank(),
    legend.position = "top",
    panel.spacing = unit(0.5, "lines") 
  )

module_heatmap
```

This is missing modules that have peek expression in the same cluster

```{r}
my_data_peak_values2 <- modules_mean_z2 %>% 
  group_by(cluster) %>%
  group_by(module) %>%
  slice_max(order_by = mean.z, n = 1, with_ties = F) %>% 
  rename(peaked_at = cluster) %>% 
  select(-mean.z, cell_type)

head(my_data_peak_values2)

number_of_peaks2 <- my_data_peak_values2 %>% 
  group_by(peaked_at) %>% 
  count() %>% 
  arrange(-n)
head(number_of_peaks2)

my_data_peak_values_reordered2 <- my_data_peak_values2 %>% 
  inner_join(number_of_peaks2, by = "peaked_at") %>% 
  arrange(-n)  
my_data_peak_values_reordered2

my_data_reordered2 <- modules_mean_z2 %>% 
  inner_join(number_of_peaks, by = c("cluster" = "peaked_at")) %>% 
  mutate(cluster = reorder(cluster, -n)) %>%  # this reorders the columns
  select(-n) %>% 
  inner_join(my_data_peak_values_reordered2, by = "module") %>% 
  mutate(peaked_at = reorder(peaked_at, -n)) %>% 
  mutate(order_rows = as.numeric(peaked_at)) %>% 
  mutate(module = reorder(module, -order_rows)) # this reorders the rows by the "peaked_at" column. 
head(my_data_reordered2)

module_heatmap2 <- my_data_reordered2 %>% 
  ggplot(aes(x = cluster, y = module)) +
  geom_tile(aes(fill = mean.z.clipped.x), color = "grey80") +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu")), limits = c(-1.5, 1.5),
                       breaks = c(-1.5, 0, 1.5), labels = c("< -1.5", "0", "> 1.5")) +
  labs(fill = "mean.z") +
  theme_classic() +
  theme(
    text = element_text(size = 14),
    axis.text = element_text(color = "black"),
    axis.text.x = element_text(color = "black"),
    strip.text = element_blank(),
    legend.position = "top",
    panel.spacing = unit(0.5, "lines") 
  )

module_heatmap2

#ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/module_heatmap2_3res.svg", height = 4.8, width = 10, bg = "white")
#ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/module_heatmap2_3res.png", height = 4.8, width = 10, bg = "white")
```

When the rows and columns are re-ordered, you can trace the signal down the diagonal from upper left to lower right. 

```{r}
leaf_cell_type_assignment <- data.frame(
  cluster = c(0:22)
) %>% 
  mutate(cell_type = case_when(
    cluster == 0 | 
      cluster == 3 | 
      cluster == 6 | 
      cluster == 7 | 
      cluster == 13 |
      cluster == 14 ~ "Mesophyll",
    cluster == 1 | 
      cluster == 9 |
      cluster == 11 | 
      cluster == 15 | 
      cluster == 18 ~ "Epidermis",
    cluster == 4 |
      cluster == 8 |
      cluster == 16 ~ "Xylem",
    cluster == 2 |
      cluster == 12 |
      cluster == 20 ~ "Phloem",
    cluster == 5 |
      cluster == 10 ~ "Proliferating",
    cluster == 19 ~ "MEP+",
      cluster == 21 ~ "IPAP",
    T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll", "Epidermis", "Guard cells",
                              "Xylem", "Phloem",
                              "Proliferating", "MEP+", "IPAP", "Unassigned"
                            ))) %>% 
  mutate(cluster = factor(cluster, levels = c(
    "0", "3", "6", "7", "13", "14", #M
    "1","9", "11", "15", "18", #Epi
    "4", "8", "16", #X
    "2", "12", "20", #Ph
    "5", "10", #prolif
    "19", ##IPAP/MEP+
    "21", #IPAP
   "17", "22" #Unknown
  )))

Leaf_cell_type_strip <- leaf_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "" )) +
  geom_tile(aes(fill = cell_type)) +
  scale_fill_manual(values = c(brewer.pal(7, "Accent"), "grey80")) +
  guides(fill = guide_legend(nrow = 2)) +
  labs(fill = "Cell type") +
  theme_void() +
  theme(
    legend.position = "bottom" ,
    text = element_text(size = 14)
  )

Leaf_cell_type_strip
graph <- wrap_plots(module_heatmap2, Leaf_cell_type_strip,
                    nrow = 2, heights = c(1, 0.03))
graph

#ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/module_heatmap2_3res.svg", height = 5.3, width = 10, bg = "white")
#ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/module_heatmap2_3res.png", height = 5.3, width = 10, bg = "white")
```


## Gene co-expression graphs 
A common data visualization for gene co-expression analyses is network graphs. 
We will be using `ggraph`, a `ggplot` extension of `igraph`. 

Our network has almost # genes and more than 1 million edges. 
It's too much to look at if we graph the full network. 
On the other hand, there is not much to look at anyway for very large networks. 
You just get messy hairballs. 

Say we want to look at genes directly co-expressed with our bait genes. 
We can pull out their neighbors using the `neighbors()` function within `igraph()`.
`igraph` comes with a set of network analysis functions that we can call. 

> (Iridoid) STR in mod 9
> (Iridoid) SLS, SLTr, and TDC in module 11
> (Iridoid) LAMT and STRtr in mod 7
> (MIA) MsSAS in module 11 
> (MIA) DCR in mod 7
> (MIA) CO1 orthologs are in modules: 4, 10, 20, 11 (top ortholog Mitsp.v1.10-1G021390 in 10)

```{r}
head(Baits)
neighbors_of_bait <- c(
  neighbors(my_network, v = "Mitsp.v1.18-2G016780"), #MsSAS ortholog
  neighbors(my_network, v = "Mitsp.v1.10-1G010920"), #MsDCR
  neighbors(my_network, v = "Mitsp.v1.22-2G007110"), #Ms10H
  neighbors(my_network, v = "Mitsp.v1.10-1G021390"), #MsCO1
  neighbors(my_network, v = "Mitsp.v1.11-1G000250"),#MsEnolMT
  neighbors(my_network, v = "Mitsp.v1.10-1G015530"), #DCS1
  neighbors(my_network, v = "Mitsp.v1.10-1G015520")) %>% #DCS2
  unique()  
length(neighbors_of_bait)
```

>2102 genes

We can make a sub-network object. 
First we subset edges in the network.
```{r}
subnetwork_edges <- edge_table_select %>% 
  filter(from %in% names(neighbors_of_bait) &
           to %in% names(neighbors_of_bait)) %>% 
  group_by(from) %>% 
  slice_max(order_by = r, n = 5) %>% 
  ungroup() %>% 
  group_by(to) %>% 
  slice_max(order_by = r, n = 5) %>% 
  ungroup()

subnetwork_genes <- c(subnetwork_edges$from, subnetwork_edges$to) %>% unique()
length(subnetwork_genes)
dim(subnetwork_edges)
```


> 2092 nodes
> 5477 edges

We can constrain the edges such that both the start and end of edges are neighbors of baits. 
I also filtered for highly correlated neighbors (top 5 edges/node based on r value). 
We still have # edges and # nodes. 
Note that the most correlated edges for each bait many have overlaps, so the total number of edges remaining will be less than what you think. 

Then we subset nodes in the network. 
```{r}
subnetwork_nodes <- node_table %>% 
  select(!functional_annotation) %>%
  filter(gene_ID %in% subnetwork_genes) %>% 
  left_join(my_network_modules, by = "gene_ID") %>% 
  left_join(module_peak_exp, by = "module") %>% #module 1 and 19 have no peak expression
  mutate(module_annotation = case_when(
    module == "9" ~ "STR",
    module == "11" ~ "SLS, SLTr, TDC, SAS",
    module == "7" ~ "LAMT, STRtr, DCR",
    module == "10" ~ "CO1",
    T ~ "other"
  ))
dim(subnetwork_nodes)

head(subnetwork_nodes)
```
I also append the data from module peak expression and add a new column called "module annotation".
> (Iridoid) STR in mod 9
> (Iridoid) SLS, SLTr, and TDC in module 11
> (Iridoid) LAMT and STRtr in mod 7
> (MIA) MsSAS in module 11 
> (MIA) DCR in mod 7
> (MIA) CO1 orthologs are in modules: 4, 10, 20, 11 (top ortholog Mitsp.v1.10-1G021390 in 10)

Then make sub-network object from subsetted edges and nodes. 
```{r}
my_subnetwork <- graph_from_data_frame(subnetwork_edges,
                                     vertices = subnetwork_nodes,
                                     directed = F)

```
Use `graph_from_data_frame()` from `igraph` to build the sub-network.
There are ways to directly filter existing networks, but I always find it more straightforward to build sub-network de novo from filtered edge and node tables.

```{r}
 my_subnetwork %>% 
  ggraph(layout = "kk", circular = F) +
  geom_edge_diagonal(color = "grey70", width = 0.5) +
  geom_node_point(color = "white", shape = 21, size = 2,
                  aes(fill = module_annotation)) + 
  scale_fill_manual(values = c(brewer.pal(8, "Accent")[c(1,3,6)], "grey30")) +
  labs(fill = "Modules") +
  guides(size = "none",
         fill = guide_legend(override.aes = list(size = 4), 
                             title.position = "top", nrow = 2)) +
  theme_void()+
  theme(
    text = element_text(size = 14), 
    legend.position = "bottom",
    legend.justification = 1,
    title = element_text(size = 12)
  )

ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/subnetwork_graph_12mar25.svg", height = 5, width = 4, bg = "white")
ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/subnetwork_graph_12mar25.png", height = 5, width = 4, bg = "white")
```

This could take a while. It is trying to draw many many lines and many dots. 
Unsurprisingly, we get a bunch of distinct hairballs. 
A good advice here is to check different graph layouts. 
The layout of the graphs can have a **huge** impact on the appearance of the network graph. 
See [igraph layouts](https://igraph.org/r/doc/layout_.html), [ggraph layouts](https://www.data-imaginist.com/2017/ggraph-introduction-layouts/), and [trying different layouts](c) for more information. 

> Left off here 12March24

# Mean separation plots for candidate genes 
## Pull out direct neighbors 
We did a bunch of analyzes, now what? 
A common "ultimate" goal for gene co-expression analyses is to find new candidate genes, which are genes co-expressed with bait genes. 
After doing network analysis, this is very easy to find. 
We can either look at what other genes are in module #, which both our bait genes are in, or we can look at direct neighbors of bait genes. 
`igraph` comes with a set of network analysis functions that we can call. 

And we already did that earlier for the sub-network. 
* "Mitsp.v1.16-1G012780" MsLAMT
* "Mitsp.v1.04-2G008810" MsSLS
* "Mitsp.v1.10-2G016180" MsDCS1
* "Mitsp.v1.18-2G016770" MsSAS
```{r}
neighbors_of_LAMT_SLS <- c(
  neighbors(my_network, v = "Mitsp.v1.16-1G012780"), # MsLAMT
  neighbors(my_network, v = "Mitsp.v1.04-2G008810") # MsSLS 
) %>% 
  unique()  

length(neighbors_of_LAMT_SLS)
```
> Early MIA genes: 220 direct neighbors of MsLAMT &  MsSLS

```{r}
neighbors_of_DCS1_SAS <- c(
  neighbors(my_network, v = "Mitsp.v1.10-2G016180"), # MsDCS1
  neighbors(my_network, v = "Mitsp.v1.18-2G016770") # MsSAS
) %>% 
  unique()  

length(neighbors_of_DCS1_SAS)
```

> Late MIA genes: 758 direct neighbors of MsDCS1 & MsSAS

Let's say you are interested in transcription factors (TFs). 
There are many types of TFs. Let's say you are particularly interested in bHLH and GRAS type TFs. 
```{r}
#my_TFs <- my_network_modules %>% 
#  filter(gene_ID %in% names(neighbors_of_LAMT_SLS)) %>% 
#  filter(str_detect(functional_annotation, "GRAS|bHLH"))

#my_TFs
```

```{r}
#TF_TPM <- Exp_table_long %>% 
#  filter(gene_ID %in% my_TFs$gene_ID) %>% 
#  inner_join(PCA_coord, by = c("library"="Run")) %>% 
#  filter(dissection_method == "Hand") %>% 
#  mutate(order_x = case_when(
#    str_detect(cluster, "5") ~ 1,
#    str_detect(cluster, "10") ~ 2,
#    str_detect(cluster, "20") ~ 3,
#    str_detect(cluster, "30") ~ 4,
#    str_detect(cluster, "MG") ~ 5,
#    str_detect(cluster, "Br") ~ 6,
#    str_detect(cluster, "Pk") ~ 7,
#    str_detect(cluster, "LR") ~ 8,
#    str_detect(cluster, "RR") ~ 9
#  )) %>% 
#  mutate(cluster = reorder(cluster, order_x)) %>% 
#  mutate(tag = str_remove(gene_ID, "Solly.M82.")) %>% 
#  mutate(tag = str_remove(tag, ".\\d+$")) %>% 
#  ggplot(aes(x = cluster, y = logTPM)) +
#  facet_grid(tag ~ tissue, scales = "free_y") +
#  geom_point(aes(fill = tissue), color = "white", size = 2, 
#             alpha = 0.8, shape = 21, position = position_jitter(0.1, seed = 666)) +
#  stat_summary(geom = "line", aes(group = gene_ID), 
#               fun = mean, alpha = 0.8, size = 1.1, color = "grey20") +
#  scale_fill_manual(values = brewer.pal(8, "Set2")) +
#  labs(x = NULL,
#       y = "log10(TPM)") +
#  theme_bw() +
#  theme(
#    legend.position = "none",
#    panel.spacing = unit(1, "lines"),
#    text = element_text(size = 14),
#    axis.text = element_text(color = "black"),
#    axis.text.x = element_blank(),
#    strip.background = element_blank(),
#    strip.text = element_text(size = 10)
#  )

#wrap_plots(TF_TPM, module_lines_color_strip, 
#           nrow = 2, heights = c(1, 0.05))

#ggsave("../Results/Candidate_genes_TPM.svg", height = 4.8, width = 8, bg = "white")
#ggsave("../Results/Candidate_genes_TPM.png", height = 4.8, width = 8, bg = "white")
```

## Write out results
Finally, I want to write out the neighbors of out bait genes as a table onto the hard drive. 
That's easy. 

```{r}
Bait_neighors <- mitr_v1.funct_anno %>% 
  filter(X1 %in% names(neighbors_of_LAMT_SLS)) %>% 
  rename(Gene_ID = X1,
         annotation = X2)

head(Bait_neighors)
write_excel_csv(Bait_neighors, "/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/neighbors_of_LAMT_SLS.csv", col_names = T)

Bait_neighors2 <- mitr_v1.funct_anno %>% 
  filter(X1 %in% names(neighbors_of_DCS1_SAS)) %>% 
  rename(Gene_ID = X1,
         annotation = X2)

head(Bait_neighors2)
write_excel_csv(Bait_neighors2, "/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/neighbors_of_DCS1_SAS.csv", col_names = T)

write_excel_csv(modules_mean_z_reorded, "/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/modules_mean_z_reordered.csv")
write_excel_csv(my_network_modules, "/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/my_network_modules.csv")
```


# Conclusions
Well, we are pretty much done. 
Now you just need to send the list of candidate genes and the nice graphics to your wet lab folks. 
Hopefully they find something interesting at the lab bench. 


