---
title: "Plumeria leaf single cell"
author: "Anne Pavia"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(Seurat) 
library(readxl)
library(RColorBrewer)
library(viridis)
library(svglite) 
library(patchwork)
library(cowplot)
library(ggpubr)
library(showtext)
```
#Notes: 
Started October 31 with Seurat v5
## leaf 
reps 1 and 2 with two flow cell runs each; protoplasts; PIPseq v5 kit, pipseeker-v3.3.0-linux; #EM algorithm #mapped on plob_v4 assembly
>PLU_BB: ~35,775 protoplast input, expect ~17,887 out
>PLU_BC: ~35,775 protoplast input, expect ~17,887 out

# Load Data for leaf

```{r}
#Note that you have to gzip these files, as Seurat expects them to be zipped to match CellRangers output (zipped in CellRanger output from v3 or higher). STARsolo did not gzip these by default

PLU_BB_raw <- Read10X(data.dir = "/scratch/ac05869/plob_sc/STARsolo/PLU_BB_Solo.out/GeneFull/raw/data_for_R/EM_EmptyDrops_Combined")
PLU_BC_raw <- Read10X(data.dir = "/scratch/ac05869/plob_sc/STARsolo/PLU_BC_Solo.out/GeneFull/raw/data_for_R/EM_EmptyDrops_Combined")

```

# Create Seurat Objects

```{r}
BB <- CreateSeuratObject(counts = PLU_BB_raw, min.cells = 3, min.features = 50,
                         project = "PLU_BB")
BC <- CreateSeuratObject(counts = PLU_BC_raw, min.cells = 3, min.features = 50,
                         project = "PLU_BC")
```
## Summary Statistics
###Before any cleaning
```{r}
summary(BB$nFeature_RNA)
summary(BB$nCount_RNA)

summary(BC$nFeature_RNA)
summary(BC$nCount_RNA)
```
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    266     757    1276    2074    2509   15284 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    597     888    1552    3282    3451   48368 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    377     663    1082    1643    1999   16029 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    498     750    1261    2290    2515   49647 

# Data QC
You may or may not be able to do organellar filtering based on what you aligned to - take notes of organellar filtering. How much % remove?
## Organelles

Data QC - Chloroplast and Mitochondrion Flags
Add column with organellar percentage to filter data on 
The [[ operator can add columns to object metadata. This is a great place to stash QC stats

Plumera obtusa SPECIFIC EXAMPLE
Any gene that doesn't start with ^Pluob is an organellar gene
#Check the minimum and maximum percentage of organellar reads in a given cell
```{r}
##rep1
BB[["percent.normal"]] <- PercentageFeatureSet(BB, pattern = "^Pluob")
BB[["percent.organellar"]] <- (100 - BB[["percent.normal"]])

min(BB[["percent.normal"]]) #34.98049
max(BB[["percent.normal"]]) #100

min(BB[["percent.organellar"]])#0
max(BB[["percent.organellar"]])#65.01951

##rep2

BC[["percent.normal"]] <- PercentageFeatureSet(BC, pattern = "^Pluob")
BC[["percent.organellar"]] <- (100 - BC[["percent.normal"]])

min(BC[["percent.normal"]])#41.72892
max(BC[["percent.normal"]])#100

min(BC[["percent.organellar"]])#0
max(BC[["percent.organellar"]])#58.27108
```

## Plotting QC Values
```{r}
VlnPlot(BB, features = c("nFeature_RNA", "nCount_RNA", "percent.organellar"), ncol = 3, raster = FALSE)
ggsave("/scratch/ac05869/plob_sc/seurat_qc/BB_violin_plots.png", height = 8, width = 8, bg = "white")

#nCount = number of UMIs per cell
#nFeature = number of Genes per cell

VlnPlot(BC, features = c("nFeature_RNA", "nCount_RNA", "percent.organellar"), ncol = 3, raster = FALSE)
ggsave("/scratch/ac05869/plob_sc/seurat_qc/BC_violin_plots.png", height = 8, width = 8, bg = "white")

#nCount = number of UMIs per cell
#nFeature = number of Genes per cell
```

## Organellar Filtering
```{r}
#this is for Chloroplast and Mitochondria, I just name it "chloro_filt" 

BB #14858 cells
BB_chloro_filt <- subset(BB, subset = percent.organellar < 5) #remove all cells with above 5% organellar reads
BB_chloro_filt # 13507

BC #13850
BC_chloro_filt <- subset(BC, subset = percent.organellar < 5) #remove all cells with above 5% organellar reads
BC_chloro_filt #12718
```

## Organellar Filtering
```{r}
#Check the new distrubution of organellar read percentage now that the data is filtered and easier to see. May need to adjust filtering!

VlnPlot(BB_chloro_filt, features = c("percent.organellar"), raster = FALSE)

#LIB_ID: 5% 
# Starts:   Ends:   Removed: 

VlnPlot(BC_chloro_filt, features = c("percent.organellar"), raster = FALSE)

#LIB_ID: 5% 
# Starts:   Ends:   Removed: 
```

## Plotting for Cutoff Selection
```{r}
#blank data frame/plot for helping to organize plots 
#every single dataset is different. BLUE: UMIs. Make so that one read to one gene is the minimum. With a lot of cells we can be more stringent and throw out more. Get rid of "extremes". What is the mean UMIs per gene? This is a metric that people might pay more attention to in the future.

blank <- data.frame(
  x = 1:10000,
  y = 1:10000
) %>% 
  ggplot(aes(x = x, y = y)) +
  theme_void()
```
###Rep1
```{r}
#You should generate a plot with test values and adjust based on the dataset!
#I plot the full dataset and the organellar filtered dataset to be able to visualize how the organellar filtering impacts the data

BB_scatter <- BB@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point() + ggtitle("PLU_BB pipseq") +
  scale_x_continuous(trans='log10') +
  scale_y_continuous(trans='log10') +
  geom_hline(yintercept=500, color = "red") + #Set minimum number of genes to be expressed per cell
  geom_hline(yintercept=10000, color = "red") + #Set maximum number of genes to be expressed per cell
  geom_vline(xintercept=500, color = "blue") + #Set minimum number of UMI's per cell
  geom_vline(xintercept=50000, color = "blue") #Set maximum number of UMI's per cell

#Below two plots should match the above values and colors
BB_count_hist <- BB@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 500, color = "blue") + 
  geom_vline(xintercept = 50000, color = "blue") +
  scale_x_log10()

BB_feature_hist <- BB@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 500, color = "red") +
  geom_vline(xintercept = 10000, color = "red") +
  scale_x_log10() +
  coord_flip()

wrap_plots(
  BB_count_hist, blank,
  BB_scatter, BB_feature_hist,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.2), 
  heights = c(0.2, 1)
)

ggsave("/scratch/ac05869/plob_sc/seurat_qc/BB_cutoffs.png", height = 8, width = 8, bg = "white")

#Same as above but use the organellar filtered dataset so I can do a comparison

BB_scatter <- BB_chloro_filt@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point() + ggtitle("PLU_BB pipseq - Organelle Filtering First") +
  scale_x_continuous(trans='log10') +
  scale_y_continuous(trans='log10') +
  geom_hline(yintercept=500, color = "red") +
  geom_hline(yintercept=10000, color = "red") +
  geom_vline(xintercept=500, color = "blue") +
  geom_vline(xintercept=50000, color = "blue")

BB_count_hist <- BB_chloro_filt@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 500, color = "blue") +
  geom_vline(xintercept = 50000, color = "blue") +
  scale_x_log10()

BB_feature_hist <- BB_chloro_filt@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 500, color = "red") +
  geom_vline(xintercept = 10000, color = "red") +
  scale_x_log10() +
  coord_flip()

wrap_plots(
  BB_count_hist, blank,
  BB_scatter, BB_feature_hist,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.2), 
  heights = c(0.2, 1)
)

ggsave("/scratch/ac05869/plob_sc/seurat_qc/BB_chloro_filt_cutoffs.png", height = 8, width = 8, bg = "white")
```

###rep 2
```{r}
#You should generate a plot with test values and adjust based on the dataset!
#I plot the full dataset and the organellar filtered dataset to be able to visualize how the organellar filtering impacts the data

BC_scatter <- BC@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point() + ggtitle("PLU_BC pipseq") +
  scale_x_continuous(trans='log10') +
  scale_y_continuous(trans='log10') +
  geom_hline(yintercept=500, color = "red") + #Set minimum number of genes to be expressed per cell
  geom_hline(yintercept=10000, color = "red") + #Set maximum number of genes to be expressed per cell
  geom_vline(xintercept=500, color = "blue") + #Set minimum number of UMI's per cell
  geom_vline(xintercept=50000, color = "blue") #Set maximum number of UMI's per cell

#Below two plots should match the above values and colors
BC_count_hist <- BC@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 500, color = "blue") + 
  geom_vline(xintercept = 50000, color = "blue") +
  scale_x_log10()

BC_feature_hist <- BC@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 500, color = "red") +
  geom_vline(xintercept = 10000, color = "red") +
  scale_x_log10() +
  coord_flip()

wrap_plots(
  BC_count_hist, blank,
  BC_scatter, BC_feature_hist,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.2), 
  heights = c(0.2, 1)
)

ggsave("/scratch/ac05869/plob_sc/seurat_qc/BC_cutoffs.png", height = 8, width = 8, bg = "white")

#Same as above but use the organellar filtered dataset so I can do a comparison

BC_scatter <- BC_chloro_filt@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point() + ggtitle("PLU_BC pipseq - Organelle Filtering First") +
  scale_x_continuous(trans='log10') +
  scale_y_continuous(trans='log10') +
  geom_hline(yintercept=500, color = "red") +
  geom_hline(yintercept=10000, color = "red") +
  geom_vline(xintercept=500, color = "blue") +
  geom_vline(xintercept=50000, color = "blue")

BC_count_hist <- BC_chloro_filt@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 500, color = "blue") +
  geom_vline(xintercept = 50000, color = "blue") +
  scale_x_log10()

BC_feature_hist <- BC_chloro_filt@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 500, color = "red") +
  geom_vline(xintercept = 10000, color = "red") +
  scale_x_log10() +
  coord_flip()

wrap_plots(
  BC_count_hist, blank,
  BC_scatter, BC_feature_hist,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.2), 
  heights = c(0.2, 1)
)

ggsave("/scratch/ac05869/plob_sc/seurat_qc/BC_chloro_filt_cutoffs.png", height = 8, width = 8, bg = "white")
```


# Data Set Filtering
```{r}
#Based on above plots filter dataset appropriately 

BB#74420 features across 14858 samples
BB <- subset(BB, subset = nFeature_RNA > 500 & nFeature_RNA < 10000 & nCount_RNA > 500 & nCount_RNA < 50000 & percent.organellar < 5)
BB#74420 features across 13357 samples within 1 assay 

##Remove genes that have organellar ID's
countsBB <- GetAssayData(BB, assay = "RNA")
keepBB <- grepl("^Pluob\\.v4", rownames(countsBB))
counts_filteredBB <- countsBB[keepBB, ]
head(counts_filteredBB)
BB<- subset(BB, features = rownames(counts_filteredBB))
BB
#After Filtering
#74377 features across 13357 samples
#43 genes removed

BC #71900 features across 13850 samples
BC <- subset(BC, subset = nFeature_RNA > 500 & nFeature_RNA < 10000 & nCount_RNA > 500 & nCount_RNA < 50000 & percent.organellar < 5)
BC #71900 features across 11829 samples

##Remove genes that have organellar ID's
countsBC <- GetAssayData(BC, assay = "RNA")
keepBC <- grepl("^Pluob\\.v4", rownames(countsBC))
counts_filteredBC <- countsBC[keepBC, ]
head(counts_filteredBC)
BC<- subset(BC, features = rownames(counts_filteredBC))
BC
#After Filtering
#71867 features across 11829 samples
#33 genes filtered out
```



## Summary Statistics
###rep1
```{r}
summary(BB$nFeature_RNA)
summary(BB$nCount_RNA)

summary(BC$nFeature_RNA)
summary(BC$nCount_RNA)

```
    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    501     783    1361    2092    2630    9979 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    598     913    1664    3225    3638   28944 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    501     754    1229    1774    2186    9968 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    528     851    1438    2469    2807   24709 
# Data Processing - QC First Look
>Notes

We want to process the data as a quick QC to check clustering
This preprocessing is also required to do Doublet Filtering 

## Normalize the Data

```{r}
#Default = LogNormalize
BC <- NormalizeData(BC)
BB <- NormalizeData(BB)
```

## Clustering

```{r}
#Select Highly Variable Features
BC <- FindVariableFeatures(BC, selection.method = "vst", nfeatures = 3000)
BB <- FindVariableFeatures(BB, selection.method = "vst", nfeatures = 3000)
```

## Scale the Data

```{r}
BC.genes <- rownames(BC)
BC <- ScaleData(BC, features = BC.genes)

BB.genes <- rownames(BB)
BB <- ScaleData(BB, features = BB.genes)
```

## PCA

```{r}
BC$Run <- BC@meta.data$orig.ident
BC <- RunPCA(BC, npcs = 75, features = VariableFeatures(object = BC), verbose = FALSE) # run PCA (principal component analysis)

BB$Run <- BB@meta.data$orig.ident
BB <- RunPCA(BB, npcs = 75, features = VariableFeatures(object = BB), verbose = FALSE) # run PCA
```

```{r}
#Elbow plot helps us quickly estimate/check for how many PC's to use for clustering
ElbowPlot(BC, ndims = 75)

ElbowPlot(BB, ndims = 75)

```

## Call Clusters

```{r}
#Adjust dimensions based on PC's in elbow plot; I chose 50 because it is where they level off at low standard deviation
#Adjust resolution based on clustering. Optimal resolution typically increases for datasets with more cells because you are more likely to capture more types of cells
##Usually need to look at the UMAP and decide if cluster numbers/splitting looks good or not. Stem will usually have more clusters than leaf
BC <- FindNeighbors(BC, dims = 1:50)
BC <- FindClusters(BC, resolution = 0.5) #resolution can be adjusted depending on how the cells cluster. Higher resolution will break the umap into more clusters
BC <- RunUMAP(BC, dims = 1:50)

BB <- FindNeighbors(BB, dims = 1:50)
BB <- FindClusters(BB, resolution = 0.5)
BB <- RunUMAP(BB, dims = 1:50)
```

## Make UMAP Plots

```{r}
UMAP1 <- DimPlot(BC, reduction = "umap", label = T, label.size = 5, repel = T, split.by = "Run") +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none",
    title = element_text(size = 10)
  ) +
  ggtitle("Grouped by Replicates\n")
UMAP1


UMAP2 <- DimPlot(BB, reduction = "umap", label = T, label.size = 5, repel = T, split.by = "Run") +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none",
    title = element_text(size = 10)
  ) +
  ggtitle("Grouped by Replicates\n")
UMAP2
#I usally dont save this UMAP since I'm usually integrating later anyway. They look like they clustered well, so I don't need to adjust the resolution from 0.5
```

## General Stats

```{r}
#Cells in each cluster: take note
table(Idents(BC))
# Cells in each replicate: take note
table(BC$orig.ident)
#How does cluster membership vary by replicate? #clusters
table(Idents(BC), BC$orig.ident)

#Cells in each cluster:
table(Idents(BB))
# Cells in each replicate: 
table(BB$orig.ident)
#How does cluster membership vary by replicate? 
table(Idents(BB), BB$orig.ident)
```

## Save "dirty" RDS before doublet removal in case needed later

```{r}
saveRDS(BC, file = "/scratch/ac05869/plob_sc/seurat_data/PLU_BC_predoublet.rds")

saveRDS(BB, file = "/scratch/ac05869/plob_sc/seurat_data/PLU_BB_predoublet.rds")

#BC <-readRDS(file = "/scratch/ac05869/plob_sc/seurat_data/PLU_BC_predoublet.rds")
#BB <-readRDS(file = "/scratch/ac05869/plob_sc/seurat_data/PLU_BB_predoublet.rds")
```

## Stat Check

```{r}
#Check various stats to determine if there is an outlier for any of the clusters
#Are there any clusters which have more doublets than others?
plot1 <- VlnPlot(BC, features = c("nFeature_RNA"), pt.size = 0.5) + ggtitle("PLU_BC - nFeature_RNA")
plot2 <-VlnPlot(BC, features = c("nCount_RNA"), pt.size = 0.5) + ggtitle("PLU_BC - nCount_RNA")

ggarrange(plot1, plot2, nrow = 2, ncol = 1, common.legend = TRUE, legend = "bottom")

plot1
plot2

plot3 <- VlnPlot(BB, features = c("nFeature_RNA"), pt.size = 0.5) + ggtitle("PLU_BB - nFeature_RNA")
plot4 <-VlnPlot(BB, features = c("nCount_RNA"), pt.size = 0.5) + ggtitle("PLU_BB - nCount_RNA")

ggarrange(plot3, plot4, nrow = 2, ncol = 1, common.legend = TRUE, legend = "bottom")

plot3
plot4

#Notes: BC and BB; nFeature is number of genes/cell, nCount is number of UMIs/cell
#BC:  13 clusters
#BB: 14 clusters
```

# Doublet Filtering
## Run Doubletfinder
```{r}
library(DoubletFinder)
#Code needs to be edited to work currently EVERY TIME in v4 but not v5
#for V4, use "trace(paramSweep, edit = T)" and "trace(doubletFinder, edit = T)" - and edit $counts to @counts
```


#BB doubles
```{r}
#Seurat Objects Pre-Processed, PCA and UMAP run 

## pK Identification (no ground-truth) ---------------------------------------------------------------------------------------
sweep.res.list_ai <- paramSweep(BB, PCs = 1:50, sct = FALSE)
sweep.stats_ai <- summarizeSweep(sweep.res.list_ai, GT = FALSE)
bcmvn_ai <- find.pK(sweep.stats_ai)

pK=as.numeric(as.character(bcmvn_ai$pK))
BCmetric=bcmvn_ai$BCmetric
pK_choose = pK[which(BCmetric %in% max(BCmetric))]

## ggplot doesn't work
showtext_auto(enable = TRUE)
png(file = "/scratch/ac05869/plob_sc/seurat_qc/PLU_BB_doubletfinder_pK_identification.png")
par(mar=c(5,4,4,8)+1,cex.main=1.2,font.main=2)
plot(x = pK, y = BCmetric, pch = 16,type="b",
col = "blue",lty=1)
abline(v=pK_choose,lwd=2,col='red',lty=2)
title("The BCmvn distributions")
text(pK_choose,max(BCmetric),as.character(pK_choose),pos = 4,col = "red")
dev.off()
showtext_auto(enable = F)


## Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
homotypic.prop <- modelHomotypic(BB@meta.data$seurat_clusters)           ## ex: annotations <- seu_kidney@meta.data$ClusteringResults
nExp_poi <- round(0.07*nrow(BB@meta.data))  ## Assuming 7% doublet formation rate - updated based on 40k cells loaded for PIPseq ##This is something that should be changed based on assumed doublet levels!
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
```

## BB doublet test
```{r}
## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
#Update pK to be pK choosen from Above steps!
BB_doublet_test <-doubletFinder(BB, #change library ID
                                PCs = 1:50, #Josh doesn't usually change this to match the number of PCs
                                pN = 0.25, 
                                pK = pK_choose, #the only number we change from earlier code (I use the variable itself)
                                nExp = nExp_poi, #this variable is from previous code code chunk
                                reuse.pANN = FALSE, sct = FALSE)

BB_doublet_test

value_BB <- paste0("DF.classifications_0.25_", pK_choose, "_", nExp_poi) #pK_choose is the variable or pK value chosen from above steps, the last value is the # doublets found
value_BB

sum(str_count(BB_doublet_test[[value_BB]], "Doublet")) #How many are doublets? 
DimPlot(BB_doublet_test,pt.size = 1,label=TRUE, label.size = 5,reduction = "umap",group.by = value_BB)+theme(aspect.ratio = 1)
#Save a UMAP of which cells were the doublets
##Does this match your expectations from the Violin plots of the clusters we made earlier?

ggsave("/scratch/ac05869/plob_sc/seurat_qc/PLU_BB_doubletfinder_UMAP.png", height = 16, width = 16, bg = "white")
```
#BC doubles
```{r}
#Seurat Objects Pre-Processed, PCA and UMAP run 

## pK Identification (no ground-truth) ---------------------------------------------------------------------------------------
sweep.res.list_ai <- paramSweep(BC, PCs = 1:50, sct = FALSE)
sweep.stats_ai <- summarizeSweep(sweep.res.list_ai, GT = FALSE)
bcmvn_ai <- find.pK(sweep.stats_ai)

pK=as.numeric(as.character(bcmvn_ai$pK))
BCmetric=bcmvn_ai$BCmetric
pK_choose = pK[which(BCmetric %in% max(BCmetric))]

## ggplot doesn't work
showtext_auto(enable = TRUE)
png(file = "/scratch/ac05869/plob_sc/seurat_qc/PLU_BC_doubletfinder_pK_identification.png")
par(mar=c(5,4,4,8)+1,cex.main=1.2,font.main=2)
plot(x = pK, y = BCmetric, pch = 16,type="b",
col = "blue",lty=1)
abline(v=pK_choose,lwd=2,col='red',lty=2)
title("The BCmvn distributions")
text(pK_choose,max(BCmetric),as.character(pK_choose),pos = 4,col = "red")
dev.off()
showtext_auto(enable = F)


## Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
homotypic.prop <- modelHomotypic(BC@meta.data$seurat_clusters)           ## ex: annotations <- seu_kidney@meta.data$ClusteringResults
nExp_poi <- round(0.07*nrow(BC@meta.data))  ## Assuming 7% doublet formation rate - updated based on 40k cells loaded for PIPseq ##This is something that should be changed based on assumed doublet levels!
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
```

## BC doublet test
```{r}
## Run DoubletFinder with varying classification stringencies ----------------------------------------------------------------
#Update pK to be pK choosen from Above steps!
BC_doublet_test <-doubletFinder(BC, #change library ID
                                PCs = 1:50, #Josh doesn't usually change this to match the number of PCs
                                pN = 0.25, 
                                pK = pK_choose, #the only number we change from earlier code (I use the variable itself)
                                nExp = nExp_poi, #this variable is from previous code code chunk
                                reuse.pANN = FALSE, sct = FALSE)

BC_doublet_test

value_BC <- paste0("DF.classifications_0.25_", pK_choose, "_", nExp_poi) #pK_choose is the variable or pK value chosen from above steps, the last value is the # doublets found
value_BC

sum(str_count(BC_doublet_test[[value_BC]], "Doublet")) #How many are doublets? 
DimPlot(BC_doublet_test,pt.size = 1,label=TRUE, label.size = 5,reduction = "umap",group.by = value_BC)+theme(aspect.ratio = 1)
#Save a UMAP of which cells were the doublets
##Does this match your expectations from the Violin plots of the clusters we made earlier?

ggsave("/scratch/ac05869/plob_sc/seurat_qc/PLU_BC_doubletfinder_UMAP.png", height = 16, width = 16, bg = "white")
```
## Save RDS Object from Doublet Finder

```{r}
saveRDS(BC_doublet_test, file = "/scratch/ac05869/plob_sc/seurat_data/PLU_BC_doublet_finder_seurat.rds")

saveRDS(BB_doublet_test, file = "/scratch/ac05869/plob_sc/seurat_data/PLU_BB_doublet_finder_seurat.rds")

#Notes:

#BC_doublet_test <- ("/scratch/ac05869/plob_sc/seurat_data/PLU_BC_doublet_finder_seurat.rds")
#BB_doublet_test <- ("/scratch/ac05869/plob_sc/seurat_data/PLU_BB_doublet_finder_seurat.rds")

```

## Filter Doublets
```{r}
#BC
value_BC
BC_clean <- subset(BC_doublet_test, subset = DF.classifications_0.25_0.005_828 == "Singlet") #BC "DF.classifications_0.25_0.005_828"
BC_clean # 

#BB
value_BB
BB_clean <- subset(BB_doublet_test, subset = DF.classifications_0.25_0.3_935 == "Singlet") #BB "DF.classifications_0.25_0.3_935"
BB_clean # 
#take note of number of doublets that were filtered out. 
```

## Summary Statistics Cleaned Data: Number of Genes  
```{r}
summary(BC_clean$nFeature_RNA)
summary(BC_clean$nCount_RNA)

summary(BB_clean$nFeature_RNA)
summary(BB_clean$nCount_RNA)
```
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    482     726    1134    1556    1939    9931 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    521     812    1311    2064    2410   24655 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    483     753    1244    1717    2256    9329 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    573     868    1489    2406    2978   24958 
# Re-Clustering Cleaned Data 
> Notes

Now that we have a clean dataset we need to re-cluster things to account for the removal of doublets 
We can either re-run the above clustering steps for single samples OR we can integrate our replicates together at the same time


## Normalize the Data
```{r}
#Default = LogNormalize
BC_clean <- NormalizeData(BC_clean)
BB_clean <- NormalizeData(BB_clean)

## Save clean data - want to eventually integrate with leaf clean normalized data for integrated dataset

saveRDS(BC_clean, file = "/scratch/ac05869/plob_sc/seurat_data/PLU_BC_clean_normalized.rds")

saveRDS(BB_clean, file = "/scratch/ac05869/plob_sc/seurat_data/PLU_BB_clean_normalized.rds")

#Notes:

```


# Integration via reciprocal PCA (RPCA)
>Notes

This method is what was suggested in Suerat v4 as of Oct 2023. In their new vignette for Seurat v5 the methodology looks to be updated. This may warrant investigation prior to running the below! I still recommend RPCA, see the linked vginette below for reasons why.

https://satijalab.org/seurat/articles/integration_rpca
We therefore,recommend RPCA during integrative analysis where: * A substantial fraction of cells in one dataset have no matching type in the other

##Data Prep
```{r}
BC_clean$sample <- "Leaf"
BB_clean$sample <- "Leaf"
leaf.list <- list(BC_clean, BB_clean)

# normalize and identify variable features for each dataset independently
leaf.list <- lapply(X = leaf.list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 3000)
})

# select features that are repeatedly variable across datasets for integration run PCA on each
# dataset using these features
features <- SelectIntegrationFeatures(object.list = leaf.list, nfeatures = 3000)
leaf.list <- lapply(X = leaf.list, FUN = function(x) {
    x <- ScaleData(x, features = features, verbose = FALSE)
    x <- RunPCA(x, features = features, verbose = FALSE)
})
```

## Integration 
```{r}
leaf.anchors <- FindIntegrationAnchors(object.list = leaf.list, anchor.features = features, reduction = "rpca")

# this command creates an 'integrated' data assay
leaf.combined <- IntegrateData(anchorset = leaf.anchors)

# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(leaf.combined) <- "integrated"
```

## Stats
```{r}
leaf.combined

summary(leaf.combined$nFeature_RNA)
summary(leaf.combined$nCount_RNA)
```
An object of class Seurat 
79566 features across 23423 samples within 2 assays 
Active assay: integrated (3000 features, 3000 variable features)
 1 layer present: data
 1 other assay present: RNA
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    482     740    1187    1642    2095    9931 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
    521     841    1399    2245    2677   24958 
## Scale the Data
```{r}
leaf.combined.genes <- rownames(leaf.combined)
leaf.combined <- ScaleData(leaf.combined, features = leaf.combined.genes)
```

## PCA
```{r}
leaf.combined$Run <- leaf.combined@meta.data$orig.ident
leaf.combined <- RunPCA(leaf.combined, npcs = 75, features = VariableFeatures(object = leaf.combined), verbose = FALSE) # run PCA
```

```{r}
ElbowPlot(leaf.combined, ndims = 75)
```

## Call Clusters
```{r}
leaf.combined <- FindNeighbors(leaf.combined, dims = 1:50) #adjust dims based on elbow plot above
leaf.combined <- FindClusters(leaf.combined, resolution = 0.5) #adjustable resolution value. Check back and forth until it looks good
leaf.combined <- RunUMAP(leaf.combined, dims = 1:50)
```

## Make UMAP Plots
```{r}
UMAP1 <- DimPlot(leaf.combined, reduction = "umap", repel = T, split.by = "Run") +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none",
    title = element_text(size = 10)
  ) +
  ggtitle("Grouped by Replicates\n")
UMAP1
#check to see that there are the same number of cluster in each rep
ggsave("/scratch/ac05869/plob_sc/seurat_qc/PLU_BB_BC_integrated_umap.png", height = 8, width = 12, bg = "white")
ggsave("/scratch/ac05869/plob_sc/seurat_qc/PLU_BB_BC_integrated_umap.svg", height = 8, width = 12, bg = "white")

ggsave("/scratch/ac05869/plob_sc/seurat_qc/PLU_BB_BC_integrated_umap_small.png", height = 4, width = 10, bg = "white")
ggsave("/scratch/ac05869/plob_sc/seurat_qc/PLU_BB_BC_integrated_umap_small.svg", height = 4, width = 10, bg = "white")
```

```{r}
UMAP2 <- DimPlot(leaf.combined, reduction = "umap", label = T,
                 label.size = 5, repel = T) +
  theme_void() +
  theme(
    text = element_text(size = 18, color = "black", face = "bold"),
    legend.position = "none",
    title = element_text(size = 10)
  )
UMAP2
ggsave("/scratch/ac05869/plob_sc/seurat_qc/PLU_BB_BC_integrated_umap_all_together.png", height = 8, width = 12, bg = "white")
ggsave("/scratch/ac05869/plob_sc/seurat_qc/PLU_BB_BC_integrated_umap_all_together.svg", height = 8, width = 12, bg = "white")

ggsave("/scratch/ac05869/plob_sc/seurat_qc/PLU_BB_BC_integrated_umap_all_together_small.png", height = 4, width = 6, bg = "white")
ggsave("/scratch/ac05869/plob_sc/seurat_qc/PLU_BB_BC_integrated_umap_all_together_small.svg", height = 4, width = 6, bg = "white")

```

## Stats - Integrated Clusters
```{r}
#Cells in each cluster
table(Idents(leaf.combined))

# Cells in each replicate
table(leaf.combined$orig.ident)

#How does cluster membership vary by replicate?
table(Idents(leaf.combined), leaf.combined$orig.ident)
```

# Save cleaned integrated RDS Object
```{r}
DefaultAssay(leaf.combined) <- "RNA"
leaf.combined = JoinLayers(leaf.combined)


#due to the error messages, I ran this and the FindAllMarkers worked. Before doing this, this chunk generated error messages when I ran it: "Warning: No DE genes identifiedWarning: The following tests were not performed: Warning: When testing 0 versus all: data layers are not joined. Please run JoinLayers"
saveRDS(leaf.combined, file = "/scratch/ac05869/plob_sc/seurat_data/PLU_leaf_cleaned_integrated.rds")
leaf.combined <- readRDS("/scratch/ac05869/plob_sc/seurat_data/PLU_leaf_cleaned_integrated.rds")
```

# Create Matrices
##Annotation data - both nuclear mitochondrial and chloroplast
```{r}
#repr_gene_models <- read_delim("/scratch/ac05869/plob_sc/STAR_index/plob_v4.repr_working.mit.chloro.agat.gtf",
#                               delim = "\t", col_names = F)
#anno <- read_delim("/scratch/ac05869/plu_comparative_final/genomes/plob_v4/plob_v4.functional_annotation.txt", col_names = c("gene_ID", "gene_anno"))

#repr_models <- repr_gene_models %>% 
#  dplyr::filter(X3 == "gene") %>% 
#  dplyr::select(X1, X4, X5, X7, X9) %>% 
#  dplyr::rename(
#    Chr = X1,
#    start = X4,
#    end = X5,
#    strand = X7
#  ) %>%
#  separate(X9, c("temp1", "ID", "gene_ID"), sep = ";") %>%
#  mutate(across(everything(), ~ str_trim(str_replace_all(.x, '"', ''))))%>%
#  separate(gene_ID, c("temp", "gene_ID"), sep = " ") %>%
#  dplyr::select(c(Chr, start, end, strand, gene_ID, ID)) %>%
#  left_join(anno, by = "gene_ID") %>%
#  select(-gene_ID)%>%
#  separate(ID, c("temp", "gene_ID"), sep = " ")%>%
#  select(-temp)

#head(repr_models)
#dim(repr_models) #130042

#write_delim(repr_models, "/scratch/ac05869/plob_sc/seurat_data/repr_gene_models.csv", delim = ",")
repr_models <- read_delim("/scratch/ac05869/plob_sc/seurat_data/repr_gene_models.csv", delim = ",")
head(repr_models)
dim(repr_models) #130042
```
# plot pathway genes
```{r}
#Need to create a file that has your marker genes and some metadata (cell types, etc.). Change all underscores to either dashes or periods
path <- read_delim("/scratch/ac05869/plu_comparative_final/pathway/plumeria_iridoid_pathway_synthelogs.txt", delim = "\t")
head(path)
path <- path %>%
  separate_rows(P.obtusa_A, P.obtusa_B, sep = ",") %>%
  mutate(P.obtusa_A = trimws(P.obtusa_A))%>%
  mutate(P.obtusa_B = trimws(P.obtusa_B)) %>%
  filter(
    !is.na(P.obtusa_A), P.obtusa_A != "",
    !is.na(P.obtusa_B), P.obtusa_B != ""
  )

pathA <- path %>%
  mutate(gene_ID = P.obtusa_A)%>%
  select(-c(P.obtusa_A, P.obtusa_B)) %>%
  distinct()
pathB <- path %>%
  mutate(gene_ID = P.obtusa_B)%>%
  select(-c(P.obtusa_A, P.obtusa_B))%>%
  distinct()

path <- rbind(pathA, pathB) %>%
  mutate(subgenome = case_when(
    str_detect(gene_ID, "aG") ~ "A",
    str_detect(gene_ID, "bG") ~"B")
  ) %>%
  arrange(order, subgenome)

path$gene_ID <- gsub("\\.\\d+$", "", path$gene_ID)

rev_path <- path[order(nrow(path):1),]
head(rev_path)

write_delim(path, "/work/crblab/ac05869/plob_v4_single_cell/pathway_analysis/plumeria_MEP_iridoid_syntelogs.tsv", delim = "\t")
```


```{r}
path <- read_delim("/work/crblab/ac05869/plob_v4_single_cell/pathway_analysis/plumeria_MEP_iridoid_syntelogs.tsv", delim = "\t")
rev_path <- path[order(nrow(path):1),]
```

## Find Markers 

> Notes

seurat told me to install presto package for "faster implementation of the Wilcoxon Rank Sum Test (default method for FindMarkers)"
devtools::install_github('immunogenomics/presto')
```{r}
DefaultAssay(leaf.combined) <- "RNA"

leaf.combined.markers <- FindAllMarkers(leaf.combined, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25) 
#min.pct is the minimum fraction that cells cluster needs to be expressing that gene for it to count. logfc.threshhold is the log-fold change in expression between clusters expressing that gene
head(leaf.combined.markers)

leaf.combined.markers_annotated <- leaf.combined.markers %>%
  left_join(repr_models, by = c("gene" = "gene_ID"), relationship = "many-to-many") %>% 
  left_join(path %>% select(gene_ID, tag, order), by = c("gene" = "gene_ID")) %>%
  group_by(cluster)

head(leaf.combined.markers_annotated)

write.table(leaf.combined.markers, "/scratch/ac05869/plob_sc/seurat_data/PLU_leaf_markers_minpct0.25_minlfc0.25.txt", quote = FALSE, sep = "\t", row.names = FALSE)
write_delim(leaf.combined.markers_annotated, "/scratch/ac05869/plob_sc/seurat_data/PLU_leaf_markers_annotated_minpct0.25_minlfc0.25.tsv", delim = "\t")
leaf.combined.markers_annotated <- read_delim("/scratch/ac05869/plob_sc/seurat_data/PLU_leaf_markers_annotated_minpct0.25_minlfc0.25.tsv", delim = "\t")
```

## Find Average Expression
```{r}
DefaultAssay(leaf.combined) <- "RNA"

# replace spaces with underscores '_' so ggplot2 doesn't fail
orig.levels <- levels(leaf.combined)
Idents(leaf.combined) <- gsub(pattern = " ", replacement = "_", x = Idents(leaf.combined))
orig.levels <- gsub(pattern = " ", replacement = "_", x = orig.levels)
levels(leaf.combined) <- orig.levels
cluster.averages <- AverageExpression(leaf.combined, return.seurat = TRUE)

# How can I extract expression matrix for all Cluster 0 cells (perhaps, to load into another package)
cluster_avg.data <- as.matrix(GetAssayData(cluster.averages, layer = "data")[, ])
#Warning: The `slot` argument of `GetAssayData()` is deprecated as of SeuratObject 5.0.0.
#Please use the `layer` argument instead.

write.table(cluster_avg.data, file = "/scratch/ac05869/plob_sc/seurat_data/PLU_leaf_cluster_av.norm.counts.txt", sep = "\t", quote = FALSE)
head(cluster_avg.data)
```
# QC Data via Marker Gene Expression Check 

```{r}
DefaultAssay(leaf.combined) <- "RNA"
rev_pathway_genes <- rev_path
dotplot_data <- DotPlot(leaf.combined, features = unique(rev_pathway_genes$gene_ID))$data

dotplot_data %>% 
  inner_join(rev_pathway_genes %>% 
               distinct(gene_ID, .keep_all = T), by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    LocusID = reorder(LocusID, -order),
    ylab = paste0(tag, ":", LocusID),  # make this new variable
    ylab = reorder(ylab, -order)
) %>%
  ggplot(aes(x = id, y = ylab)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Cluster",
       title = "P.obtusa syntelogs - Leaf") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

#ggsave("/scratch/ac05869/plob_sc/pathway_analysis/PLU_leaf_pathway_expression_251110_1.png", height = 10, width = 8)
#ggsave("/scratch/ac05869/plob_sc/pathway_analysis/PLU_leaf_pathway_expression_251110_1.svg", height = 10, width = 8)

```
#Cell type annotations
Read in genespace run to find syntelogs of markers in Plumeria
```{r}
load(file = "/scratch/ac05869/plu_comparative_final/Genespace/Apocynaceae_with_outgroup/wd/results/pangenes_athaliana_ref.rda")

all_syntelogs <- pangenes %>%
  mutate(
    across(where(is.list),~ map_chr(.x, ~ paste(.x, collapse = ",")))) %>%
  separate_rows(A.thaliana, sep = ",") %>%
  mutate(A.thaliana = trimws(A.thaliana)) 

all_syntelogs$A.thaliana <- gsub("\\.\\d+$", "", all_syntelogs$A.thaliana)
all_syntelogs$C.roseus <- gsub("\\.\\d+$", "", all_syntelogs$C.roseus)

write_delim(all_syntelogs, file = "/scratch/ac05869/plob_sc/pathway_analysis/markers/Athaliana_syntelogs_pangenes.tsv", delim = "\t")

TAIR <- read_delim(file = "/scratch/ac05869/plob_sc/pathway_analysis/markers/TAIR10_functional_descriptions.txt", delim = "\t")

annotated_syntelogs <- all_syntelogs %>%
  left_join(TAIR %>% select(-Type), by = c("repGene"= "Model_name"))
View(annotated_syntelogs)
write_delim(annotated_syntelogs, file = "/scratch/ac05869/plob_sc/pathway_analysis/markers/Athaliana_syntelogs_pangenes_annotated.tsv", delim = "\t")

```
#Read in marker genes 
```{r}
#/scratch/ac05869/plob_sc/pathway_analysis/markers
pop_epi <- read_delim("/work/crblab/ac05869/plob_v4_single_cell/cluster_anno/EPDIERMIS_POP_shoot_epi_clean_anno_20251104.tsv", delim = "\t")
head(pop_epi)
markers <- read_delim("/scratch/ac05869/plob_sc/pathway_analysis/markers/Dec2025_compiled_stem_leaf_markers_annotated.tsv", delim = "\t")
head(markers)
colnames(markers)

markers2 <- markers %>% select(-DOI1, -DOI2, -DOI3, -DOI4, -Type, -Computational_description) %>%
  rename(c(gene_ID =Arabidopsis, cell_type = Cell_type))
colnames(markers2)

pop_epi2 <- pop_epi %>% select(Arabidopsis, gene_name, Best.Arabidopsis.BLAST.hit, cell_type ) %>%
  rename(c(gene_ID=Arabidopsis, Name = gene_name)) %>%
  left_join(TAIR %>% select(-Type), by = c("gene_ID"= "Model_name")) %>%
  select(c(-Best.Arabidopsis.BLAST.hit, -Computational_description))
pop_epi2$gene_ID <- gsub("\\.\\d+$", "", pop_epi2$gene_ID)
pop_epi2$Sources <- "Poplar markers"
pop_epi2$Organ <- "Leaf"
colnames(pop_epi2)
head(pop_epi2)

markers_comb <- markers2 %>%
  filter(!gene_ID %in% pop_epi2$gene_ID)%>%
  bind_rows(pop_epi2)
```

## Markers to syntelogs
```{r}
colnames(all_syntelogs)
syntenic_markers.1 <- all_syntelogs %>%
  filter(A.thaliana %in%  markers_comb$gene_ID)%>%
  rename(Marker_ID = A.thaliana) %>%
  filter(
    !is.na(P.obtusa_A), P.obtusa_A != "",
    !is.na(P.obtusa_B), P.obtusa_B != ""
  )%>%
  separate_rows(P.obtusa_A, sep = ",") %>%
  mutate(P.obtusa_A = trimws(P.obtusa_A))%>%
  separate_rows(P.obtusa_B, sep = ",") %>%
  mutate(P.obtusa_B = trimws(P.obtusa_B)) %>%
  select(-start, -end, -pgID, -interpChr, -interpOrd, -og, -repGene, -genome, -chr, -P.rubra_A, -P.rubra_B, -H.carnosa, -A.pictum, -C.roseus)


syntenic_markers.2 <- all_syntelogs %>%
  filter(C.roseus %in%  markers_comb$gene_ID)%>%
  rename(Marker_ID = C.roseus) %>%
  filter(
    !is.na(P.obtusa_A), P.obtusa_A != "",
    !is.na(P.obtusa_B), P.obtusa_B != ""
  )%>%
  separate_rows(P.obtusa_A, sep = ",") %>%
  mutate(P.obtusa_A = trimws(P.obtusa_A))%>%
  separate_rows(P.obtusa_B, sep = ",") %>%
  mutate(P.obtusa_B = trimws(P.obtusa_B)) %>%
  select(-start, -end, -pgID, -interpChr, -interpOrd, -og, -repGene, -genome, -chr, -P.rubra_A, -P.rubra_B, -H.carnosa, -A.pictum, -A.thaliana)

syntenic_markers <- rbind(syntenic_markers.1, syntenic_markers.2)

A_markers <- syntenic_markers %>%
  select(-P.obtusa_B) %>%
  mutate(gene_ID = P.obtusa_A) %>%
  select( -P.obtusa_A) %>%
  left_join(markers_comb, by = c("Marker_ID" ="gene_ID")) %>%
  distinct()

B_markers <- syntenic_markers %>%
  select(-P.obtusa_A)%>%
  mutate(gene_ID = P.obtusa_B)%>%
  select( -P.obtusa_B) %>%
  left_join(markers_comb, by = c("Marker_ID" ="gene_ID")) %>%
  distinct()

markers3 <- rbind(A_markers, B_markers) %>%
  group_by(gene_ID) %>%
  slice(1) %>%
  ungroup()%>%
  mutate(cell_type2 = case_when(
    str_detect(cell_type, "Abaxial epidermal") ~ "Abaxial epidermis",
        str_detect(cell_type, "Abaxial Epidermis") ~ "Abaxial epidermis",
    str_detect(cell_type, "abaxial") ~ "Abaxial epidermis",
    str_detect(cell_type, "Companion cells") ~ "Companion cell",
    T ~ cell_type
  ))%>%
  mutate(cell_type2 = case_when(
    str_detect(cell_type, "epidermis") ~ "Epidermis",
        str_detect(cell_type, "Epidermis") ~ "Epidermis",
        str_detect(cell_type, "XP") ~ "Xylem",
    str_detect(cell_type, "PP") ~ "Phloem",
        str_detect(cell_type, "Bundle") ~ "Vasculature",
        str_detect(cell_type, "Xylem") ~ "Xylem",
            str_detect(cell_type, "xylem") ~ "Xylem",
                str_detect(cell_type, "endodermis") ~ "Endodermis",
            str_detect(cell_type, "Adaxial Epidermis") ~ "Adaxial Epidermis",
    str_detect(cell_type, "adaxial") ~ "Adaxial Epidermis",
        str_detect(cell_type, "Sieve") ~ "Vasculature",
            str_detect(cell_type, "Mesophyll") ~ "Mesophyll",
            str_detect(cell_type, "Cortex") ~ "Cortex",
            str_detect(cell_type, "mesophyll") ~ "Mesophyll",
            str_detect(cell_type, "Vessel") ~ "Vasculature",
    T ~ cell_type
  ))%>%
  mutate(subgenome = case_when(
    str_detect(gene_ID, "aG") ~ "A",
    str_detect(gene_ID, "bG") ~"B")) %>%
  arrange(cell_type2, Name)%>%
#  filter(!(cell_type2 == "Epidermis" & is.na(Short_description) & Sources == "Lee et al 2025")) %>%
  distinct()
    
markers3$gene_ID <- gsub("\\.\\d+$", "", markers3$gene_ID) 
dim(markers3)#2309

write_tsv(markers3, file = '/scratch/ac05869/plob_sc/pathway_analysis/markers/March2026_compiled_stem_leaf_markers_annotated_plumeria_syntelogs.tsv')
#markers3 <- read_delim(file = '/scratch/ac05869/plob_sc/pathway_analysis/markers/Dec2025_compiled_stem_leaf_markers_annotated_plumeria_syntelogs.tsv')
```

```{r}
poplar_in_plu <- markers3 %>%
  filter(Sources == "Poplar markers")
```

#Seurat cluster markers with markers
```{r}
leaf.combined.markers_annotated <- read_delim("/scratch/ac05869/plob_sc/seurat_data/PLU_leaf_markers_annotated_minpct0.25_minlfc0.25.tsv", delim = "\t")

leaf_clust_anno <- leaf.combined.markers_annotated %>%
  select(-Chr, -start, -end, -strand, -tag, -order)%>%
  inner_join(markers3 , by = c('gene' = 'gene_ID')) %>%
  rename('gene_ID' = 'gene')

leaf_clust_anno_pop <- leaf.combined.markers_annotated %>%
  select(-Chr, -start, -end, -strand, -tag, -order)%>%
  inner_join(poplar_in_plu , by = c('gene' = 'gene_ID')) %>%
  rename('gene_ID' = 'gene')

write_delim(leaf_clust_anno, file = '/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/Plumeria_leaf_cluster_markers_annotated.tsv', delim = '\t')
write_delim(leaf_clust_anno_pop, file = '/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/Plumeria_leaf_cluster_markers_annotated_poplar.tsv', delim = '\t')

```

Annotated the clusters based on known markers and de novo markers from clustering, and pathway gene expression of MEP pathway (7DLH and G8H for IPAP cells)


```{r}
expressed_markers <- markers3 %>% filter(gene_ID %in% Features(leaf.combined))
dim(expressed_markers)#2363
dotplot <- DotPlot(leaf.combined, features = unique(expressed_markers$gene_ID))$data
dotplot
```

```{r}
celltypes <- dotplot %>% 
  inner_join(leaf_clust_anno %>%
               mutate(order = as.numeric(as.factor(cell_type2))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type,':', Name, "-", features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Leaf Cluster",
       title = "") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/seurat_marker_expression_20Mar26.svg", height = 24, width = 10)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/seurat_marker_expression_20Mar26.png", height = 24, width = 10)

```
```{r}
celltypes <- dotplot %>% 
  inner_join(markers3 %>%
    filter(!str_detect(Sources, "Lee et al 2025")) %>%
               mutate(order = as.numeric(as.factor(cell_type2))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type,':', Name, "-", features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Leaf Cluster",
       title = "") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/marker_cells_dotplot_no_Lee_20Mar26.svg", height = 20, width = 14)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/marker_cells_dotplot_no_Lee_20Mar26.png", height = 20, width = 14)

```
```{r}
celltypes <- dotplot %>% 
inner_join(markers3 %>%
    filter(str_detect(Sources, "Lee et al 2025")) %>%
               mutate(order = as.numeric(as.factor(cell_type2))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type,':', Name, "-", features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Leaf Cluster",
       title = "") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/marker_cells_dotplot_only_Lee_20Mar26.svg", height = 50, width = 14, limitsize = FALSE)
```
```{r}
library(patchwork)

# Create the base data first
celltypes_data <- dotplot %>% 
  inner_join(markers3 %>%
    filter(str_detect(Sources, "Lee et al 2025")) %>%
    mutate(order = as.numeric(as.factor(cell_type2))),
    by = c("features.plot" = "gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),
    ylab2 = paste0(cell_type, ':', Name, "-", features.plot),
    ylab2 = reorder(ylab2, -order)
  )

all_genes <- levels(celltypes_data$ylab2)
n <- length(all_genes)

e1_genes <- all_genes[1:floor(n/8)]
e2_genes <- all_genes[(floor(n/8)+1):floor(2*n/8)]
e3_genes <- all_genes[(floor(2*n/8)+1):floor(3*n/8)]
e4_genes <- all_genes[(floor(3*n/8)+1):floor(4*n/8)]
e5_genes <- all_genes[(floor(4*n/8)+1):floor(5*n/8)]
e6_genes <- all_genes[(floor(5*n/8)+1):floor(6*n/8)]
e7_genes <- all_genes[(floor(6*n/8)+1):floor(7*n/8)]
e8_genes <- all_genes[(floor(7*n/8)+1):n]

p1 <- make_panel(celltypes_data, e1_genes, show_legend = FALSE)
p2 <- make_panel(celltypes_data, e2_genes, show_legend = FALSE)
p3 <- make_panel(celltypes_data, e3_genes, show_legend = FALSE)
p4 <- make_panel(celltypes_data, e4_genes, show_legend = FALSE)
p5 <- make_panel(celltypes_data, e5_genes, show_legend = FALSE)
p6 <- make_panel(celltypes_data, e6_genes, show_legend = FALSE)
p7 <- make_panel(celltypes_data, e7_genes, show_legend = FALSE)
p8 <- make_panel(celltypes_data, e8_genes, show_legend = FALSE)

# 2x2 layout
final_plot1 <- p1 + p2
final_plot2 <- p3+p4
final_plot3<- p5+p5
final_plot4<- p7+p8

ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/marker_cells_dotplot_only_Lee_20Mar26_1.png",
       final_plot1, height = 30, width = 28, limitsize = FALSE)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/marker_cells_dotplot_only_Lee_20Mar26_2.png",
       final_plot2, height = 30, width = 28, limitsize = FALSE)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/marker_cells_dotplot_only_Lee_20Mar26_3.png",
       final_plot3, height = 30, width = 28, limitsize = FALSE)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/marker_cells_dotplot_only_Lee_20Mar26_4.png",
       final_plot4, height = 30, width = 28, limitsize = FALSE)
```
#Plot curated markers to annotate clusters
```{r}
leaf_markers <- read_delim(file = "/work/crblab/ac05869/plob_v4_single_cell/cluster_anno/plumeria_leaf_marker_list.tsv", delim = "\t")
dotplot <- DotPlot(leaf.combined, features = unique(leaf_markers$gene_ID))$data
dotplot
celltypes <- dotplot %>% 
  inner_join(leaf_markers %>%
               mutate(order = as.numeric(as.factor(cell_type))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", Name),  # make this new variable
    ylab2 = paste0(cell_type2,':', Name, "-", features.plot),  # make this new variable
    ylab3 = reorder(ylab, -order)
) %>%
  ggplot(aes(x = id, y = ylab3)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Leaf Cluster",
       title = "") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/Selected_markers_23March26.png", height = 10, width = 12)
```

Annotate clusters
#Reorder the columns by cell types
```{r}
Marker_dotplot <- DotPlot(leaf.combined, 
                          features = unique(leaf_markers$gene_ID))
unique(leaf_markers$Assigned_cell_type)

Marker_dotplot_nice_df <- Marker_dotplot$data %>% 
  mutate(gene_ID = row.names(.)) %>% 
  full_join(leaf_markers %>% 
              distinct(gene_ID, .keep_all = T),
             by = join_by(features.plot == gene_ID)) %>% 
    mutate(gene_ID = str_sub(gene_ID, end = -8)) %>% 
  mutate(cell_type3 = factor(Assigned_cell_type, levels = c(
    "Mesophyll", "Epidermis", "Protodermal", "Proliferating", "Vessels", "Sieve", "Guard cell", "IPAP"
  ))) %>% 
  mutate(cell_type4 = as.numeric(cell_type3)) %>% 
  mutate(gene_ID = reorder(gene_ID, -cell_type4))  %>% 
  mutate(LocusID = str_remove(
               str_sub(features.plot, end = -8),
               "^Pluob\\.v4\\."),
    ylab = paste0(features.plot),
    ylab2 = paste0(cell_type2,': ', Name),  # make this new variable
    ylab2 = reorder(ylab2, -cell_type4)) %>%
  mutate(id = factor(id, levels = c(
    "0", "2", "6", "8", "11", #Mesophyll
    "9","13", #Epidermis
    "4","12", #Protodermal
    "5", "7", "10", #Proliferating
    "1", "14", #Vessels
    "3", #Vasculature Sieve
    "15", #Guard cells
    "16" #IPAP
  ))) %>% 
  dplyr::filter(is.na(avg.exp.scaled) == F)


```

0	Mesophyll
1	Vessels
2	Mesophyll
3	Sieve
4	Protodermal
5	Proliferating
6	Mesophyll
7	Proliferating
8	Mesophyll
9	Epidermis
10	Proliferating
11	Mesophyll
12	Protodermal
13	Epidermis
14	Vessels
15	Guard cell
16	IPAP

```{r}
Marker_dot <- Marker_dotplot_nice_df %>% ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_fill_viridis(option = "A", begin = 0, end = 0.9) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"), nrow = 2)) +
  labs(y = "",
       x = "Leaf Clusters",
       fill = "Avg. Exp.",
       size = "% Exp.") +
  scale_y_discrete(labels = setNames(Marker_dotplot_nice_df$ylab2, Marker_dotplot_nice_df$Marker_ID)) + #replace the actual y axis with a new label
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "top",
        legend.box = "horizontal",
        legend.key.height = unit(0.7, "lines"))

Marker_dot
```

##reorder clusters according to hypothesized cell-types
    "0", "2", "6", "8", "11", #Mesophyll
    "9","13", #Epidermis
    "4","12", #Protodermal
    "5", "7", "10", #Proliferating
    "1", "14", #Vessels
    "3", #Vasculature Sieve
    "15", #Guard cells
    "16" #IPAP
```{r}
leaf_cell_type_assignment <- data.frame(
  cluster = c(0:16)
) %>% 
  mutate(cell_type = case_when(
    cluster == 0 | 
      cluster == 2 | 
      cluster == 6 |
      cluster == 8 |
      cluster == 11 ~ "Mesophyll",
    cluster == 1 |
      cluster == 14 ~ "Vessels",
    cluster == 3 ~ "Sieve",
    cluster == 4 |
      cluster == 12 ~ "Protodermal",
      cluster == 9 |
      cluster == 13  ~ "Epidermis",
      cluster == 5 |  
      cluster == 7 |
      cluster == 10 ~ "Proliferating",
    cluster == 15  ~ "Guard cell",
    cluster == 16 ~ "IPAP",
    T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
    "Mesophyll", "Epidermis", "Protodermal", "Proliferating", "Vessels", "Sieve", "Guard cell", "IPAP"
  ))) %>% 
  mutate(cluster = factor(cluster, levels = c(
    "0", "2", "6", "8", "11", #Mesophyll
    "9","13", #Epidermis
    "4","12", #Protodermal
    "5", "7", "10", #Proliferating
    "1", "14", #Vessels
    "3", #Vasculature Sieves
    "15", #Guard cells
    "16" #IPAP
  )))

Leaf_cell_type_strip <- leaf_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "" )) +
  geom_tile(aes(fill = cell_type)) +
  scale_fill_manual(values = c(brewer.pal(8, "Paired"), "grey80")) +
  guides(fill = guide_legend(nrow = 2)) +
  labs(fill = "Cell type") +
  theme_void() +
  theme(
    legend.position = "bottom" ,
    text = element_text(size = 14)
  )

Leaf_cell_type_strip
ggsave(file = "/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/leaf_cell_strip.svg", height = 1, width = 12)
```


```{r}
Leaf_cell_type_mkr_text <- Marker_dotplot_nice_df %>% 
  dplyr::filter(id == 0) %>% 
  ggplot(aes(x = "", y = ylab2)) +
  geom_text(aes(label = Marker_ID), fontface = "italic") +
  theme_void()
Leaf_cell_type_mkr_text

blank <- data.frame(
  x = 0,
  y = 0
) %>% 
  ggplot(aes(x, y)) +
  theme_void()

Marker_dot_strip <- wrap_plots(Marker_dot, Leaf_cell_type_mkr_text,
          Leaf_cell_type_strip, blank,
          heights = c(1, 0.03), 
          widths = c(1, 0.2), nrow = 2, ncol = 2)

Marker_dot_strip
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/plumeria_leaf_cluster_assignment_23March26.png", height = 10, width = 12)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/plumeria_leaf_cluster_assignment_23March26.svg", height = 10, width = 12)
```

# UMAP with cell type info
```{r}
leaf.combined <- readRDS("/scratch/ac05869/plob_sc/seurat_data/PLU_leaf_cleaned_integrated.rds")

leaf.combined@meta.data <- leaf.combined@meta.data %>% 
  mutate(cell_type = case_when(
    seurat_clusters == 0 | 
      seurat_clusters == 2 | 
      seurat_clusters == 6 |
      seurat_clusters == 8 |
      seurat_clusters == 11 ~ "Mesophyll",
    seurat_clusters == 1 |
      seurat_clusters == 14 ~ "Vessels",
    seurat_clusters == 3 ~ "Sieve",
    seurat_clusters == 4 |
      seurat_clusters == 12 ~ "Protodermal",
      seurat_clusters == 9 |
      seurat_clusters == 13  ~ "Epidermis",
      seurat_clusters == 5 |  
      seurat_clusters == 7 |
      seurat_clusters == 10 ~ "Proliferating",
    seurat_clusters == 15  ~ "Guard cell",
    seurat_clusters == 16 ~ "IPAP",
    T ~ "Unassigned"
  ))%>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c( "Mesophyll", "Epidermis", "Protodermal", "Proliferating", "Vessels", "Sieve", "Guard cell", "IPAP"
                            )))

DimPlot(leaf.combined, reduction = "umap", label = T,
        group.by = "cell_type", label.size = 5, repel = T, #alpha = 0.8
        ) +
  theme_void() +
  scale_color_manual(values = c(brewer.pal(8, "Paired"), "grey80")) +
  labs(title = "") +
  theme(
    text = element_text(size = 14, color = "black"),
    legend.position = "none")

c(brewer.pal(8, "Paired"), "grey80")

UMAP3 <- DimPlot(leaf.combined, reduction = "umap", label = T, label.size = 5, repel = T, alpha = 0.8
        ) +
  theme_void() +
#  scale_color_manual(values = c(brewer.pal(8, "Paired"), "grey80")) +
  labs(title = "") +
  theme(
    text = element_text(size = 18, color = "black"),
    legend.position = "none") +
  scale_color_manual(values = c("0" = "#A6CEE3",
                                "2" = "#A6CEE3",
                                "6" = "#A6CEE3",
                                "8" = "#A6CEE3",
                                "11" = "#A6CEE3",
                                "9" = "#1F78B4",
                                "13" = "#1F78B4",
                                "4" = "#B2DF8A",
                                "12" = "#B2DF8A",
                                "5" = "#33A02C",
                                "7" = "#33A02C",
                                "10" = "#33A02C",
                                "1" = "#FB9A99",
                                "14" = "#FB9A99",
                                "3" = "#E31A1C",
                                "15" = "#FDBF6F",
                                "16" = "#FF7F00"
  ))

UMAP3
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/UMAP_cell_types_23March26.svg", height = 4, width = 6, bg = "white")
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/UMAP_cell_types_23March26.png", height = 4, width = 6, bg = "white")
```

```{r}
leaf_assignment_graph <- leaf_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "")) +
  facet_wrap(~ cell_type, scales = "free", ncol = 2) +
  geom_point(size = 5, aes(color = cell_type)) +
  geom_text(aes(label = cluster)) +
  labs(x = NULL,
       y = NULL) +
  theme_minimal() +
  scale_color_manual(values = c(brewer.pal(8, "Paired"))) +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    text = element_text(size = 14, color = "black"),
    axis.text.x = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
    )

leaf_assignment_graph

RNA_umap <- wrap_plots(UMAP3 +
                         labs(tag = ""), leaf_assignment_graph, 
           nrow = 1, ncol = 2, widths = c(1, 0.7) , heights = c(1))

RNA_umap
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/UMAP_cell_types_legend_23March26.svg", height = 5, width = 8, bg = "white")
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/UMAP_cell_types_legend_23March26.png", height = 5, width = 8, bg = "white")
```

#stats by cell type
```{r}
stats <- leaf.combined@meta.data %>% 
  group_by(cell_type) %>% 
  dplyr::count() %>% 
  mutate(total = nrow(leaf.combined@meta.data)) %>% 
  mutate(percentage = n/total) %>% 
  ungroup()
stats

write.csv(stats, "/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/cell_type_proportions_table_23March26.csv", row.names = FALSE)
```


```{r}
RNA_by_rep <- leaf.combined@meta.data %>% 
  group_by(cell_type, Run) %>% 
  dplyr::count() %>% 
  inner_join(
    leaf.combined@meta.data %>% 
  group_by(Run) %>% 
  dplyr::count() %>% 
  ungroup() %>% 
  dplyr::rename(total = n), 
  by = "Run"
  ) %>% 
  mutate(percentage = n/total * 100) %>% 
  ungroup() %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c("Mesophyll", "Epidermis", "Protodermal", "Proliferating", "Vessels", "Sieve", "Guard cell", "IPAP"
                            ))) %>% 
  mutate(cell_type = fct_rev(cell_type)) %>% 
  mutate(Run = case_when(
    Run == "PLU_BB" ~ "Rep1",
    Run == "PLU_BC" ~ "Rep2"
  ))#
write.csv(RNA_by_rep, "/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/cell_type_proportions_reps_table_23March26.csv", row.names = FALSE)

RNA_by_rep %>% 
  ggplot(aes(x = Run, y = percentage)) +
  geom_bar(stat = "identity", aes(fill = cell_type)) +
  scale_fill_manual(values = rev(c(brewer.pal(8, "Paired")))) +
  labs(x = "", 
       y = "% cells",
       fill = "cell type") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, color = "black"))

ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/cell_type_proportions_23March26.svg", height = 6, width = 3, bg = "white")
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/cell_type_proportions_23March26.png", height = 6, width = 3, bg = "white")
```

# Plot Pathway genes
```{r}
dotplot_data <- DotPlot(leaf.combined, features = unique(rev_path$gene_ID))$data

p1_nice_df<- dotplot_data %>% 
  inner_join(rev_path %>% 
               distinct(gene_ID, .keep_all = T), by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    LocusID = reorder(LocusID, -order),
    ylab = paste0(features.plot),  #for plot y axis
    ylab2 = paste0(tag, ": ", LocusID),  # make this new variable
    ylab = reorder(ylab, -order),
    ylab = as.factor(ylab),
    subgenome = paste("Subgenome", subgenome)
) #
p1 <- p1_nice_df %>%
  ggplot(aes(x = id, y = ylab)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "",
#      title = "Plumeria sp. leaf syntelogs",
#       subtitle = "MEP and Iridoid genes"
      ) +
#  facet_wrap(~subgenome) +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.height = unit(0.7, "lines"),
        legend.position = "top",
        legend.box = "horizontal")
p1
```

```{r}
p1_text <- p1_nice_df %>% 
  dplyr::filter(id == 0) %>% 
  ggplot(aes(x = "", y = ylab)) +
  geom_text(aes(label = tag), fontface = "italic") +
  theme_void()
p1_text

blank <- data.frame(
  x = 0,
  y = 0
) %>% 
  ggplot(aes(x, y)) +
  theme_void()

p1_annotated <- wrap_plots(p1, p1_text,
          Leaf_cell_type_strip, blank,
          heights = c(1, 0.03), 
          widths = c(1, 0.2), nrow = 2, ncol = 2)
p1_annotated


ggsave("/scratch/ac05869/plob_sc/pathway_analysis/plumeria_leaf_pathway_23March26.png", height = 10, width = 12)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/plumeria_leaf_pathway_23March26.svg", height = 10, width = 12)
```

##Expression by subgenome
```{r}
label_map <- p1_nice_df %>%
  distinct(ylab, LocusID)

p2 <- p1_nice_df %>%
  ggplot(aes(x = id, y = ylab)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  scale_y_discrete(labels = setNames(p1_nice_df$ylab2, p1_nice_df$ylab)) +
  labs(y = "Genes",
       x = "",
      ) +
  facet_wrap(~subgenome, scales = "free_y") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.height = unit(0.7, "lines"),
        legend.position = "top",
        legend.box = "horizontal")
p2
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/plumeria_leaf_pathway_subgenomes_23March26.png", height = 8, width = 12)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/plumeria_leaf_pathway_subgenomes_23March26.svg", height = 8, width = 12)
```

## Cell type strip
```{r}
Leaf_cell_type_strip2 <- leaf_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "" )) +
  geom_tile(aes(fill = cell_type)) +
  scale_fill_manual(values = c(brewer.pal(8, "Paired"), "grey80")) +
  guides(fill = guide_legend(nrow = 2)) +
  labs(fill = "Cell type") +
  theme_void() +
  theme(legend.position = "none",
    strip.text = element_blank(),
    strip.background = element_blank()
  )

Leaf_cell_type_strip2
ggsave(file = "/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/leaf_cell_strip.svg", height = 1, width = 12)
```


```{r}
p2_strip_A <- leaf_cell_type_assignment %>% 
  mutate(subgenome = "A")
p2_strip_B <- leaf_cell_type_assignment %>% 
  mutate(subgenome = "B")

p2_strip <- bind_rows(p2_strip_A, p2_strip_B) %>%
  ggplot(aes(x = cluster, y = "" )) +
  geom_tile(aes(fill = cell_type)) +
  scale_fill_manual(values = c(brewer.pal(8, "Paired"), "grey80")) +
  facet_wrap(~subgenome) +
  theme_void() +
  theme(legend.position = "none",
    strip.text = element_blank(),
    strip.background = element_blank()
  )

p2_strip


p2_annotated <- wrap_plots(p2,
          p2_strip,
          heights = c(1, 0.03), 
          widths = c(1), nrow = 2, ncol = 1)
p2_annotated

```














_____________old analyses

```{r}
celltypes <- dotplot %>% 
  inner_join(markers3 %>%
               filter(cell_type2 == 'Companion cell') %>%
               mutate(order = as.numeric(as.factor(cell_type2))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type,':', Name, "-", features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Leaf Cluster",
       title = "") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/companion_cells_17Dec25.svg", height = 24, width = 10)
```
```{r}
celltypes <- dotplot %>%
  inner_join(
    markers3 %>%
      filter(
        str_detect(Name, regex("guard", ignore_case = TRUE)) |
        str_detect(cell_type, regex("guard", ignore_case = TRUE)) |
        str_detect(Curator_summary, regex("guard", ignore_case = TRUE))
      ) %>% mutate(order = as.numeric(as.factor(cell_type2))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type,':', Name, "-", features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Leaf Cluster",
       title = "") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/guard_cells_17Dec25.svg", height = 24, width = 10)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/guard_cells_17Dec25.png", height = 24, width = 10)

```
```{r}
celltypes <- dotplot %>% 
  inner_join(markers3 %>%
               filter(cell_type == 'Trichome') %>%
               mutate(order = as.numeric(as.factor(cell_type2))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type,':', Name, "-", features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Leaf Cluster",
       title = "") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/trichome_17Dec25.svg", height = 24, width = 10)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/trichome_17Dec25.png", height = 24, width = 10)

```
```{r}
celltypes <- dotplot %>% 
  inner_join(markers3 %>%
               filter(cell_type2 == 'Vasculature') %>%
               mutate(order = as.numeric(as.factor(cell_type2))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type,':', Name, "-", features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Leaf Cluster",
       title = "") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/vasculature_17Dec25.svg", height = 24, width = 10)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/vasculature_17Dec25.png", height = 24, width = 10)

```
```{r}
celltypes <- dotplot %>% 
  inner_join(markers3 %>%
               filter(cell_type2 == 'Phloem' | cell_type2 == "Procambium" | cell_type2 == "Procambium-PP") %>%
               mutate(order = as.numeric(as.factor(cell_type2))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type,':', Name, "-", features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Leaf Cluster",
       title = "") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/phloem_17Dec25.svg", height = 24, width = 10)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/phloem_17Dec25.png", height = 24, width = 10)

```
```{r}
celltypes <- dotplot %>% 
  inner_join(markers3 %>%
               filter(cell_type2 == 'Xylem' | cell_type2 == "Procambium-XP") %>%
               mutate(order = as.numeric(as.factor(cell_type2))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type,':', Name, "-", features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Genes",
       x = "Leaf Cluster",
       title = "") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/xylem_17Dec25.svg", height = 24, width = 10)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/xylem_17Dec25.png", height = 24, width = 10)

```

Assign clusters according to hypothesized cell-types
Cluster						
0	Mesophyll	Pluob.v4.9bG009550		LOX3		
		Pluob.v4.1bG042490		Lhcb5		
		Pluob.v4.17bG004630		GER3		
		Pluob.v4.9aG007430		CA1		
1	Vasculature	Pluob.v4.7bG024310		VND1		
		Pluob.v4.7aG007320		VND7		
		Pluob.v4.11bG024660		ATHB-14		
		Pluob.v4.3aG039710		ARF5		
2	Mesophyll	Same as 1				
3	Bundle-sheath	Pluob.v4.10bG024590		CALS7		
		Pluob.v4.8bG000240		APL		
		Pluob.v4.12aG020160		BAM3		
		Pluob.v4.1aG004360		XYP1		
4	Adaxial Epidermis	Pluob.v4.14bG020300		LTPG1		
		Pluob.v4.4bG029890		LACS2		
		Pluob.v4.10bG000380		KCS1		
		Pluob.v4.1bG021330		protodermal factor 1		
		Pluob.v4.17bG007530		YORE-YORE 1		
5	Proliferating	Pluob.v4.6aG007520		AT1G63100		
		Pluob.v4.11bG036540		gamma histone variant H2AX	Leaf	seurat denovo
6	Mesophyll	same as 1				
7	Proliferating	Pluob.v4.6aG007520		AT1G63100		
		Pluob.v4.8bG003230		AT3G20150		
8	Mesophyll	Pluob.v4.11bG007140		ENH1		
		Pluob.v4.10aG017320		Lhcb4.1		
		Pluob.v4.9aG013120		RbcS2B		
		Pluob.v4.5bG009320		STOMAGEN		
9	Unassigned	Pluob.v4.8bG022770		AT1G67360		
		Pluob.v4.13bG014160		MET2B		
		Pluob.v4.2aG021410		AT1G18210		
10	Proliferating	same as 5 and 7				
11	Mesophyll	Pluob.v4.5bG027550		rubisco activase	Leaf	seurat denovo
		Pluob.v4.3bG033470		chlorophyll binding family	Leaf	seurat denovo
12	Abaxial Epidermis	Pluob.v4.5aG027150		ANAC034		
13	Unknown	Same as 9				
14	Vasculature	same as 1 and 3				
15	Guard cell	Pluob.v4.3aG021150		KAT1	potassium channel in Arabidopsis thaliana 1
		Pluob.v4.4bG041960		KCS1	3-ketoacyl-CoA synthase 1
		Pluob.v4.8aG005190		GPT1	glucose 6-phosphate/phosphate translocator 1
		Pluob.v4.17aG023920		PBD2	20S proteasome beta subunit D2
		Pluob.v4.11aG032320		SAD1	Small nuclear ribonucleoprotein family protein
		Pluob.v4.17bG003800		AT2G19450	membrane bound O-acyl transferase (MBOAT) family protein
		Pluob.v4.11bG037190		TPK1	thiamin pyrophosphokinase1
		Pluob.v4.10bG021090		CAX7	calcium exchanger 7
		Pluob.v4.14aG008020		mMDH2	Lactate/malate dehydrogenase family protein
16	IPAP	Pluob.v4.18aG008600		IO
  Pluob.v4.4bG014130		7DLH		

Filter by the following genes in marker list:
c("Pluob.v4.9bG009550",	"Pluob.v4.1bG042490",	"Pluob.v4.17bG004630",	"Pluob.v4.9aG007430",	"Pluob.v4.7bG024310",	"Pluob.v4.7aG007320",	"Pluob.v4.11bG024660",	"Pluob.v4.3aG039710",	"Pluob.v4.10bG024590",	"Pluob.v4.8bG000240",	"Pluob.v4.12aG020160",	"Pluob.v4.1aG004360",	"Pluob.v4.14bG020300",	"Pluob.v4.4bG029890",	"Pluob.v4.10bG000380",	"Pluob.v4.1bG021330",	"Pluob.v4.17bG007530",	"Pluob.v4.6aG007520",	"Pluob.v4.11bG036540",	"Pluob.v4.6aG007520",	"Pluob.v4.8bG003230",	"Pluob.v4.11bG007140",	"Pluob.v4.10aG017320",	"Pluob.v4.9aG013120",	"Pluob.v4.5bG009320",	"Pluob.v4.8bG022770",	"Pluob.v4.13bG014160",	"Pluob.v4.2aG021410",	"Pluob.v4.5bG027550",	"Pluob.v4.3bG033470",	"Pluob.v4.5aG027150",	"Pluob.v4.3aG021150",	"Pluob.v4.4bG041960",	"Pluob.v4.6aG007160",	"Pluob.v4.17aG023920",	"Pluob.v4.11aG032320",	"Pluob.v4.17bG003800",	"Pluob.v4.11bG037190",	"Pluob.v4.10bG021090",	"Pluob.v4.14aG008020",	"Pluob.v4.15bG023870",	"Pluob.v4.7bG029150",	"Pluob.v4.18aG008600",	"Pluob.v4.4bG014130")

```{r}
patterns <- c("Pluob.v4.9bG009550",	"Pluob.v4.1bG042490",	"Pluob.v4.17bG004630",	"Pluob.v4.9aG007430",	"Pluob.v4.7bG024310",	"Pluob.v4.7aG007320",	"Pluob.v4.11bG024660",	"Pluob.v4.3aG039710",	"Pluob.v4.10bG024590",	"Pluob.v4.8bG000240",	"Pluob.v4.12aG020160",	"Pluob.v4.1aG004360",	"Pluob.v4.14bG020300",	"Pluob.v4.4bG029890",	"Pluob.v4.10bG000380",	"Pluob.v4.1bG021330",	"Pluob.v4.17bG007530",	"Pluob.v4.6aG007520",	"Pluob.v4.11bG036540",	"Pluob.v4.6aG007520",	"Pluob.v4.8bG003230",	"Pluob.v4.11bG007140",	"Pluob.v4.10aG017320",	"Pluob.v4.9aG013120",	"Pluob.v4.5bG009320",	"Pluob.v4.8bG022770",	"Pluob.v4.13bG014160",	"Pluob.v4.2aG021410",	"Pluob.v4.5bG027550",	"Pluob.v4.3bG033470",	"Pluob.v4.5aG027150",	"Pluob.v4.3aG021150",	"Pluob.v4.4bG041960",	"Pluob.v4.6aG007160",	"Pluob.v4.17aG023920",	"Pluob.v4.11aG032320",	"Pluob.v4.17bG003800",	"Pluob.v4.11bG037190",	"Pluob.v4.10bG021090",	"Pluob.v4.14aG008020",	"Pluob.v4.15bG023870",	"Pluob.v4.7bG029150",	"Pluob.v4.18aG008600",	"Pluob.v4.4bG014130")
pattern_regex <- str_c(patterns, collapse = "|")
leaf_markers <- expressed_markers %>% filter(str_detect(gene_ID, pattern_regex)) %>%
  mutate(Name2 = case_when(
    str_detect(Name, "^AT\\d+G") ~ Short_description,
    T ~ Name
  ))%>%
  mutate(Name = case_when(
    str_detect(Name2, "Calcium-binding EF-hand family protein") ~ "CA bind EF-hand",
    str_detect(Name2, "Rubber elongation factor") ~ "REF",
    str_detect(Name2, "GRAS family transcription factor") ~ "GRAS",
    str_detect(Name2, "Kinesin motor family protein") ~ "Kinesin motor",
    str_detect(Name2, "KNOTTED") ~ "KNOTTED",
    str_detect(Name2, "O-fucosyltransferase") ~ "(O-FucT)",
    str_detect(Name2, "S-adenosyl-L-methionine-dependent") ~ "SAM-dependent MT",
    str_detect(Name2, "MBOAT") ~ "MBOAT",
    T ~ Name2
  ))

#Add genes not in marker list
denovo <- tibble(
  Name       = c("H2AX", "rubisco activase","chlorophyll binding"),
  cell_type2  = c("Cell division", "Mesophyll","Mesophyll"),
  cell_type  = c("Cell division", "Mesophyll","Mesophyll"),
  Organ      = c("Leaf", "Leaf","Leaf"),
  gene_ID    = c("Pluob.v4.11bG036540", "Pluob.v4.5bG027550","Pluob.v4.3bG033470"),
  subgenome  = c("B", "B","B")
)

leaf_markers <- bind_rows(leaf_markers, denovo)
```


```{r}
dotplot <- DotPlot(leaf.combined, features = unique(leaf_markers$gene_ID))$data
celltypes2 <- dotplot %>% 
  inner_join(leaf_markers %>%
               mutate(order = as.numeric(as.factor(cell_type2))),
             by = c("features.plot"="gene_ID"), multiple = 'first') %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type2, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type2,':', features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Markers",
       x = "Leaf Clusters") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
       # axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes2
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/selective_marker_expression_17Dec25.png", height = 9, width = 10)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/selective_marker_expression_17Dec25.svg", height = 9, width = 10)

```
save markers as table

```{r}
write_delim(leaf_markers, file = "/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/markers_used_17Dec25.csv", delim = ",")
```

Reorder the columns by cell types
```{r}
Marker_dotplot <- DotPlot(leaf.combined, 
                          features = unique(leaf_markers$gene_ID))

Marker_dotplot_nice_df <- Marker_dotplot$data %>% 
  mutate(gene_ID = row.names(.)) %>% 
  full_join(leaf_markers %>% 
              distinct(gene_ID, .keep_all = T),
             by = join_by(features.plot == gene_ID)) %>% 
    mutate(gene_ID = str_sub(gene_ID, end = -8)) %>% 
   mutate(cell_type3 = case_when(
    str_detect(cell_type2, "Meso") ~ "Mesophyll",
    str_detect(cell_type2, "PP") ~ "Phloem Precursor",
    str_detect(cell_type2, "XP") ~ "Xylem Precursor",
    str_detect(cell_type2, "Com|Vasc|Bundle") ~ "Vasculature",
    str_detect(cell_type2, "Xylem") ~ "Xylem",
    str_detect(cell_type2, "Phloem") ~ "Phloem",
    str_detect(cell_type2, "Trich") ~ "Trichome",
    str_detect(cell_type2, "Epi") ~ "Epiderm",
    str_detect(cell_type2, "Guard") ~ "Guard",
    str_detect(cell_type2, "IPAP") ~ "IPAP",
    str_detect(cell_type2, "Cell division|Embryo")  ~ "Prolif",
    T ~ cell_type2
  )) %>% 
  mutate(cell_type3 = factor(cell_type3, levels = c(
    "Mesophyll", "Epiderm", "Trichome", "Xylem Precursor", "Phloem Precursor", "Xylem", "Phloem", "Vasculature", "Prolif","Guard", "IPAP"
  ))) %>% 
  mutate(cell_type4 = as.numeric(cell_type3)) %>% 
  mutate(gene_ID = reorder(gene_ID, -cell_type4))  %>% 
  mutate(LocusID = str_remove(
               str_sub(features.plot, end = -8),
               "^Pluob\\.v4\\."),
    ylab = paste0(features.plot),
    ylab2 = paste0(cell_type3,':', features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -cell_type4)) %>%
  mutate(id = factor(id, levels = c(
    "0", "2", "6", "8", "11", #Mesophyll
     "4","12", #Protodermal
    "1", "3", "14", #Vasculature
    "5", "7", "10", #Proliferating
    "16", #IPAP
    "9","13", #Epidermis
    "15" #Guard cells
  ))) %>% 
  dplyr::filter(is.na(avg.exp.scaled) == F)

```

0	Mesophyll
1	Vasculature
2	Mesophyll
3	Vasculature
4	Protodermal
5	Proliferating
6	Mesophyll
7	Proliferating
8	Mesophyll
9	Epidermis
10	Proliferating
11	Mesophyll
12	Protodermal
13	Epidermis
14	Vasculature
15	Guard cell
16	IPAP

```{r}
Marker_dot <- Marker_dotplot_nice_df %>% ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_fill_viridis(option = "A", begin = 0, end = 0.9) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"), nrow = 2)) +
  labs(y = "",
       x = "Leaf Clusters",
       fill = "Avg. Exp.",
       size = "% Exp.") +
  scale_y_discrete(labels = setNames(Marker_dotplot_nice_df$ylab2, Marker_dotplot_nice_df$ylab)) +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "top",
        legend.box = "horizontal",
        legend.key.height = unit(0.7, "lines"))

Marker_dot
```

reorder clusters according to hypothesized cell-types
    "0", "2", "6", "8", "11", #Mesophyll
     "4", #Adaxial Epidermis
    "12", #Abaxial Epidermis
    "1", "14", #Vasculature
    "3", #Vasculature 
    "5", "7", "10", #Proliferating
    "16", #IPAP
    "9","13", "15" #Unassigned

```{r}
leaf_cell_type_assignment <- data.frame(
  cluster = c(0:16)
) %>% 
  mutate(cell_type = case_when(
    cluster == 0 | 
      cluster == 2 | 
      cluster == 6 |
      cluster == 8 |
      cluster == 11 ~ "Mesophyll",
    cluster == 1 |
      cluster == 14 ~ "Vasculature",
    cluster == 3 ~ "Vasculature",
    cluster == 4 |
      cluster == 12 ~ "Epidermal",
      cluster == 9 |
      cluster == 13  ~ "Unassigned",
      cluster == 5 |  
      cluster == 7 |
      cluster == 10 ~ "Proliferating",
    cluster == 15  ~ "Unassigned",
    cluster == 16 ~ "IPAP",
    T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll", "Epidermal",
                              "Vasculature",
                              "Proliferating", "IPAP", "Unassigned"
                            ))) %>% 
  mutate(cluster = factor(cluster, levels = c(
    "0", "2", "6", "8", "11", #Mesophyll
     "4", #Adaxial Epidermis
    "12", #Abaxial Epidermis
    "1", "14", #Vasculature
    "3", #Vasculature 
    "5", "7", "10", #Proliferating
    "16", #IPAP
    "9","13", "15" #Unassigned
  )))

Leaf_cell_type_strip <- leaf_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "" )) +
  geom_tile(aes(fill = cell_type)) +
  scale_fill_manual(values = c(brewer.pal(5, "Paired"), "grey80")) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(fill = "Cell type") +
  theme_void() +
  theme(
    legend.position = "bottom" ,
    text = element_text(size = 14)
  )

Leaf_cell_type_strip
```


```{r}
Leaf_cell_type_mkr_text <- Marker_dotplot_nice_df %>% 
  dplyr::filter(id == 0) %>% 
  ggplot(aes(x = "", y = ylab2)) +
  geom_text(aes(label = Name), fontface = "italic") +
  theme_void()
Leaf_cell_type_mkr_text

blank <- data.frame(
  x = 0,
  y = 0
) %>% 
  ggplot(aes(x, y)) +
  theme_void()

Marker_dot_strip <- wrap_plots(Marker_dot, Leaf_cell_type_mkr_text,
          Leaf_cell_type_strip, blank,
          heights = c(1, 0.03), 
          widths = c(1, 0.2), nrow = 2, ncol = 2)

Marker_dot_strip
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/plumeria_leaf_cluster_assignment_17Dec25.png", height = 8, width = 14)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/plumeria_leaf_cluster_assignment_17Dec25.svg", height = 8, width = 14)
```

UMAP with cell type info
```{r}
leaf.combined <- readRDS("/scratch/ac05869/plob_sc/seurat_data/PLU_leaf_cleaned_integrated.rds")

leaf.combined@meta.data <- leaf.combined@meta.data %>% 
  mutate(cell_type = case_when(
    seurat_clusters == 0 | 
      seurat_clusters == 2 | 
      seurat_clusters == 6 |
      seurat_clusters == 8 |
      seurat_clusters == 11 ~ "Mesophyll",
    seurat_clusters == 1 |
      seurat_clusters == 14 ~ "Vasculature",
    seurat_clusters == 3 ~ "Vasculature",
    seurat_clusters == 4 |
      seurat_clusters == 12 ~ "Epidermal",
      seurat_clusters == 9 |
      seurat_clusters == 13  ~ "Unassigned",
      seurat_clusters == 5 |  
      seurat_clusters == 7 |
      seurat_clusters == 10 ~ "Proliferating",
    seurat_clusters == 15  ~ "Unassigned",
    seurat_clusters == 16 ~ "IPAP",
    T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll", "Epidermal",
                              "Vasculature",
                              "Proliferating", "IPAP", "Unassigned"
                            )))

DimPlot(leaf.combined, reduction = "umap", label = T,
        group.by = "cell_type", label.size = 5, repel = T
        ) +
  theme_void() +
  scale_color_manual(values = c(brewer.pal(5, "Paired"), "grey80")) +
  labs(title = "") +
  theme(
    text = element_text(size = 14, color = "black"),
    legend.position = "none")

c(brewer.pal(5, "Paired"), "grey80")

UMAP3 <- DimPlot(leaf.combined, reduction = "umap", label = T, label.size = 5, repel = T
        ) +
  theme_void() +
#  scale_color_manual(values = c(brewer.pal(8, "Paired"), "grey80")) +
  labs(title = "") +
  theme(
    text = element_text(size = 14, color = "black"),
    legend.position = "none") +
  scale_color_manual(values = c("0" = "#A6CEE3",
                                "1" = "#B2DF8A",
                                "2" = "#A6CEE3",
                                "3" = "#B2DF8A",
                                "4" = "#1F78B4",
                                "5" = "#33A02C",
                                "6" = "#A6CEE3",
                                "7" = "#33A02C",
                                "8" = "#A6CEE3",
                                "9" =  "grey80",
                                "10" = "#33A02C",
                                "11" = "#A6CEE3",
                                "12" = "#1F78B4",
                                "13" = "grey80",
                                "14" = "#B2DF8A",
                                "15" = "grey80",
                                "16" = "#FB9A99"
                                ))

UMAP3
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/UMAP_cell_types_18Dec25.svg", height = 6, width = 5, bg = "white")
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/UMAP_cell_types_18Dec25.png", height = 6, width = 5, bg = "white")
```

```{r}
leaf_assignment_graph <- leaf_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "")) +
  facet_wrap(~ cell_type, scales = "free", ncol = 2) +
  geom_point(size = 5, aes(color = cell_type)) +
  geom_text(aes(label = cluster)) +
  labs(x = NULL,
       y = NULL) +
  theme_minimal() +
  scale_color_manual(values = c(brewer.pal(8, "Paired"), "grey80")) +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    text = element_text(size = 14, color = "black"),
    axis.text.x = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
    )

leaf_assignment_graph

RNA_umap <- wrap_plots(UMAP3 +
                         labs(tag = ""), leaf_assignment_graph, 
           nrow = 2 , heights = c(1, 0.7))

RNA_umap
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/UMAP_cell_types_legend_18Dec25.svg", height = 6, width = 3, bg = "white")
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/UMAP_cell_types_legend_18Dec25.png", height = 6, width = 3, bg = "white")
```

stats by cell type
```{r}
leaf.combined@meta.data %>% 
  group_by(cell_type) %>% 
  dplyr::count() %>% 
  mutate(total = nrow(leaf.combined@meta.data)) %>% 
  mutate(percentage = n/total) %>% 
  ungroup()
```


```{r}
RNA_by_rep <- leaf.combined@meta.data %>% 
  group_by(cell_type, Run) %>% 
  dplyr::count() %>% 
  inner_join(
    leaf.combined@meta.data %>% 
  group_by(Run) %>% 
  dplyr::count() %>% 
  ungroup() %>% 
  dplyr::rename(total = n), 
  by = "Run"
  ) %>% 
  mutate(percentage = n/total * 100) %>% 
  ungroup() %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll", "Epidermal",
                              "Vasculature",
                              "Proliferating", "IPAP", "Unassigned"
                            ))) %>% 
  mutate(cell_type = fct_rev(cell_type)) %>% 
  mutate(Run = case_when(
    Run == "PLU_BB" ~ "Rep1",
    Run == "PLU_BC" ~ "Rep2"
  )) %>% 
  ggplot(aes(x = Run, y = percentage)) +
  geom_bar(stat = "identity", aes(fill = cell_type)) +
  scale_fill_manual(values = rev(c(brewer.pal(5, "Paired"), "grey80"))) +
  labs(x = "", 
       y = "% cells by cell type",
       fill = "cell type") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, color = "black"))

RNA_by_rep
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/cell_type_proportions_18Dec25.svg", height = 6, width = 3, bg = "white")
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/cell_type_proportions_18Dec25.png", height = 6, width = 3, bg = "white")
```

For annotating the single cell data with only Li's markers
```{r}
cro_plum_synt <- all_syntelogs %>%
  select(repGene, C.roseus, P.spp_A, P.spp_B)
cro_plum_synt$repGene <- gsub("\\.\\d+$", "", cro_plum_synt$repGene)
cro_markers <- read_tsv(file = "/scratch/ac05869/plob_sc/cluster_anno/leaf_markers_croseus_v3.tsv")

cell_Type_markers <- cro_plum_synt%>%
  right_join(cro_markers, by = "repGene") %>% 
  filter(
    !is.na(P.spp_A), P.spp_A != "",
    !is.na(P.spp_B), P.spp_B != ""
  )%>%
  separate_rows(P.spp_A, sep = ",") %>%
  mutate(P.spp_A = trimws(P.spp_A))%>%
  separate_rows(P.spp_B, sep = ",") %>%
  mutate(P.spp_B = trimws(P.spp_B)) %>%
  filter(!grepl("\\+", P.spp_A))%>%
  filter(!grepl("\\+", P.spp_B))

a.mark <- cell_Type_markers %>%
  select(!c(C.roseus, P.spp_B)) %>%
  unique()%>%
  rename(gene_ID = P.spp_A)
dim(a.mark)
b.mark <- cell_Type_markers %>%
  select(!c(C.roseus, P.spp_A))%>%
  unique() %>%
  rename(gene_ID = P.spp_B)

both.mark <- rbind(a.mark, b.mark) %>% unique()
both.mark$gene_ID <- gsub("\\.\\d+$", "", both.mark$gene_ID)

```

```{r}
dotplot <- DotPlot(leaf.combined, features = unique(both.mark$gene_ID))$data
celltypes2 <- dotplot %>% 
  inner_join(both.mark %>%
               mutate(order = as.numeric(as.factor(cell_type))),
             by = c("features.plot"="gene_ID")) %>%
  mutate(
    LocusID = str_sub(features.plot, end = -8),
    ylab = paste0(cell_type, ":", LocusID),  # make this new variable
    ylab2 = paste0(cell_type,':', features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -order)
) %>%
  ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
#  scale_y_continuous(
#    name = "Marker",
#    sec.axis = sec_axis(., name = ylab2)) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  labs(y = "Markers",
       x = "Leaf Clusters") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
       # axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.8, "lines"))

celltypes2
```


```{r}
Marker_dotplot_nice_df <- dotplot %>% 
  mutate(gene_ID = row.names(.)) %>% 
  full_join(both.mark %>% 
              distinct(gene_ID, .keep_all = T),
             by = join_by(features.plot == gene_ID)) %>% 
    mutate(gene_ID = str_sub(gene_ID, end = -8)) %>% 
  mutate(cell_type = factor(cell_type)) %>% 
  mutate(cell_type2 = as.numeric(cell_type)) %>% 
  mutate(gene_ID = reorder(gene_ID, -cell_type2))  %>% 
  mutate(LocusID = str_remove(
               str_sub(features.plot, end = -8),
               "^Pluob\\.v4\\."),
    ylab = paste0(features.plot),
    ylab2 = paste0(cell_type,':', features.plot),  # make this new variable
    ylab2 = reorder(ylab2, -cell_type2)) %>%
  mutate(id = factor(id, levels = c(
    "0", "2", "6", "8", "11", #Mesophyll
     "4", #Adaxial Epidermis
    "12", #Abaxial Epidermis
    "1", "14", #Vasculature
    "3", #Vasculature 
    "5", "7", "10", #Proliferating
    "16", #IPAP
    "9","13", "15" #Unassigned
  ))) %>% 
  dplyr::filter(is.na(avg.exp.scaled) == F)

Marker_dot <- Marker_dotplot_nice_df %>% ggplot(aes(x = id, y = ylab2)) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_fill_viridis(option = "A", begin = 0, end = 0.9) +
  scale_color_gradientn(colors = viridis(10, option = "A")[1:9]) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"), nrow = 2)) +
  labs(y = "",
       x = "Leaf Clusters",
       fill = "Avg. Exp.",
       size = "% Exp.") +
  scale_y_discrete(labels = setNames(Marker_dotplot_nice_df$ylab2, Marker_dotplot_nice_df$ylab)) +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "top",
        legend.box = "horizontal",
        legend.key.height = unit(0.7, "lines"))


Leaf_mrkr_txt <- Marker_dotplot_nice_df %>% 
  dplyr::filter(id == 0) %>% 
  ggplot(aes(x = "", y = ylab2)) +
  geom_text(aes(label = Symbol), fontface = "italic") +
  theme_void()
Leaf_mrkr_txt

blank <- data.frame(
  x = 0,
  y = 0
) %>% 
  ggplot(aes(x, y)) +
  theme_void()

Marker_dot_strip <- wrap_plots(Marker_dot, Leaf_mrkr_txt, blank,
          heights = c(1, 0.03), 
          widths = c(1, 0.2), nrow = 2, ncol = 2)

Marker_dot_strip
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/catharanthus_marker_expression_22jan26.png", height = 9, width = 10)
ggsave("/scratch/ac05869/plob_sc/pathway_analysis/markers/leaf/catharanthus_marker_expression_22jan26.svg", height = 9, width = 10)

```
