---
title: "UGA_kratom_leaf_multiome_pipseq_coexp"
author: "Anne Pavia"
date: "2025-06-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_knit$set(root.dir = '/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/wd')
#setwd("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/wd")
```

#Packages
```{r}
library(tidyverse)
library(igraph)
library(ggraph)
library(Seurat)
library(readxl)
library(patchwork)
library(RColorBrewer)
library(viridis)
library(showtext)
library(ape)
#showtext_auto(enable = T)
set.seed(666)
options(bitmapType='cairo')
```

#Load data
```{r}
KRT_RNA <- readRDS("/scratch/ac05869/10X_Multiome/KRT_leaf/seurat_out/Integrated_RNA_Data/KRT_RNA_integrated_5june25.rds")
DefaultAssay(KRT_RNA) <- "RNA"
cluster.averages <- AverageExpression(KRT_RNA, return.seurat = TRUE)
cluster.averages
cluster_avg.data <- as.matrix(GetAssayData(cluster.averages, layer = "data")[, ]) 

Exp_table <- cluster_avg.data #"As of Seurat v5, we recommend using AggregateExpression to perform pseudo-bulk analysis."
colnames(Exp_table) <- as.factor(c(0:21)) #changes based on cluster numbers

head(Exp_table)
dim(Exp_table) #75499 genes
```
#format pathway genes
```{r}
#Annotation
anno <- read.delim(file = "/scratch/ac05869/nih/kratom/mitr_v1_anno/mitr_v1.functional_annotation.txt", header = FALSE, sep = "\t", col.names = c("gene", "functional_annotation"))
#Don't remove transcript syntax because I want to select only the representative transcript
#anno$gene <- gsub("\\.\\d+$", "", anno$gene)
head(anno)
nrow(anno) #172819
anno <- anno %>% rename(gene_ID = gene)

##Read in repr gff file
working_models <- read.gff("/work/crblab/ac05869/genomes/mitr_v1.working_models.repr.gff3", na.strings = c(".", "?"), GFF3 = TRUE)
head(working_models)
#parse apart attributes column
head(attributes <- working_models$attributes)
head(parsed_attributes <- strsplit(attributes, ";"))
extract_attribute_value <- function(attribute_list, key) {
  key_value <- grep(paste0("^", key, "="), attribute_list, value = TRUE)
  if (length(key_value) > 0) {
    return(sub(paste0("^", key, "="), "", key_value))
  } else {
    return(NA)
  }
}
working_models$Name <- sapply(parsed_attributes, extract_attribute_value, key = "Name")
working_models$ID <- sapply(parsed_attributes, extract_attribute_value, key = "ID")
working_models$Parent <- sapply(parsed_attributes, extract_attribute_value, key = "Parent")

working_models2 <- working_models %>%
  select(c(ID, seqid, type, start, end, strand)) %>% 
  filter(type == "mRNA") %>%
  rename(chr = seqid) %>%
  rename(gene_ID = ID)
#working_models2$gene <- gsub("\\.\\d+$", "", working_models2$gene) #Don't remove transcript syntax yet. Need to select repr transcript in annotation file
head(working_models2)
dim(working_models2) #103160

repr_anno <- anno %>%
  right_join(working_models2, by = "gene_ID") %>%
  rename("gene" = "gene_ID") %>%
  mutate(gene_ID = gsub("\\.\\d+$", "", gene)) %>%
  mutate(gene_ID = gsub("_", "-", gene_ID))
dim(repr_anno)#103160

write_delim(repr_anno, file = "/scratch/ac05869/nih/kratom/mitr_v1_anno/mitr_v1.working_models.repr.functional_annotation.txt")
repr_anno <- read_delim(file = "/scratch/ac05869/nih/kratom/mitr_v1_anno/mitr_v1.working_models.repr.functional_annotation.txt")


## genes 
pathway_genes <- read_delim("/scratch/ac05869/10X_Multiome/KRT_leaf/seurat_out/Data/pathway_genes_allwins_and_cro_orthologs.tsv") #contains both cro orthologs and allwins best hits
pathway_genes <- pathway_genes %>%
  left_join(repr_anno, by = "gene_ID")
#write_delim(pathway_genes, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/seurat_out/Data/annotated_kratom_pathway.tsv", delim = "\t")
pathway_genes <- read_delim(file = "/scratch/ac05869/10X_Multiome/KRT_leaf/seurat_out/Data/annotated_kratom_pathway.tsv")
rev_pathway_genes <- pathway_genes[order(nrow(pathway_genes):1),]
head(rev_pathway_genes)

Baits <- read_delim("/scratch/ac05869/10X_Multiome/KRT_leaf/seurat_out/Data/allwin_goi_top_blast.txt", delim = "\t")
Baits <- Baits %>% left_join(repr_anno, by = "gene_ID")
head(Baits)
#write_delim(Baits, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/seurat_out/Data/annotated_allwin_goi_top_blast.txt")
Baits <- read_delim(file = "/scratch/ac05869/10X_Multiome/KRT_leaf/seurat_out/Data/annotated_allwin_goi_top_blast.txt")
```
#Load pathway genes
```{r}
repr_anno <- read_delim(file = "/scratch/ac05869/nih/kratom/mitr_v1_anno/mitr_v1.working_models.repr.functional_annotation.txt")
Baits <- read_delim(file = "/scratch/ac05869/10X_Multiome/KRT_leaf/seurat_out/Data/annotated_allwin_goi_top_blast.txt")
# replace spaces with underscores '_' so ggplot2 doesn't fail
Baits$gene_ID <- gsub(pattern = "_", replacement = "-", x = Baits$gene_ID)
# remove transcript ID so it matches the seurat genes
Baits$gene_ID <- gsub("\\.\\d+$", "", x = Baits$gene_ID)
pathway_genes <- read_delim(file = "/scratch/ac05869/10X_Multiome/KRT_leaf/seurat_out/Data/annotated_kratom_pathway.tsv")
# replace spaces with underscores '_' so ggplot2 doesn't fail
pathway_genes$gene_ID <- gsub(pattern = "_", replacement = "-", x = pathway_genes$gene_ID)
# remove transcript ID so it matches the seurat genes
pathway_genes$gene_ID <- gsub("\\.\\d+$", "", x = pathway_genes$gene_ID)
rev_pathway_genes <- pathway_genes[order(nrow(pathway_genes):1),]
head(rev_pathway_genes)
```

```{r}
Exp_table_long_averaged <- Exp_table %>%
  as.data.frame() %>%
  rownames_to_column("gene_ID") %>% 
  pivot_longer(cols = !gene_ID, names_to = "cluster", values_to = "avg_exp") 
head(Exp_table_long_averaged)
```

## Z score
It standardize the expression pattern of each gene to mean = 0, sd = 1. It is not absolutely necessary, but I have found including this step to produce results that better capture the underlying biology.
```{r}
Exp_table_long_averaged_z <- Exp_table_long_averaged %>% 
  group_by(gene_ID) %>% 
  mutate(z.score = (avg_exp - mean(avg_exp))/sd(avg_exp)) %>% 
  ungroup()

head(Exp_table_long_averaged_z)
```
## Gene-wise correlation

Now we can correlate each gene to every other gene. The essence of this workflow is simple, so we will use a simple correlation. If you want, you can use fancier methods such as GENIE3

We will use the cor() function in R. But the cor() only take vector or matrix as input, so we need to go from long to wide again.

```{r}
#need wide z-score
z_score_wide <- Exp_table_long_averaged_z %>% 
  select(gene_ID, cluster, z.score) %>% 
  pivot_wider(names_from = cluster, values_from = z.score) %>% 
  as.data.frame()

row.names(z_score_wide) <- z_score_wide$gene_ID
head(z_score_wide)

number_of_clusters <- ncol(z_score_wide) - 1
number_of_clusters #22

```

# Generate Edge Table for Co-expression networks with > 30,000 genes

```{r}
cor_matrix <- cor(t(z_score_wide[, -1]))
dim(cor_matrix) #75499
```

Before we select edges (correlations), we need to deal with some redundant data. The correlation matrix is symmetrical along its diagonal. The diagonal will be 1, because it is correlating with itself. Everything else appears twice. We can take care of that by setting the upper (or lower) triangle of this matrix to NA. This step can take a while. The larger the matrix, the slower it is.

```{r}
cor_matrix_upper_tri <- cor_matrix
cor_matrix_upper_tri[lower.tri(cor_matrix_upper_tri)] <- NA

# Create an empty dataframe to store the results
edge_table <- data.frame()

chunk_size <- 1000  # Define the chunk size

num_rows <- nrow(cor_matrix_upper_tri) #number of rows in correlation matrix
num_chunks <- ceiling(num_rows / chunk_size) # number of chunks for data set

for (i in 1:num_chunks) {
  start <- (i - 1) * chunk_size + 1 
  end <- min(i * chunk_size, num_rows)
  
  # Extract a chunk of rows
  chunk <- cor_matrix_upper_tri[start:end, ]
  
  #Wide to long form
  chunk <- chunk %>%
    as.data.frame() %>%
    mutate(from = row.names(chunk)) %>%
    pivot_longer(cols = !from, names_to = "to", values_to = "r") %>%
    filter(is.na(r) == F) %>%
    filter(from != to)
  
  chunk <- chunk %>%
  mutate(t = r*sqrt((number_of_clusters-2)/(1-r^2))) %>%
  mutate(p.value = case_when(
    t > 0 ~ pt(t, df = number_of_clusters-2, lower.tail = F),
    t <=0 ~ pt(t, df = number_of_clusters-2, lower.tail = T)
  )) %>%
  mutate(FDR = p.adjust(p.value, method = "fdr"))
  
  chunk <-chunk %>% 
  filter(r > 0) %>% 
  filter(FDR < 0.05)
  
  # Add the processed chunk to the result dataframe
  edge_table <- bind_rows(edge_table, chunk)
  
  # Print the current process
  cat("Processed chunk", i, "of", num_chunks, "\n")
}

#saveRDS(edge_table, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/leaf_edge_table_pos_corr_p0.05.rds")
edge_table <- readRDS(file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/leaf_edge_table_pos_corr_p0.01.rds")
dim(edge_table2) #35200647
```


You can look at various adjusted p value cutoffs and the corresponding r value before proceeding. Let's say we just look at positively correlated genes.
I already filtered at FDR < 0.01 in order to process the chuncks above, because the dataframe was too big for R
```{r}
edge_table %>% 
  filter(r > 0) %>% 
  filter(r > 0.65) %>% 
  slice_min(order_by = abs(r), n = 10)

edge_table %>% 
  filter(r > 0) %>% 
  filter(FDR < 0.05) %>% 
  slice_min(order_by = abs(r), n = 10)
```


> FDR cutoff at 0.01, r values are 0.6459088  or larger
> cutoff at 0.05, r values are 0.5570972 or larger

### Empirical determination using bait genes and rank distribution 
If I go into this analysis not knowing any biology, then I would proceed with a t approximation followed by some p value cutoff.
I think in real life, this is hardly the case. We usually know something a priori. 
This is where bait genes can be helpful. 
You can use the bait genes to determine the cutoff if you know two bait genes are involved in the same process. 
The underlying assumption is if two bait genes are involved in the same process, they might be co-expressed. 
Because this selection method is based on empirical observations, I argue this is better than using an arbitrary p value cutoff. 
> Skip this until after filtering the edge_table so it doesn't crash


You can also look at the distribution of r values. 
Won't work in this case, because I already filtered
```{r}
edge_table %>% 
  slice_sample(n = 20000) %>% 
  ggplot(aes(x = r)) +
  geom_histogram(color = "white", bins = 100) +
  geom_vline(xintercept = 0.7, color = "tomato1", linewidth = 1.2) +
  theme_classic() +
  theme(
    text = element_text(size = 14),
    axis.text = element_text(color = "black")
  )

#ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/r_histogram_r.7.svg", height = 3.5, width = 5, bg = "white")
#ggsave("/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/r_histogram_r.7.png", height = 3.5, width = 5, bg = "white")
```

Cutoff at r = 0.7

Note that there are many negatively correlated genes, we can look at those at well.
But for the sake of this example, let's just look at positively correlated genes. 

#Create edge table
```{r}
edge_table_select <- edge_table #I don't want to filter it any more than I did

dim(edge_table_select)#35200647 for edges p<0.01; 72531411 for p<0.05
head(edge_table_select)
#saveRDS(edge_table_select, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/UGA_leaf_edge_table_select_p0.05.rds")
edge_table_select2 <- readRDS(file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/UGA_leaf_edge_table_select_p0.05.rds")
#saveRDS(edge_table_select, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/UGA_leaf_edge_table_select_p0.01.rds")
edge_table_select <- readRDS("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/UGA_leaf_edge_table_select_p0.01.rds")

```

> We are now down to 72531411 edges (for p<0.05 dataset), and 35200647 for p<0.01 dataset

--- left off here

### Check bait genes after filtering for high correlated genes
```{r}
head(Baits)
edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.11-1G000250") & #enolMT
           str_detect(to,"Mitsp.v1.10-2G022130") | #CO1
         str_detect(from, "Mitsp.v1.10-2G022130") &
           str_detect(to,"Mitsp.v1.11-1G000250")) 
#None
edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.11-1G000250") & #enolMT
           str_detect(to,"Mitsp.v1.10-1G021390") | #CO2
         str_detect(from, "Mitsp.v1.10-1G021390") &
           str_detect(to,"Mitsp.v1.11-1G000250")) 
#None
edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.11-1G000250") & #enolMT
           str_detect(to,"Mitsp.v1.01-2G031870") | #THAS
         str_detect(from, "Mitsp.v1.01-2G031870") &
           str_detect(to,"Mitsp.v1.11-1G000250")) 
#None
edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.10-2G022130") & #CO1
           str_detect(to,"Mitsp.v1.10-2G012340") | #DCR2
         str_detect(from, "Mitsp.v1.10-2G012340") &
           str_detect(to,"Mitsp.v1.10-2G022130")) 
#None between CO1 and DCR1 haplotypes 1 and 2, or with DCR2

edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.10-1G021390") & #CO2
           str_detect(to,"Mitsp.v1.10-2G012340") | #DCR2
         str_detect(from, "Mitsp.v1.10-2G012340") &
           str_detect(to,"Mitsp.v1.10-1G021390")) 
#None between CO2 and DCR1, but 0.6738474 corr with DCR2

edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.10-1G010980") & #DCR1
           str_detect(to,"Mitsp.v1.18-2G016770") | #SAS
         str_detect(from, "Mitsp.v1.18-2G016770") &
           str_detect(to,"Mitsp.v1.10-1G010980")) 
#0.7045289	

edge_table_select %>%
  filter(str_detect(from, "Mitsp.v1.18-2G016770") & #SAS
           str_detect(to,"Mitsp.v1.10-2G012340") | #DCR2
         str_detect(from, "Mitsp.v1.10-2G012340") &
           str_detect(to,"Mitsp.v1.18-2G016770")) 
#0.7937364

```



coexpression notes:
>enolMT has no correlation with CO1, CO2, or THAS (because they are in different cell types)
>correlations pairwise between CO1, CO2, and DCR1, DCR2 have only correlation between CO2 and DCR2 (0.6738474)
>DCR1 and DCR2 have correlations with SAS (DCR2 is higher at 0.7937364)
>

## Module detection
The main goal of a gene co-expression analysis to detect gene co-expression modules, groups of highly co-expressed genes. 
We will be the Leiden algorithm to detect module, which is a graph based clustering method. 
The Leiden method produces clusters in which members are highly interconnected. 
In gene co-expression terms, it looks for groups of genes that are highly correlated with each other. 
If you are interested, you can read more about it in this [review](https://www.nature.com/articles/s41598-019-41695-z ).

### Build graph object 
We will be using `igraph` to do some of the downstream analyses. It will do a lot of the heavy lifting for us. 
While you can get Leiden as a standalone package, Leiden is also part of the `igraph` package. 
The first thing to do is producing a graph object, also known as a network object. 

To make a graph object, you need a edge table. 
We already made that, which is `edge_table_select`, a edge table that we filtered based on some kind of r cutoff. 
Optionally, we can also provide a node table, which contains information about all the nodes present in this network. 
We can make that. 

We need to two things. 

1. Non-redundant gene IDs from the edge table
2. Functional annotation, which I [downloaded](http://spuddb.uga.edu/m82_uga_v1_download.shtml ).
```{r}
mitr_v1.funct_anno <- repr_anno #(/scratch/ac05869/nih/kratom/mitr_v1_anno/mitr_v1.working_models.repr.functional_annotation.txt)
mitr_v1.funct_anno$gene_ID <- gsub(pattern = "_", replacement = "-", mitr_v1.funct_anno$gene_ID)
mitr_v1.funct_anno$gene_ID <- gsub("\\.\\d+$", "", mitr_v1.funct_anno$gene_ID)

head(mitr_v1.funct_anno)
```



```{r}
node_table <- data.frame(
  gene_ID = c(edge_table_select$from, edge_table_select$to) %>% 
    unique()
  ) %>% 
  left_join(mitr_v1.funct_anno, by = "gene_ID")

node_table2 <- data.frame(
  gene_ID = c(edge_table_select2$from, edge_table_select2$to) %>% 
    unique()
  ) %>% 
  left_join(mitr_v1.funct_anno, by = "gene_ID")

head(node_table)
dim(node_table) #75463 for p<0.01 dataset, 
dim(node_table2) #75469 for p<0.05 dataset, 

saveRDS(node_table2, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/Integrated_leaf_node_table_p0.05.rds")
saveRDS(node_table, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/Integrated_leaf_node_table_p0.01.rds")

node_table <- readRDS(file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/Integrated_leaf_node_table_p0.01.rds")
```

> We have 75463 genes in this network, along with 35200647 edges (p<0.01)
Now let's make the network object. 

```{r}
my_network <- graph_from_data_frame(
  edge_table_select,
  vertices = node_table,
  directed = F
)

#saveRDS(my_network, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/Integrated_leaf_network_p0.05.rds")
#my_network <- readRDS("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/Integrated_leaf_network_p0.05.rds")
saveRDS(my_network, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/Integrated_leaf_network_p0.01.rds")
my_network <- readRDS("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/Integrated_leaf_network_p0.01.rds")
```

`graph_from_data_frame()` is a function from the `igraph` package. 
It takes your edge table and node table and produce a graph (aka network) from it. 
Note that I selected the `directed = F` argument, because we made our network using correlation.
Correlation is non-directional, because cor(A,B) = cor(B,A). 

### What is the optimal resolution for module detection? 
The optimal resolution for module detection differs between networks. 
A key factor that contributes to the difference in optimal resolution is to what extent are nodes inter-connected. 

Since this is a simple workflow, we can determine the optimal resolution using heuristics. 
We can test a range of resolutions and monitor two key performance indexes:

1. Optimize number of modules that have >= 5 genes.
2. Optimize number of genes that are contained in modules that have >= 5 genes. 

Because: 

* Too low resolution leads to forcing genes with different expression patterns into the same module.
* Too high resolution leads to many genes not contained in any one module. 

```{r}
optimize_resolution <- function(network, resolution){
  modules = network %>% 
    cluster_leiden(resolution = resolution,
                   objective_function = "modularity")
  
  parsed_modules = data.frame(
    gene_ID = names(membership(modules)),
    module = as.vector(membership(modules)) 
    )
  
  num_module_5 = parsed_modules %>% 
    group_by(module) %>% 
    count() %>% 
    arrange(-n) %>% 
    filter(n >= 5) %>% 
    nrow() %>% 
    as.numeric()
  
  num_genes_contained = parsed_modules %>% 
    group_by(module) %>% 
    count() %>% 
    arrange(-n) %>% 
    filter(n >= 5) %>% 
    ungroup() %>% 
    summarise(sum = sum(n)) %>% 
    as.numeric()
  
  cbind(num_module_5, num_genes_contained) %>% 
    as.data.frame()

}
```

Here I wrote a function to detect module, pull out number of modules that have >= 5 genes, and count number of genes contained in modules that have >= 5 genes. All in one function. 

Then I can test a list of resolutions in this function. 
Let's test a range of resolution from 0.25 to 5, in steps of 0.25.  
```{r}
optimization_results <- purrr::map_dfr(
  .x = seq(from = 0.25, to = 5, by = 0.25),
  .f = optimize_resolution, 
  network = my_network
) %>% 
  cbind(
   resolution = seq(from = 0.25, to = 5, by = 0.25)
  ) %>% 
  as.data.frame() %>% 
  rename(num_module = num_module_5,
         num_contained_gene = num_genes_contained)

head(optimization_results)
```

This could take a while. 
We have the results organized into one tidy data table. We can graph it. 
```{r}
Optimize_num_module <- optimization_results %>% 
  ggplot(aes(x = resolution, y = num_module)) +
  geom_line(linewidth = 1.1, color = "dodgerblue2") +
  geom_point(size = 3) +
  geom_vline(xintercept = 2, size = 1, linetype = 4) +
  labs(x = "resolution parameter",
       y = "num. modules\nw/ >=5 genes") +
  theme_classic() +
  theme(
    text = element_text(size = 14),
    axis.text = element_text(color = "black")
  )

Optimize_num_gene <- optimization_results %>% 
  ggplot(aes(x = resolution, y = num_contained_gene)) +
  geom_line(linewidth = 1.1, color = "violetred2") +
  geom_point(size = 3) +
  geom_vline(xintercept = 2, size = 1, linetype = 4) +
  labs(x = "resolution parameter",
       y = "num. genes in\nmodules w/ >=5 genes") +
  theme_classic() +
  theme(
    text = element_text(size = 14),
    axis.text = element_text(color = "black")
  )

wrap_plots(Optimize_num_module, Optimize_num_gene, nrow = 2)

ggsave("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/Optimize_resolution.svg", height = 5, width = 3.2, bg ="white")
ggsave("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/Optimize_resolution.png", height = 5, width = 3.2, bg ="white")
```

You can see that there is a big jump for num. modules w/ >= 5 genes going from # to # resolution.
The number of modules stabilizes at resolution >=#
However, if you look at number of contained genes, the story is a little different. 
The number of contained genes is very stable until resolution > 2.5, after which the number of genes continues to diminish. 

> In this case going with res. = 2

### Graph based clustering
The next step is detect modules from the graph object. 
```{r}
modules <- cluster_leiden(my_network, resolution = 2, 
                          objective_function = "modularity")

```

`cluster_leiden()` runs the Leiden algorithm for you. 
`resolution_parameter` controls how many clusters you will get. The larger it is, the more clusters. 
You can play around with the resolution and see what you get. 
The underlying math of `objective_function` is beyond me, but it specifies how the modules are computed. 

Let's say we move on with module detection using a resolution of # 
Next, we need to link the module membership to the gene IDs.
```{r}
my_network_modules <- data.frame(
  gene_ID = names(membership(modules)),
  module = as.vector(membership(modules)) 
) %>% 
  inner_join(node_table, by = "gene_ID")

my_network_modules %>% 
  group_by(module) %>% 
  count() %>% 
  arrange(-n) %>% 
  filter(n >= 5)

my_network_modules %>% 
  group_by(module) %>% 
  count() %>% 
  arrange(-n) %>% 
  filter(n >= 5) %>% 
  ungroup() %>% 
  summarise(sum = sum(n))

```

Looks like there are ~16 modules that have 5 or more genes, comprising ~75469 genes.

Not all genes are contained in modules. They are just lowly connected genes. 

Note that Leiden clustering has a stochastic aspect. The membership maybe slightly different every time you run it. 
Moving forward we will only use modules that have 5 or more genes. 

```{r}
module_5 <- my_network_modules %>% 
  group_by(module) %>% 
  count() %>% 
  arrange(-n) %>% 
  filter(n >= 5)

my_network_modules <- my_network_modules %>% 
  filter(module %in% module_5$module) %>%
  select(gene_ID, module)

head(my_network_modules)
#View(my_network_modules)

saveRDS(my_network_modules, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/my_network_modules_5_genes.rds")
```

```{r}
my_network_modules <- readRDS(file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/my_network_modules_5_genes.rds")
```

### Module quality control
We have a bunch of different modules now, how do we know if they make any sense? 
One way to QC these modules is looking at our bait genes. 
#Filter edge table and save to hard drive
```{r}
#from, to, r, t, p.value, FDR, from func. anno, to func. anno, from module, to module, pathway gene from, pathway gene to, 
head(edge_table_select)
head(mitr_v1.funct_anno)
Baits2 <- Baits %>% select(gene_ID, order, tag)

edge_table_select_filtered <- edge_table_select %>%
  filter(from %in% Baits$gene_ID | to %in% Baits$gene_ID) #%>%
dim(edge_table_select_filtered) #7936

edge_table_select_filtered2 <- edge_table_select2 %>%
  filter(from %in% Baits$gene_ID | to %in% Baits$gene_ID) #%>%
dim(edge_table_select_filtered2) #21349
```


```{r}
edge_table_select_p0.01_annotated <- edge_table_select_filtered %>%
  left_join(mitr_v1.funct_anno, join_by("from"=="gene_ID")) %>%
  rename(from_annotation = functional_annotation) %>%
  rename(from_chr = chr)%>%
  rename(from_start = start)%>%
  rename(from_end = end)%>%
  rename(from_strand = strand) %>%
  left_join(Baits2, join_by("from"=="gene_ID")) %>%
  left_join(my_network_modules, join_by("from"=="gene_ID")) %>%
  rename(from_module = module) %>%
  rename(from_pathway = tag) %>%
#  select(from, from_annotation, from_module, from_pathway, to, r, t, p.value, FDR) %>%
  left_join(mitr_v1.funct_anno, join_by("to"=="gene_ID")) %>%
  rename(to_annotation = functional_annotation) %>%
  rename(to_chr = chr)%>%
  rename(to_start = start)%>%
  rename(to_end = end)%>%
  rename(to_strand = strand) %>%
  left_join(Baits2, join_by("to"=="gene_ID")) %>%
  rename(to_pathway = tag) %>%
  left_join(my_network_modules, join_by("to"=="gene_ID")) %>%
  rename(to_module = module) %>%
  select(from, from_annotation, from_module, from_pathway, to, to_annotation, to_module, to_pathway, r, t, p.value, FDR, from_chr, from_start, from_end, from_strand, to_chr, to_start, to_end, to_strand)

write_excel_csv(edge_table_select_p0.01_annotated, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/edge_table_select_p0.01_pathway_genes_28July25.csv", col_names = T)
edge_table_select_p0.01_annotated <- read_delim(file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/edge_table_select_p0.01_pathway_genes_28July25.csv", delim = ",")

edge_table_select_p0.05_annotated <- edge_table_select_filtered2 %>%
  left_join(mitr_v1.funct_anno, join_by("from"=="gene_ID")) %>%
  rename(from_annotation = functional_annotation) %>%
  rename(from_chr = chr)%>%
  rename(from_start = start)%>%
  rename(from_end = end)%>%
  rename(from_strand = strand) %>%
  left_join(Baits2, join_by("from"=="gene_ID")) %>%
  left_join(my_network_modules, join_by("from"=="gene_ID")) %>%
  rename(from_module = module) %>%
  rename(from_pathway = tag) %>%
#  select(from, from_annotation, from_module, from_pathway, to, r, t, p.value, FDR) %>%
  left_join(mitr_v1.funct_anno, join_by("to"=="gene_ID")) %>%
  rename(to_annotation = functional_annotation) %>%
  rename(to_chr = chr)%>%
  rename(to_start = start)%>%
  rename(to_end = end)%>%
  rename(to_strand = strand) %>%
  left_join(Baits2, join_by("to"=="gene_ID")) %>%
  rename(to_pathway = tag) %>%
  left_join(my_network_modules, join_by("to"=="gene_ID")) %>%
  rename(to_module = module) %>%
  select(from, from_annotation, from_module, from_pathway, to, to_annotation, to_module, to_pathway, r, t, p.value, FDR, from_chr, from_start, from_end, from_strand, to_chr, to_start, to_end, to_strand)

write_excel_csv(edge_table_select_p0.05_annotated, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/edge_table_select_p0.05_pathway_genes_24July25.csv", col_names = T)
edge_table_select_p0.05_annotated <- read_delim(file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/edge_table_select_p0.05_pathway_genes_24July25.csv", delim = ",")
```


```{r}
head(Baits)
(Iridoids <- my_network_modules %>% 
  filter(gene_ID == "Mitsp.v1.21-2G006690" | #Iridoid MsLAMT, first switch of clusters from MEP
           gene_ID == "Mitsp.v1.16-2G020400" | #SLTr
           gene_ID == "Mitsp.v1.21-1G011770" | #STR
           gene_ID == "Mitsp.v1.12-1G013480")) #STRtr

(GOI_1 <- my_network_modules %>% 
  filter(gene_ID == "Mitsp.v1.10-1G015530" | ##MsDCS2 (MIA)
           gene_ID == "Mitsp.v1.10-1G015520" | ##MsDCS1 (MIA)
           gene_ID == "Mitsp.v1.18-2G016780" | ##MsSAS (MIA)
           gene_ID == "Mitsp.v1.10-2G022130" | ##MsCO1.1 (MIA)
           gene_ID == "Mitsp.v1.10-1G021390" | ##MsCO1.2 (MIA)
           gene_ID == "Mitsp.v1.10-1G021550" | ##MsCO1.3 (MIA)
           gene_ID == "Mitsp.v1.08-2G007110" | ##Ms10H
           gene_ID == "Mitsp.v1.10-1G010920" | ##MsDCR1
           gene_ID == "Mitsp.v1.10-2G012340")) ##MsDCR2

(bait_0_17 <- my_network_modules %>%
    filter(gene_ID == "Mitsp.v1.18-2G016780" | #MsSAS 
             gene_ID == "Mitsp.v1.10-1G010920" | #MsDCR1
             gene_ID == "Mitsp.v1.10-2G012340" |#MsDCR2
gene_ID == "Mitsp.v1.22-2G007110" | #Ms10H
gene_ID == "Mitsp.v1.10-2G022130" |# ~ "MsCO1.1",
gene_ID == "Mitsp.v1.10-1G021390" | # ~ "MsCO1.2",
gene_ID == "Mitsp.v1.10-1G021550")) # ~ "MsCO1.3")


(bait_21 <- my_network_modules %>%
  filter(gene_ID == "Mitsp.v1.21-1G011770" |  # "STR"
  gene_ID == "Mitsp.v1.10-1G015530" |  # ~ "MsDCS1",
  gene_ID == "Mitsp.v1.10-1G015520" |  # ~ "MsDCS2",
  gene_ID == "Mitsp.v1.02-1G013220" |  # ~ "MsSGD",
  gene_ID == "Mitsp.v1.11-1G000250"
  ))#MsEnolMT



```

If they are highly correlated, then they are in the same module

## Module-treatment correspondance
The next key task is understanding the expression pattern of the clusters. 
Again, the essence of this workflow is simple, so we will use a simple method: peak expression.
To do that, we append the module membership data back to the long table containing z scores. 

```{r}
Exp_table_long_averaged_z_modules <- Exp_table_long_averaged_z %>% 
  inner_join(my_network_modules, by = "gene_ID") %>%
  left_join(mitr_v1.funct_anno, join_by(gene_ID))
  
head(mitr_v1.funct_anno)
head(Exp_table_long_averaged_z_modules)
#View(Exp_table_long_averaged_z_modules)
```

```{r}
#Annotate clusters by cell type
Exp_table_long_averaged_z_modules <- Exp_table_long_averaged_z_modules %>%
    mutate(cell_type = case_when(
    cluster == 0 | 
      cluster == 2 | 
      cluster == 4 |
      cluster == 6 | 
      cluster == 12 ~ "Mesophyll",
    cluster == 1 | 
      cluster == 8 |
      cluster == 10 | 
      cluster == 16 ~ "Epidermis",
      cluster == 3 |
      cluster == 5 |
      cluster == 15  |
      cluster == 17 |
      cluster == 21  ~ "Vasculature",
    cluster == 9 |
      cluster == 11 ~ "Proliferating",
    cluster == 19 ~ "IPAP",
      cluster == 13 ~ "Guard cells",
    T ~ "Unassigned"))
    
Exp_table_long_averaged_z_modules

#--_excel_csv(Exp_table_long_averaged_z_modules, "/scratch/ac05869/MIT_KRT_sc_analysis/coexp_out/Exp_table_long_averaged_z_modules_r0.7_3res.csv")
```

Now we can produce summary statistics for each cluster and look at their expression pattern using mean. 
```{r}
modules_mean_z <- Exp_table_long_averaged_z_modules %>% 
  group_by(module, cluster, cell_type) %>% 
  summarise(mean.z = mean(z.score)) %>% 
  ungroup()

head(modules_mean_z)
#View(modules_mean_z)

modules_mean_z2 <- Exp_table_long_averaged_z_modules %>% 
  group_by(cluster, module, cell_type) %>% 
  summarise(mean.z = mean(z.score)) %>% 
  ungroup()

head(modules_mean_z2)
#View(modules_mean_z2)
```

Then we look at at which cluster is each module most highly expressed. 
```{r}
module_peak_exp <- modules_mean_z %>% 
  group_by(cluster) %>% 
  group_by(module) %>% #added this because some modules did not have peak expression when only grouped by cluster
  slice_max(order_by = mean.z, n = 1)

module_peak_exp
#View(module_peak_exp)
```
Again, `group_by()` is doing a lot of heavy lifting here. 

### More module QC
You can also QC the clusters via a line graph
> Notes

```{r}
ggplot(Exp_table_long_averaged_z_modules, aes(x=cluster, y=z.score)) +
  geom_line(aes(group = gene_ID), color = "grey70") +
  geom_line(
    data = modules_mean_z,
    aes(y = mean.z, group = module), 
   linewidth = 1.1
  ) +
  labs(x = NULL,
       y = "z score") +
  theme_classic() +
  facet_wrap( ~ module, ncol = 4) +
  theme(
    text = element_text(size = 12),
    axis.text = element_text(color = "black"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8),
    panel.spacing = unit(1, "line")
  )

ggsave("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/module_line_plots_2res.svg", height = 4, width = 10, bg = "white")
ggsave("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/module_line_plots_2res.png", height = 4, width = 10, bg = "white")
```

### Heat map representation of clusters 
A good way to present these modules is to make a heat map. 
To make an effective heatmap though, we need to take care of a few things.

* reorder x and y axis
* take care of outliers 

#### Check outliers 
Let's take care of outliers first 
```{r}
modules_mean_z$mean.z %>% summary()

quantile(modules_mean_z$mean.z, 0.95)
```
You can see that the distribution of averaged z scores are more or less symmetrical from the 1st to 3rd quartiles. 

The 95th percentile of averaged z score is 1.447974 . We can probably roughly clipped the z-scores at 1.5 or -1.5

> 1.45, clipped at 1.5

```{r}
modules_mean_z <- modules_mean_z %>% 
  mutate(mean.z.clipped = case_when(
    mean.z > 1.5 ~ 1.5,
    mean.z < -1.5 ~ -1.5,
    T ~ mean.z
  ))

head(modules_mean_z)
```

> This sets z scores > 1.5 or < -1.5 to 1.5 or -1.5, respectively. The rest remain unchanged.  

#### Reorder rows and columns 
Let's say we graph modules on y axis, and stage/tissue (cluster) on x-axis.
Reordering columns are easy, we just do it by hand. 
We already did it before. We can copy and paste that down here. 
```{r}
modules_mean_z2 <- modules_mean_z %>% 
#  mutate(cluster = factor(cluster, levels = c(0:22))) %>%
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll", "Epidermis", "Guard cells",
                              "Vasculature",
                              "Proliferating", "IPAP", "Unassigned"
                            ))) %>% 
  mutate(cluster = factor(cluster, levels = c(
     "0", "2", "4", "6", "12", #M
    "1","8", "10", "16", #Epi
    "3", "5", "15", "17", "21", #vasculature
   "9", "11", #prolif
    "13", #Guard cells
   "19", ##IPAP
    "7", "14", "18", "20" #Unknown
  )))
head(modules_mean_z2)
```

```{r}
my_data_peak_values <- modules_mean_z2 %>% 
  group_by(cluster) %>% #this used to be module
  slice_max(order_by = mean.z, n = 1, with_ties = F) %>% 
  rename(peaked_at = cluster) %>% 
  select(-mean.z, cell_type)

head(my_data_peak_values)

number_of_peaks <- my_data_peak_values %>% 
  group_by(peaked_at) %>% 
  count() %>% 
  arrange(-n)
head(number_of_peaks)

my_data_peak_values_reordered <- my_data_peak_values %>% 
  inner_join(number_of_peaks, by = "peaked_at") %>% 
  arrange(-n)  
my_data_peak_values_reordered

my_data_reordered <- modules_mean_z2 %>% 
  inner_join(number_of_peaks, by = c("cluster" = "peaked_at")) %>% 
  mutate(cluster = reorder(cluster, -n)) %>%  # this reorders the columns
  select(-n) %>% 
  inner_join(my_data_peak_values_reordered, by = "module", multiple = "first") %>% 
  mutate(peaked_at = reorder(peaked_at, -n)) %>% 
  mutate(order_rows = as.numeric(peaked_at)) %>% 
  mutate(module = reorder(module, -order_rows)) # this reorders the rows by the "peaked_at" column. 
head(my_data_reordered)

module_heatmap <- my_data_reordered %>% 
  ggplot(aes(x = cluster, y = module)) +
  geom_tile(aes(fill = mean.z.clipped.x), color = "grey80") +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu")), limits = c(-1.5, 1.5),
                       breaks = c(-1.5, 0, 1.5), labels = c("< -1.5", "0", "> 1.5")) +
  labs(fill = "mean.z") +
  theme_classic() +
  theme(
    text = element_text(size = 16),
    axis.text = element_text(color = "black"),
    axis.text.x = element_text(color = "black"),
    strip.text = element_blank(),
    legend.position = "top",
    panel.spacing = unit(0.5, "lines") 
  )

module_heatmap
```

```{r}
my_data_peak_values2 <- modules_mean_z2 %>% 
  group_by(cluster) %>%
  group_by(module) %>%
  slice_max(order_by = mean.z, n = 1, with_ties = F) %>% 
  rename(peaked_at = cluster) %>% 
  select(-mean.z, cell_type)

head(my_data_peak_values2)

number_of_peaks2 <- my_data_peak_values2 %>% 
  group_by(peaked_at) %>% 
  count() %>% 
  arrange(-n)
head(number_of_peaks2)

my_data_peak_values_reordered2 <- my_data_peak_values2 %>% 
  inner_join(number_of_peaks2, by = "peaked_at") %>% 
  arrange(-n)  
my_data_peak_values_reordered2

my_data_reordered2 <- modules_mean_z2 %>% 
  inner_join(number_of_peaks, by = c("cluster" = "peaked_at")) %>% 
  mutate(cluster = reorder(cluster, -n)) %>%  # this reorders the columns
  select(-n) %>% 
  inner_join(my_data_peak_values_reordered2, by = "module") %>% 
  mutate(peaked_at = reorder(peaked_at, -n)) %>% 
  mutate(order_rows = as.numeric(peaked_at)) %>% 
  mutate(module = reorder(module, -order_rows)) # this reorders the rows by the "peaked_at" column. 
head(my_data_reordered2)

module_heatmap2 <- my_data_reordered2 %>% 
  ggplot(aes(x = cluster, y = module)) +
  geom_tile(aes(fill = mean.z.clipped.x), color = "grey80") +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu")), limits = c(-1.5, 1.5),
                       breaks = c(-1.5, 0, 1.5), labels = c("< -1.5", "0", "> 1.5")) +
  labs(fill = "mean.z") +
  theme_classic() +
  theme(
    text = element_text(size = 14),
    axis.text = element_text(color = "black"),
    axis.text.x = element_text(color = "black"),
    strip.text = element_blank(),
    legend.position = "top",
    panel.spacing = unit(0.5, "lines") 
  )

module_heatmap2

ggsave("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/module_heatmap.svg", height = 4.8, width = 10, bg = "white")
ggsave("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/module_heatmap.png", height = 4.8, width = 10, bg = "white")
```

When the rows and columns are re-ordered, you can trace the signal down the diagonal from upper left to lower right. 

```{r}
leaf_cell_type_assignment <- data.frame(
  cluster = c(0:22)
) %>% 
  mutate(cell_type = case_when(
    cluster == 0 | 
      cluster == 2 | 
      cluster == 4 |
      cluster == 6 | 
      cluster == 12 ~ "Mesophyll",
    cluster == 1 | 
      cluster == 8 |
      cluster == 10 | 
      cluster == 16 ~ "Epidermis",
      cluster == 3 |
      cluster == 5 |
      cluster == 15  |
      cluster == 17 |
      cluster == 21  ~ "Vasculature",
    cluster == 9 |
      cluster == 11 ~ "Proliferating",
    cluster == 19 ~ "IPAP",
      cluster == 13 ~ "Guard cells",
    T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll", "Epidermis", "Guard cells",
                              "Vasculature",
                              "Proliferating", "IPAP", "Unassigned"
                            ))) %>% 
  mutate(cluster = factor(cluster, levels = c(
     "0", "2", "4", "6", "12", #M
    "1","8", "10", "16", #Epi
    "3", "5", "15", "17", "21", #vasculature
   "9", "11", #prolif
    "13", #Guard cells
   "19", ##IPAP
    "7", "14", "18", "20" #Unknown
  )))

Leaf_cell_type_strip <- leaf_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "" )) +
  geom_tile(aes(fill = cell_type)) +
  scale_fill_manual(values = c(brewer.pal(7, "Accent"), "grey80")) +
  guides(fill = guide_legend(nrow = 2)) +
  labs(fill = "Cell type") +
  theme_void() +
  theme(
    legend.position = "bottom" ,
    text = element_text(size = 14)
  )

Leaf_cell_type_strip
graph <- wrap_plots(module_heatmap2, Leaf_cell_type_strip,
                    nrow = 2, heights = c(1, 0.03))
graph

ggsave("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/module_heatmap_celltype.svg", height = 5.3, width = 10, bg = "white")
ggsave("/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/module_heatmap_celltype.png", height = 5.3, width = 10, bg = "white")
```
#Make node table for candidate genes
```{r}
edge_table_filtered2 <- edge_table_select %>%
  filter(from %in% pathway_genes$gene_ID) %>%
  left_join(pathway_genes, by = c("from" = "gene_ID")) %>%
  select("from", "tag", "to", "r", "t", "p.value", "FDR") %>%
  mutate(
    stats = paste0(
      "r = ", round(r, 3),
      "; t = ", signif(t, 3),
      "; p = ", signif(p.value, 3),
      "; FDR = ", signif(FDR, 3))) %>%
  pivot_wider(names_from = tag,
              values_from = stats)%>%
  select(!c("from", "r", "t", "p.value", "FDR")) %>%
  rename("gene_ID" = "to")

edge_table_filtered3 <- edge_table_select %>%
  filter(to %in% pathway_genes$gene_ID) %>%
  left_join(pathway_genes, by = c("to" = "gene_ID")) %>%
  select("from", "tag", "to", "r", "t", "p.value", "FDR") %>%
  mutate(
    stats = paste0(
      "r = ", round(r, 3),
      "; t = ", signif(t, 3),
      "; p = ", signif(p.value, 3),
      "; FDR = ", signif(FDR, 3))) %>%
  pivot_wider(names_from = tag,
              values_from = stats) %>%
  select(!c("to", "r", "t", "p.value", "FDR"))%>%
  rename("gene_ID" = "from")

edge_table_merged <- bind_rows(edge_table_filtered2, edge_table_filtered3) %>%
  group_by(gene_ID) %>%
  summarise(across(everything(), ~ paste(unique(na.omit(.x)), collapse = " | ")), .groups = "drop")

  

mod_8_genes <- my_network_modules %>%
  filter(module == 8) %>%
  right_join(edge_table_merged, by = "gene_ID") %>%
  left_join(repr_anno, by = "gene_ID") %>% 
  select(c(gene, functional_annotation, !gene_ID))
write_excel_csv(mod_8_genes, file = "/scratch/ac05869/10X_Multiome/KRT_leaf/coex_out/mod_8_genes_28July25.csv", col_names = T)
```


